---
title: "Homework N — Short title"
subtitle: "70-467 · individual · notebook + at most six slides + at most five minutes video"
author: "Andrés Castaño Zuluaga"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: right
    body-classes: "mlba-page"
    code-fold: false
    code-tools: true
execute:
  echo: true
  warning: false
  message: false
---

<!--
COPY when writing a real homework (M8 / M11).
Individual (locked 2026-08-15). Four separate cases. No Lending Club
semester machine. New case, same method as the week — not the lab table.

Package: runnable R notebook + at most six slides + a video of at most five minutes. Canvas only.

Approve LOOK before filling any homework.
-->

<!-- HW3 required: cutoff is a decision. Parts must price a false alarm
     vs a miss and say who you flag. Do not treat 0.5 as the answer. -->
<!-- HW4: trees + boosting + PCA if needed.
     Trees stay in scope. Dimension reduction only as covered.
     Do not add extra methods to fill the missing Wednesday. -->

## Due

| | |
|--|--|
| Released | {{date from schedule.qmd}} |
| Due | {{date}} · Canvas · 23:59 |
| Work | Individual |
| Time budget | About one focused evening plus the short video |

## The case

{{One manager question. What should they do this week?}}

You recommend and you state a limit. A model score is not the recommendation.

## Data

We ship the files with this homework. Do not use a private vendor pull. Do not reuse a previous homework's table.

| File | Role |
|------|------|
| `data/{{HW-FILE.csv}}` | The table (when the homework is real) |
| `codebook.md` | Variable, type, meaning, known at decision time? |

## Codebook

See `codebook.md` in this pack. Before you fit anything, mark which columns would be known at the moment of the decision.

## Starter

Setup is filled. The analysis below is incomplete on purpose — you have to decide the split (and, on HW3, the cutoff).

```{r}
#| label: setup
#| message: false
# library(tidyverse)
# dat <- read.csv("data/HW-FILE.csv")

# Placeholder so the starter knits. Replace with the shipped table.
```

```{r}
#| label: incomplete
# This fits on all rows. That is not the answer.
# You decide: how to split, what to keep, what to tell the desk.
# HW3: the cutoff is a choice. HW4: trees + boosting + PCA if needed.
```

## Parts

Numbered. Each part leaves a pick a chatbot cannot finish for you.

### 1.

{{Look at the table in a way that serves the case — not a fishing trip. Name one column you would drop, and why.}}

### 2.

{{Fit the method for *this* homework only. State the one modeling choice you made (split, penalty, cutoff, …).}}

### 3.

{{Honesty check for this homework. HW2: a full cross-validation pipeline is not required. HW3: costs of a false alarm vs a miss; who you flag. HW4: trees + boosting + PCA if needed.}}

### 4.

{{The call. Who you treat differently, and one thing you must not claim.}}

Write the manager paragraph here, not only in the slides.

## Submit

Three pieces, Canvas only.

1. **This notebook**, knitted, all code runnable.
2. **At most six slides:** the problem, the evidence, the decision, and the limit.
3. **≤ 5:00 video** of the same story. Phone recording is fine.

## AI

Allowed as an assistant **with disclosure**. Write what you used and what you changed. You verify every number. You must be able to explain every chart in an oral without the tool in the room.

<!-- Rubric: M2. Do not invent weights in the template. -->
