---
title: "Lab NN — Short title"
subtitle: "70-467 · ~75 minutes · bring a laptop"
author: "Andrés Castaño Zuluaga"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: right
    body-classes: "mlba-page"
    code-fold: false
    code-tools: true
execute:
  echo: true
  warning: false
  message: false
---

<!--
COPY to site/labs/labNN-slug.qmd
E2E-style starter walkthrough (PICKS: Lab A + thinking questions).
Not from-scratch. Not coding-TODO blanks as the hour.
Setup filled. Demo chunks filled. Students tweak and answer Think 1–4.
Default: no Canvas submit for lab.
Default meeting: Wednesday. Exceptions: see schedule.qmd.

Locks:
- Week 3 / 5 / 8 honesty in a Think item or a tweak comment.
- Week 8: cutoff as a cost; who you would treat differently.
- Week 12: clustering + short modern-AI. No market-basket exercise.
- Same question language as the pre-class problem and the notes.
- Homework is a different case.

Approve LOOK before filling any week.
-->

## Setup

```{r}
#| label: setup
#| message: false
# library(tidyverse)
# When this lab is real: shipped file next to the notebook.
# dat <- read.csv("data/WEEK-FILE.csv")

# Placeholder so the starter knits. Replace with this week's table.
n <- 20
y <- c(1:n)
```

Confirm the file reads. Look at the first rows when the csv is real.

## Walkthrough

Same question as the one-pager: {{paste the one sentence}}.

```{r}
#| label: demo-1
# Filled on purpose. Run it. Then stop.
# tweak: change this number once and say what moved
k <- 5
mean(y[seq_len(k)])
```

{{One line: what this chunk is doing for the question. Not a new method.}}

```{r}
#| label: demo-2
# Filled. This is the honest version of the idea above, when the week is real.
# tweak: if this is week 3, 5, or 8 — change the split / fold / cutoff,
#         not the model family
mean(y)
```

## Think 1

{{Interpret what you just ran. What does the number or picture say?}}

Write two or three sentences here.

## Think 2

{{A picture. Sketch it in R or on paper. What should a manager see?}}

```{r}
#| label: think-2
# Hint, not a blank hour. Tweak the plot; do not start from zero.
# plot(y)
```

## Think 3

{{What would you tell the desk? One recommendation and one limit.}}

## Think 4

{{Second interpret or business ask. Week 3 / 5: one sentence on leakage
or "available at decision time." Week 8: pick a cutoff using costs, not
0.5, and say who you treat differently. Week 12: what a neural net /
chatbot is *not*.}}

## Last 10 minutes

Someone says **Think 1** out loud. Then we hear one limit from Think 3.

## After class

{{What "done enough" looks like.}} No Canvas submit for lab unless that week says otherwise. This file is not the homework.
