---
title: "Logistic regression"
author: "Andrés Castaño Zuluaga"
format:
  revealjs:
    theme: simple
    css:
      - ../_theme/mlba-reveal.css
      - session05.css
    slide-number: true
    embed-resources: true
    html-math-method: katex
    width: 1280
    height: 720
    margin: 0.04
    center: false
    auto-stretch: false
    min-scale: 0.2
    max-scale: 1.4
    transition: none
    navigation-mode: linear
    controls-tutorial: false
    hash-type: number
    highlight-style: arrow
    code-copy: false
    code-overflow: wrap
    title-slide-attributes:
      data-background-color: "#182C4B"
      class: "quarto-title-block"
    footer: "70-467 · Logistic regression"
    template-partials:
      - title-slide.html
execute:
  echo: false
  warning: false
  message: false
fig-dpi: 140
---

```{r}
#| label: setup
#| include: false
library(ggplot2)

# ---- File -------------------------------------------------------------------
oj_path_candidates <- c(
  "OJ.csv",
  "../../_project/data_pack/teaching/w05_slides/OJ.csv",
  "../../../_project/data_pack/teaching/w05_slides/OJ.csv",
  "../../_project/data_pack/teaching/backups/OJ.csv",
  "../../../_project/data_pack/teaching/backups/OJ.csv"
)
oj_path <- oj_path_candidates[file.exists(oj_path_candidates)][1]
if (is.na(oj_path)) stop("OJ.csv not found next to the deck or in the data pack")

oj <- read.csv(oj_path, stringsAsFactors = FALSE)
# Some exports keep an unnamed row-id in the first column. Drop it.
if (!names(oj)[1] %in% c("Purchase", "LoyalCH", "PriceDiff")) {
  oj <- oj[, -1, drop = FALSE]
}
oj$Purchase <- factor(oj$Purchase, levels = c("CH", "MM"))
oj$y <- as.integer(oj$Purchase == "CH")

n <- nrow(oj)
n_ch <- sum(oj$y)
n_mm <- n - n_ch
share_ch <- mean(oj$y)

# ---- Plot theme (same ink as Session 04) ------------------------------------
weaver <- "#182C4B"; red <- "#C41230"; iron <- "#6D6E71"; paper <- "#F4F6F8"; teal <- "#008F91"
theme_mlba <- function(base = 15) {
  theme_minimal(base_family = "sans", base_size = base) +
    theme(
      plot.background = element_rect(fill = paper, color = NA),
      panel.background = element_rect(fill = paper, color = NA),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "#E0E0E0", linewidth = 0.3),
      axis.title = element_text(color = iron, size = base - 2),
      axis.text = element_text(color = weaver, size = base - 3),
      legend.position = "none",
      plot.margin = margin(6, 10, 4, 4),
      plot.caption = element_text(color = iron, size = base - 5, hjust = 0),
      strip.text = element_text(color = weaver, face = "bold", size = base - 2),
      strip.background = element_rect(fill = "#E0E0E0", color = NA)
    )
}
brand_fill <- c(CH = "#D6DEE8", MM = "#E4E5E7")

# ---- Locked teaching fits: CH = 1, MM = 0 -----------------------------------
fit_lin <- lm(y ~ LoyalCH, data = oj)
lin_hat <- predict(fit_lin)
lin_min <- min(lin_hat)
lin_max <- max(lin_hat)

fit1 <- glm(y ~ LoyalCH, data = oj, family = binomial)
fit2 <- glm(y ~ LoyalCH + PriceDiff, data = oj, family = binomial)

oj$loyal_bin <- cut(
  oj$LoyalCH,
  breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),
  include.lowest = TRUE,
  right = FALSE
)
bin_df <- data.frame(
  bin = levels(oj$loyal_bin),
  n = as.integer(tapply(oj$y, oj$loyal_bin, length)),
  rate = as.numeric(tapply(oj$y, oj$loyal_bin, mean)),
  stringsAsFactors = FALSE
)
bin_df$label <- sprintf("%.2f", bin_df$rate)

loyal_means <- tapply(oj$LoyalCH, oj$Purchase, mean)
grid_loyal <- data.frame(LoyalCH = seq(0, 1, length.out = 240))
grid_loyal$p_hat <- as.numeric(predict(fit1, newdata = grid_loyal, type = "response"))
grid_loyal$lin <- as.numeric(predict(fit_lin, newdata = grid_loyal))

new3 <- data.frame(
  LoyalCH = c(0.5, 0.5, 0.9),
  PriceDiff = c(0.2, -0.2, 0)
)
new3$p_hat <- as.numeric(predict(fit2, newdata = new3, type = "response"))
new3$who <- factor(
  c("LoyalCH 0.5\nPriceDiff +0.2", "LoyalCH 0.5\nPriceDiff −0.2", "LoyalCH 0.9\nPriceDiff 0"),
  levels = c("LoyalCH 0.5\nPriceDiff +0.2", "LoyalCH 0.5\nPriceDiff −0.2", "LoyalCH 0.9\nPriceDiff 0")
)
# Rounded glass coefficients. Steps below use these, so the arithmetic matches the board.
b1_glass <- round(coef(fit1), 2)
b2_glass <- round(coef(fit2), 2)
odds_loyal_step <- exp(0.1 * unname(b1_glass[2]))
odds_price_step <- exp(0.10 * unname(b2_glass[3]))
eta_hand <- unname(b2_glass[1] + b2_glass[2] * 0.5 + b2_glass[3] * 0.2)
p_hand <- as.numeric(plogis(eta_hand))
p_glass1 <- function(x) {
  plogis(unname(b1_glass[1]) + unname(b1_glass[2]) * x)
}
p_lo <- p_glass1(0.32)
p_lo_step <- p_glass1(0.42)
p_hi <- p_glass1(0.72)
p_hi_step <- p_glass1(0.82)
p_pd0 <- as.numeric(predict(
  fit2, newdata = data.frame(LoyalCH = 0.5, PriceDiff = 0), type = "response"
))
p_pd1 <- as.numeric(predict(
  fit2, newdata = data.frame(LoyalCH = 0.5, PriceDiff = 0.1), type = "response"
))
odds_file <- round(share_ch, 2) / (1 - round(share_ch, 2))

price_grid <- data.frame(
  LoyalCH = 0.5,
  PriceDiff = seq(-0.4, 0.55, length.out = 180)
)
price_grid$p_hat <- as.numeric(predict(fit2, newdata = price_grid, type = "response"))

glass_curve <- data.frame(LoyalCH = seq(0, 1, length.out = 240))
glass_curve$p_hat <- p_glass1(glass_curve$LoyalCH)

r2 <- function(x) sprintf("%.2f", x)
r3 <- function(x) sprintf("%.3f", x)

# Hard lock. If a future OJ export drifts, stop rather than teach a new number.
stopifnot(n == 1070L, n_ch == 653L, n_mm == 417L)
stopifnot(r2(share_ch) == "0.61")
stopifnot(sprintf("%.2f", odds_file) == "1.56")
stopifnot(r2(loyal_means["CH"]) == "0.72", r2(loyal_means["MM"]) == "0.32")
stopifnot(r3(lin_min) == "0.036", r3(lin_max) == "1.051")
stopifnot(all(bin_df$label == c("0.12", "0.28", "0.53", "0.77", "0.96")))
stopifnot(r2(coef(fit1)[1]) == "-2.76", r2(coef(fit1)[2]) == "6.09")
stopifnot(r2(coef(fit2)[1]) == "-3.25", r2(coef(fit2)[2]) == "6.40", r2(coef(fit2)[3]) == "2.86")
stopifnot(all(r2(new3$p_hat) == c("0.63", "0.35", "0.92")))
stopifnot(r2(odds_loyal_step) == "1.84")
stopifnot(r2(odds_price_step) == "1.33")
stopifnot(r3(eta_hand) == "0.522")
stopifnot(r2(p_hand) == "0.63")
stopifnot(r2(p_lo) == "0.31", r2(p_lo_step) == "0.45")
stopifnot(r2(p_hi) == "0.84", r2(p_hi_step) == "0.90")
stopifnot(r2(p_pd0) == "0.49", r2(p_pd1) == "0.56")
```

## Where Cas and Abdullah left us

```{=html}
<div class="pair-card">
  <p><strong>Cas</strong> and <strong>Abdullah</strong> opened logistic regression on Monday.</p>
  <p>Today we continue — same topic, one store file, one decision.</p>
  <p>We will not re-teach their intro from a cold start.</p>
</div>
```

::: {.notes}
Bridge only. Thank them by name. One sentence on what they opened, then the grid and the OJ file.

Do not steal their examples onto later slides. Do not put the Monday date on the title. Book pages stay in these notes: ISLR 2nd ed. §4.1–4.3 (~pp. 129–141). Logistics stay in your mouth or the final next-slide.
:::

## Where are we on the grid?

<p class="q-prompt">Last week we shrank slopes for a dollar balance. What changes when the thing we want to predict is a <strong>yes/no</strong> at the shelf?</p>

```{=html}
<div class="grid-wrap">
<table class="mlba-grid">
  <colgroup>
    <col class="topic" /><col class="desc" /><col class="tools" />
  </colgroup>
  <thead>
    <tr><th>Topic</th><th>Description</th><th>Tools</th></tr>
  </thead>
  <tbody>
    <tr class="band"><td colspan="3">Fit &amp; evaluate</td></tr>
    <tr class="done"><td>Linear regression</td><td>Predict a number</td><td>OLS</td></tr>
    <tr class="done"><td>Metrics</td><td>Size of the errors</td><td>RMSE, R²</td></tr>
    <tr class="done"><td>Train/test, leakage, cross-validation</td><td>Error on new data</td><td>Holdout, k-fold</td></tr>
    <tr class="done"><td>Selection + regularization</td><td>Select variables or shrink coefficients</td><td>Best subset, ridge, lasso</td></tr>
    <tr class="band"><td colspan="3">Classify &amp; decide</td></tr>
    <tr class="here"><td>Logistic</td><td>Predict a yes/no outcome</td><td>Logistic regression</td></tr>
    <tr><td>Thresholds</td><td>Turn a probability into a decision</td><td>ROC, cutoff</td></tr>
    <tr class="band"><td colspan="3">Flexible &amp; unsupervised</td></tr>
    <tr><td>Trees</td><td>Split the predictors in sequence</td><td>Decision trees</td></tr>
    <tr><td>Forests / boosting</td><td>Combine many trees</td><td>Random forest, boosting</td></tr>
    <tr><td>PCA</td><td>Fewer combinations of the predictors</td><td>Principal components</td></tr>
    <tr><td>Clustering</td><td>Group similar observations</td><td>K-means</td></tr>
  </tbody>
</table>
</div>
```

<p class="caption">The marked cell is this hour: a probability of a brand choice. Thresholds stay later. Do not reopen ridge or lasso.</p>

::: {.notes}
Open block — no click. The Logistic row is already marked. You name “probability of a brand choice.”

The Thresholds cell is the later week on the same roadmap. Do not open it. No confusion matrix. ISLR Ch. 4 starts classification here; we stop at $\hat p$.
:::

## Can we predict which OJ brand a shopper buys?

:::: {.columns}
::: {.column .spaced-list width="52%"}
- One row = one store purchase of orange juice ($n$ = `r format(n, big.mark = ",")`).
- $Y$: `Purchase` — **CH** (Citrus Hill) or **MM** (Minute Maid).
- Measured mix: CH `r n_ch` (`r sprintf("%.1f%%", 100 * n_ch / n)`) / MM `r n_mm` (`r sprintf("%.1f%%", 100 * n_mm / n)`). No missing values.
- Decision this hour: a **new** visit is about to happen. What is the chance they pick Citrus Hill?
:::
::: {.column width="48%"}
```{r}
#| fig-width: 5.6
#| fig-height: 3.55
#| fig-alt: "Bar chart of 653 Citrus Hill purchases and 417 Minute Maid purchases."
brand_n <- data.frame(
  Purchase = factor(c("CH", "MM"), levels = c("CH", "MM")),
  n = c(n_ch, n_mm)
)
ggplot(brand_n, aes(Purchase, n, fill = Purchase)) +
  geom_col(width = 0.62) +
  geom_text(aes(label = format(n, big.mark = ",")), vjust = -0.35,
            fontface = "bold", color = weaver, size = 5) +
  scale_fill_manual(values = c(CH = weaver, MM = iron)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.14)), limits = c(0, 820)) +
  labs(x = NULL, y = "Purchases") +
  theme_mlba(15)
```
:::
::::

<p class="q-prompt">If headquarters only said “most people buy CH,” is that enough for <strong>this</strong> shopper, with a loyalty score and a price gap?</p>

::: {.notes}
Sit with the question. The answer comes back on the remember slide. Do not say the file name on the title.

The CH share is about 0.61. Name it as the mix in the file, not as a classification score. ISLR 4.1: a qualitative response. OJ only.
:::

## What is in the file?

```{=html}
<table class="cb">
  <thead>
    <tr><th>Variable</th><th>Type</th><th>Unit</th><th>Description</th></tr>
  </thead>
  <tbody>
    <tr><td>Purchase</td><td>CH / MM</td><td>brand</td><td>Which OJ they bought (Y)</td></tr>
    <tr><td>LoyalCH</td><td>numeric</td><td>0–1 score</td><td>Customer loyalty toward Citrus Hill</td></tr>
    <tr><td>PriceDiff</td><td>numeric</td><td>dollars</td><td>Sale price MM − sale price CH</td></tr>
    <tr><td>SalePriceCH</td><td>numeric</td><td>dollars</td><td>CH shelf / sale price</td></tr>
    <tr><td>SalePriceMM</td><td>numeric</td><td>dollars</td><td>MM shelf / sale price</td></tr>
  </tbody>
</table>
```

<p class="caption">Workhorses this hour: <strong>LoyalCH</strong> and <strong>PriceDiff</strong>. List prices, discounts, special-display flags, and store ids stay in the file. We will not use them today.</p>

::: {.notes}
Point at LoyalCH and PriceDiff. If asked: PriceCH, PriceMM, DiscCH, DiscMM, SpecialCH, SpecialMM, StoreID, STORE, Store7, PctDiscCH, PctDiscMM, ListPriceDiff. WeekofPurchase stays off unless asked.
:::

## What we will do today

<ol class="today-list">
<li>Why a straight line fails for a brand choice, on this file</li>
<li>The curve, the odds, and what a coefficient multiplies</li>
<li>Fit loyalty, then the price gap with loyalty held fixed</li>
<li>Score one new visit — a probability — and stop</li>
</ol>

::: {.notes}
Four lines. No book section numbers on the glass. No cutoff, no hit/miss table, no other dataset. The closing slide names the lab and the midterm; this slide does not.
:::

## Brand choice is a probability, not a line {.section-start}

<p class="act-kicker">This file</p>

<p class="setup-line">We want the chance a shopper buys Citrus Hill. A straight line was built for a number that can sit anywhere — a dollar balance, not a brand.</p>

::: {.notes}
Section start: weaver band, one sentence, then the OJ pictures. Not a blank dark slide.

ISLR ~pp. 129–133: a binary response, then why not least squares on 0/1. OJ only.
:::

## Example on OJ — loyal shoppers and CH

:::: {.columns}
::: {.column .spaced-list width="42%"}
- Mean `LoyalCH` ≈ **`r r2(loyal_means["CH"])`** among CH buyers.
- Mean `LoyalCH` ≈ **`r r2(loyal_means["MM"])`** among MM buyers.
- Loyalty already separates the two brands. We still need a probability.

```{r}
#| echo: true
round(tapply(oj$LoyalCH, oj$Purchase, mean), 2)
```
:::
::: {.column width="58%"}
::: {.fragment}
```{r}
#| fig-width: 6.4
#| fig-height: 3.7
#| fig-alt: "Boxplots of LoyalCH for CH and MM buyers, with means 0.72 and 0.32."
means_df <- data.frame(
  Purchase = factor(c("CH", "MM"), levels = c("CH", "MM")),
  LoyalCH = as.numeric(loyal_means[c("CH", "MM")])
)
ggplot(oj, aes(Purchase, LoyalCH)) +
  geom_boxplot(aes(fill = Purchase), width = 0.55, color = weaver,
               outlier.size = 0.8, outlier.alpha = 0.35) +
  geom_point(data = means_df, color = red, size = 3.2, shape = 18) +
  geom_text(
    data = means_df,
    aes(label = sprintf("mean %.2f", LoyalCH)),
    color = red, fontface = "bold", size = 4.2, nudge_x = 0.28, nudge_y = 0.04
  ) +
  scale_fill_manual(values = brand_fill) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(x = NULL, y = "LoyalCH") +
  theme_mlba(14)
```
:::
:::
::::

::: {.fragment}
<p class="q-prompt">Before any formula: which shopper would you bet on for CH — loyalty near 0.3, or near 0.7?</p>
:::

::: {.notes}
Claim first, then the boxplots. The two means are the picture: 0.72 versus 0.32.

Association only. PriceDiff also runs higher among CH buyers — do not quote a new mean. No cause.
:::

## Example on OJ — a fitted “chance” of 1.05

:::: {.columns}
::: {.column width="44%"}
- Treat $Y$ like a dollar amount: `lm(y ~ LoyalCH)`.
- On **this** file the fitted values run from about **`r r3(lin_min)`** to about **`r r3(lin_max)`**.
- A number above 1, or below 0, is not a chance.

```{r}
#| echo: true
fit_lin <- lm(y ~ LoyalCH, oj)
range(round(predict(fit_lin), 3))
```
:::
::: {.column width="56%"}
::: {.fragment}
```{r}
#| fig-width: 6.4
#| fig-height: 3.85
#| fig-alt: "Least-squares line crossing the fences at 0 and 1. The fitted values run from about 0.036 to about 1.051."
fence_lab <- data.frame(
  LoyalCH = c(0.18, 0.72),
  y = c(lin_min + 0.10, 1.08),
  lab = c(sprintf("min ≈ %.3f", lin_min), sprintf("max ≈ %.3f", lin_max))
)
ggplot(oj, aes(LoyalCH, y)) +
  geom_jitter(height = 0.035, width = 0, alpha = 0.22, color = weaver, size = 1.25) +
  geom_line(data = grid_loyal, aes(LoyalCH, lin), color = red, linewidth = 1.05) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed", color = iron, linewidth = 0.4) +
  geom_text(data = fence_lab, aes(label = lab), color = red, fontface = "bold", size = 4) +
  scale_y_continuous(breaks = c(0, 0.5, 1), limits = c(-0.08, 1.2)) +
  labs(x = "LoyalCH", y = "Least-squares fit") +
  theme_mlba(13)
```
:::
:::
::::

::: {.fragment}
<p class="q-prompt">Would you report “`r r3(lin_max)` chance of buying CH” to the store?</p>
:::

::: {.notes}
Claim first: 0.036 to 1.051, measured on this file. Then the picture of the line leaving the fences.

The class says no. That is the why-not-linear beat, in OJ. Do not read a new metric off the plot.
:::

## Example on OJ — CH share by loyalty bin

:::: {.columns}
::: {.column width="40%"}
The CH rate climbs with loyalty. The steps are not equal.

```{=html}
<table class="cb odds-mini">
  <thead><tr><th>LoyalCH</th><th>CH rate</th></tr></thead>
  <tbody>
```
```{r}
#| results: asis
cat(paste0(
  "<tr><td>", bin_df$bin, "</td><td>", bin_df$label, "</td></tr>",
  collapse = "\n"
))
```
```{=html}
  </tbody>
</table>
```

```{r}
#| echo: true
round(tapply(oj$y, oj$loyal_bin, mean), 2)
```
:::
::: {.column width="60%"}
::: {.fragment}
```{r}
#| fig-width: 6.6
#| fig-height: 3.9
#| fig-alt: "Bar chart of the Citrus Hill rate in five LoyalCH bins: about 0.12, 0.28, 0.53, 0.77, 0.96."
ggplot(bin_df, aes(bin, rate)) +
  geom_col(width = 0.72, fill = weaver) +
  geom_text(aes(label = label), vjust = -0.35, fontface = "bold", color = weaver, size = 4.4) +
  scale_y_continuous(limits = c(0, 1.12), breaks = seq(0, 1, 0.25), expand = c(0, 0)) +
  labs(x = "LoyalCH bin", y = "Share who bought CH") +
  theme_mlba(13)
```
:::
:::
::::

::: {.fragment}
<p class="q-prompt">Is the step from the first bin to the second the same size as the step near the top?</p>
:::

::: {.notes}
Claim is the five rates: 0.12, 0.28, 0.53, 0.77, 0.96. Then the bars.

You say: a straight line cannot tell this climb. Steeper through the middle, flatter near the ends. That is the shape we fit next.
:::

## The curve stays between 0 and 1 {.section-start}

:::: {.columns}
::: {.column width="44%"}
For a score $\eta$, the logistic function is a probability:

::: {.formula}
$$
p = \frac{e^{\eta}}{1 + e^{\eta}}, \qquad 0 < p < 1.
$$
:::

$\eta$ will be a line in the shopper’s features. $p$ is $\mathrm{P}(\mathrm{CH})$ for that shopper.
:::
::: {.column width="56%"}
::: {.r-stack}
::: {.fragment .fade-in-then-out}
<p class="caption">The function. It cannot cross 0 or 1.</p>
```{r}
#| fig-width: 6.3
#| fig-height: 3.35
#| fig-alt: "A smooth curve of fitted P(CH) against LoyalCH, staying between 0 and 1."
ggplot(grid_loyal, aes(LoyalCH, p_hat)) +
  geom_line(color = weaver, linewidth = 1.25) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed", color = iron, linewidth = 0.35) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
  labs(x = "LoyalCH", y = "P(CH)") +
  theme_mlba(13)
```
:::
::: {.fragment .fade-in}
<p class="caption">Same curve, with this file’s purchases.</p>
```{r}
#| fig-width: 6.3
#| fig-height: 3.35
#| fig-alt: "The same fitted curve with each purchase drawn at 0 or 1."
ggplot(oj, aes(LoyalCH, y)) +
  geom_jitter(height = 0.04, width = 0, alpha = 0.22, color = iron, size = 1.15) +
  geom_line(data = grid_loyal, aes(LoyalCH, p_hat), color = weaver, linewidth = 1.25) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed", color = iron, linewidth = 0.35) +
  scale_y_continuous(limits = c(-0.08, 1.08), breaks = c(0, 0.5, 1)) +
  labs(x = "LoyalCH", y = "Purchase (0/1) and P(CH)") +
  theme_mlba(13)
```
:::
:::
:::
::::

::: {.notes}
Section start with the function, then the picture. First click: the curve alone. Second click: the 0/1 purchases from this file laid on it. That is the ISLR Figure 4.2 move, on OJ.

The curve is `glm(y ~ LoyalCH)` on this file. Do not read the coefficients yet — they are the next fit. Do not name another dataset.
:::

## Odds come out of that function

- If $p = \mathrm{P}(\mathrm{CH})$, the odds of CH are $p / (1 - p)$.
- Substitute the logistic function. The odds equal $e^{\eta}$.
- We still report $p$. The odds are the scale the line is written on.

::: {.formula}
$$
\frac{p}{1-p} = e^{\eta}.
$$
:::

::: {.fragment}
<p class="q-prompt">If $\eta$ rises by $\log 2$, what happens to the odds of CH?</p>
:::

::: {.fragment}
<p class="today-note">They double. $p$ does not double. The next slides separate those two.</p>
:::

::: {.notes}
Algebra only. No new OJ metric on this slide. The file’s own odds are the ladder after this.

ISLR ~pp. 133–136: logistic function, then odds. Glass stays in CH / MM.
:::

## This file’s share, on the odds ladder

<p class="setup-line">A probability of 0.61 is not odds of 0.61. On this file the CH share is about **`r r2(share_ch)`**, so the odds of CH are about **`r sprintf("%.2f", odds_file)`**.</p>

```{=html}
<table class="cb">
  <thead>
    <tr><th>p = P(CH)</th><th>Odds p / (1 − p)</th><th>In words</th></tr>
  </thead>
  <tbody>
    <tr><td>0.25</td><td>0.33</td><td>one to three against CH</td></tr>
    <tr><td>0.50</td><td>1</td><td>even</td></tr>
    <tr class="file-row"><td>0.61</td><td>1.56</td><td>CH share in this file</td></tr>
    <tr><td>0.75</td><td>3</td><td>three to one for CH</td></tr>
  </tbody>
</table>
```

```{r}
#| echo: true
p <- round(mean(oj$y), 2)
round(p / (1 - p), 2)
```

::: {.fragment}
<p class="q-prompt">The code prints 1.56. Which row is this file, and which rows are only the language of odds?</p>
:::

::: {.notes}
The highlighted row is the measured pair: 0.61 ↔ 1.56. The other rows teach the ladder. Do not invent a second share.

If the odds are 1, $p = 0.5$. Say that if someone asks. Do not open a classification cutoff.
:::

## A coefficient multiplies the odds, not $p$

- Write the score as a line: $\eta = \beta_0 + \beta_1 \mathrm{LoyalCH}$.
- A step of size $\Delta$ in loyalty **adds** $\beta_1 \Delta$ to the log-odds.
- That **multiplies** the odds of CH by $e^{\beta_1 \Delta}$.
- It does **not** add $\beta_1 \Delta$ to the probability.

::: {.formula}
$$
\frac{\text{odds after}}{\text{odds before}} = e^{\beta_1 \Delta}.
$$
:::

<p class="caption">The size of the change in $p$ still depends on where you start. The bins already showed unequal steps.</p>

::: {.notes}
The rule, before the estimated number. Do not quote $e^{6.09}$ as a “600 times” story — LoyalCH only moves inside 0 to 1, and the worked step is +0.1 on the example after the fit.

Association, not a cause. ISLR reading-a-coefficient beat, held until we have $\hat\beta$ on OJ.
:::

## What to remember — and what we do next

:::: {.two-box}
::: {.fragment}
::: {.box-card}
<p class="card-h">Remember</p>
<ul>
<li>On this file a straight line left (0, 1): about 0.036 to 1.051.</li>
<li>The logistic curve stays inside (0, 1).</li>
<li>A coefficient multiplies the <strong>odds</strong>. It does not add a constant to $p$.</li>
</ul>
:::
:::
::: {.fragment}
::: {.box-card .alt}
<p class="card-h">Next</p>
<ul>
<li>Fit that curve on OJ. Loyalty first.</li>
<li>Then the price gap, with loyalty held fixed.</li>
<li>Score <strong>one</strong> new visit as a probability. Stop there.</li>
</ul>
:::
:::
::::

::: {.notes}
Two clicks, one card each. Do not add a third card.

Then maximum likelihood, lightly, and the measured coefficients −2.76 and 6.09.
:::

## R picks the coefficients that fit these purchases {.section-start}

<p class="act-kicker">One idea</p>

<p class="setup-line">Maximum likelihood chooses $\beta$ so the CH and MM purchases in this file are as likely as possible under the curve. `glm` does the search. We read the result.</p>

::: {.notes}
Section start. Light. No derivation, no likelihood product on the glass.

ISLR estimation, one sentence. Then the measured table. OJ only.
:::

## Example on OJ — loyalty alone

:::: {.columns}
::: {.column width="46%"}
`glm(y ~ LoyalCH)` on this file:

- Intercept ≈ **`r r2(coef(fit1)[1])`**
- LoyalCH ≈ **`r r2(coef(fit1)[2])`**

::: {.formula}
$$
\hat\eta = -2.76 + 6.09\,\mathrm{LoyalCH}.
$$
:::

```{r}
#| echo: true
fit1 <- glm(y ~ LoyalCH, oj, family = binomial)
round(coef(fit1), 2)
```
:::
::: {.column width="54%"}
::: {.fragment}
```{r}
#| fig-width: 6.2
#| fig-height: 3.7
#| fig-alt: "Fitted probability of CH against LoyalCH, from the one-predictor logistic regression."
ggplot(grid_loyal, aes(LoyalCH, p_hat)) +
  geom_line(color = weaver, linewidth = 1.25) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed", color = iron, linewidth = 0.3) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
  labs(x = "LoyalCH", y = "Fitted P(CH)") +
  theme_mlba(13)
```
:::
:::
::::

::: {.fragment}
<p class="q-prompt">The slope is about 6.09. Is that “+6.09 probability” when loyalty goes from 0 to 1?</p>
:::

::: {.notes}
Claim first: −2.76 and 6.09. Then the curve.

Answer: no. 6.09 is on the log-odds scale. The next slide measures a step of +0.1: the odds multiply by about 1.84, and $\Delta p$ depends on the starting loyalty.
:::

## Example on OJ — same +0.1, two changes in $p$

:::: {.columns}
::: {.column width="42%"}
- From the fit you just saw, a **+0.1** step in `LoyalCH` multiplies the odds of CH by about **`r r2(odds_loyal_step)`**.
- That multiplier does not depend on where you start.
- The change in the **probability** does.

- Start at 0.32 (MM buyers’ mean): about **`r r2(p_lo)`** → **`r r2(p_lo_step)`**.
- Start at 0.72 (CH buyers’ mean): about **`r r2(p_hi)`** → **`r r2(p_hi_step)`**.
:::
::: {.column width="58%"}
::: {.r-stack}
::: {.fragment .fade-in-then-out}
```{r}
#| fig-width: 6.4
#| fig-height: 3.55
#| fig-alt: "Logistic curve with a +0.1 loyalty step starting at 0.32. The chance moves from about 0.31 to about 0.45."
seg_lo <- data.frame(x = 0.32, xend = 0.42, y = p_lo, yend = p_lo_step)
pts_lo <- data.frame(LoyalCH = c(0.32, 0.42), p_hat = c(p_lo, p_lo_step))
ggplot(glass_curve, aes(LoyalCH, p_hat)) +
  geom_line(color = weaver, linewidth = 1.15) +
  geom_segment(
    data = seg_lo,
    aes(x = x, xend = xend, y = y, yend = yend),
    inherit.aes = FALSE, color = red, linewidth = 1.15
  ) +
  geom_point(data = pts_lo, color = red, size = 3.1) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed", color = iron, linewidth = 0.3) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
  labs(x = "LoyalCH", y = "Chance of CH", caption = "Start at 0.32") +
  theme_mlba(13)
```
:::
::: {.fragment .fade-in}
```{r}
#| fig-width: 6.4
#| fig-height: 3.55
#| fig-alt: "The same curve with a +0.1 loyalty step starting at 0.72. The chance moves from about 0.84 to about 0.90."
seg_hi <- data.frame(x = 0.72, xend = 0.82, y = p_hi, yend = p_hi_step)
pts_hi <- data.frame(LoyalCH = c(0.72, 0.82), p_hat = c(p_hi, p_hi_step))
ggplot(glass_curve, aes(LoyalCH, p_hat)) +
  geom_line(color = weaver, linewidth = 1.15) +
  geom_segment(
    data = seg_hi,
    aes(x = x, xend = xend, y = y, yend = yend),
    inherit.aes = FALSE, color = red, linewidth = 1.15
  ) +
  geom_point(data = pts_hi, color = red, size = 3.1) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed", color = iron, linewidth = 0.3) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
  labs(x = "LoyalCH", y = "Chance of CH", caption = "Start at 0.72") +
  theme_mlba(13)
```
:::
:::
:::
::::

::: {.notes}
Odds multiplier first, in the claim: about 1.84, from $e^{0.1 \times 6.09}$ using the rounded slope. It is the same at both starts.

Then the pictures, one click each. At 0.32 the chance goes about 0.31 → 0.45. At 0.72 it goes about 0.84 → 0.90. Same step in loyalty, different $\Delta p$. Those four probabilities use the rounded coefficients −2.76 and 6.09.

Do not call 1.84 a change in the probability.
:::

## Add the price gap, loyalty held fixed {.section-start}

<p class="act-kicker">Same shopper, different shelf</p>

- PriceDiff = sale price MM − sale price CH.
- The model is now $\hat\eta = \hat\beta_0 + \hat\beta_1 \mathrm{LoyalCH} + \hat\beta_2 \mathrm{PriceDiff}$.
- **Held fixed** means two visits with the same `LoyalCH` and a different price gap.
- On this file: intercept ≈ **`r r2(coef(fit2)[1])`**, LoyalCH ≈ **`r r2(coef(fit2)[2])`**, PriceDiff ≈ **`r r2(coef(fit2)[3])`**.

```{r}
#| echo: true
fit2 <- glm(y ~ LoyalCH + PriceDiff, oj, family = binomial)
round(coef(fit2), 2)
```

<p class="caption">The PriceDiff coefficient is that comparison, on the log-odds scale. It is an association for prediction. It is not a cause.</p>

::: {.notes}
Section start. Read “held fixed” as the partial association: same loyalty, different price gap. Do not teach a confounding story, a causal diagram, or “we removed bias.” Prediction only.

Measured coefficients: −3.25, 6.40, 2.86. The +$0.10 step is the next slide.
:::

## Example on OJ — ten cents on the price gap

:::: {.columns}
::: {.column width="44%"}
- Hold `LoyalCH` fixed. Raise PriceDiff by **$0.10** (MM becomes 10 cents more expensive relative to CH).
- The odds of CH are multiplied by about **`r r2(odds_price_step)`**.
- At LoyalCH = 0.5, that step moves the chance from about **`r r2(p_pd0)`** to about **`r r2(p_pd1)`**.

::: {.formula}
$$
e^{0.10 \times 2.86} \approx 1.33.
$$
:::
:::
::: {.column width="56%"}
::: {.fragment}
```{r}
#| fig-width: 6.3
#| fig-height: 3.7
#| fig-alt: "Fitted P(CH) against PriceDiff with LoyalCH held at 0.5. A ten-cent step is marked."
seg_pd <- data.frame(x = 0, xend = 0.1, y = p_pd0, yend = p_pd1)
pts_pd <- data.frame(PriceDiff = c(0, 0.1), p_hat = c(p_pd0, p_pd1))
ggplot(price_grid, aes(PriceDiff, p_hat)) +
  geom_line(color = weaver, linewidth = 1.2) +
  geom_segment(
    data = seg_pd,
    aes(x = x, xend = xend, y = y, yend = yend),
    inherit.aes = FALSE, color = red, linewidth = 1.15
  ) +
  geom_point(data = pts_pd, color = red, size = 3.1) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
  labs(
    x = "PriceDiff (dollars)",
    y = "Fitted P(CH)",
    caption = "LoyalCH held at 0.5"
  ) +
  theme_mlba(13)
```
:::
:::
::::

::: {.fragment}
<p class="q-prompt">Holding loyalty fixed, does a larger MM–CH price gap push the chance of CH up or down?</p>
:::

::: {.notes}
Claim first: odds × about 1.33, from $e^{0.10 \times 2.86}$ on the rounded PriceDiff coefficient. Then the curve at LoyalCH = 0.5.

Up. At loyalty 0.5 the chance goes about 0.49 → 0.56. That pair is `predict` on this fit, not a new model. Do not quote a mean PriceDiff.
:::

## Score one visit by hand {.section-start}

<p class="act-kicker">LoyalCH = 0.5, PriceDiff = +0.2</p>

Use the rounded coefficients already on the board.

::: {.formula}
$$
\hat\eta = -3.25 + 6.40(0.5) + 2.86(0.2) = `r r3(eta_hand)`.
$$
:::

::: {.fragment}
::: {.formula}
$$
\hat p = \frac{e^{0.522}}{1 + e^{0.522}} \approx `r r2(p_hand)`.
$$
:::
:::

::: {.fragment}
```{r}
#| echo: true
new <- data.frame(LoyalCH = 0.5, PriceDiff = 0.2)
round(predict(fit2, new, type = "response"), 2)
```
:::

<p class="caption">The hand score uses the rounded coefficients. `predict` on the same row agrees at two decimals: about 0.63.</p>

::: {.notes}
Section start. Three beats: write $\hat\eta = 0.522$, turn it into $\hat p \approx 0.63$, then show `predict` prints 0.63.

Do not also print the unrounded linear predictor. The 0.522 is the arithmetic from −3.25, 6.40, and 2.86. Association for this visit, not a shelf rule.
:::

## Example on OJ — same loyalty, opposite price

:::: {.columns}
::: {.column width="42%"}
- LoyalCH 0.5, PriceDiff **+0.2** → $\hat p$ ≈ **`r r2(new3$p_hat[1])`**
- LoyalCH 0.5, PriceDiff **−0.2** → $\hat p$ ≈ **`r r2(new3$p_hat[2])`**
- LoyalCH 0.9, PriceDiff **0** → $\hat p$ ≈ **`r r2(new3$p_hat[3])`**

```{r}
#| echo: true
visits <- data.frame(
  LoyalCH = c(0.5, 0.5, 0.9),
  PriceDiff = c(0.2, -0.2, 0)
)
round(predict(fit2, visits, type = "response"), 2)
```
:::
::: {.column width="58%"}
::: {.r-stack}
::: {.fragment .fade-in-then-out}
```{r}
#| fig-width: 6.4
#| fig-height: 3.6
#| fig-alt: "Two bars: predicted P(CH) about 0.63 and 0.35 at the same loyalty and opposite price gaps."
ggplot(new3[1:2, ], aes(who, p_hat, fill = who)) +
  geom_col(width = 0.62) +
  geom_text(aes(label = sprintf("%.2f", p_hat)), vjust = -0.35,
            fontface = "bold", color = weaver, size = 5) +
  scale_fill_manual(values = c(weaver, iron, teal)) +
  scale_y_continuous(limits = c(0, 1.12), breaks = c(0, 0.5, 1), expand = c(0, 0)) +
  labs(x = NULL, y = "Predicted P(CH)") +
  theme_mlba(12)
```
:::
::: {.fragment .fade-in}
```{r}
#| fig-width: 6.4
#| fig-height: 3.6
#| fig-alt: "Three bars: predicted P(CH) about 0.63, 0.35, and 0.92."
ggplot(new3, aes(who, p_hat, fill = who)) +
  geom_col(width = 0.68) +
  geom_text(aes(label = sprintf("%.2f", p_hat)), vjust = -0.35,
            fontface = "bold", color = weaver, size = 4.6) +
  scale_fill_manual(values = c(weaver, iron, teal)) +
  scale_y_continuous(limits = c(0, 1.12), breaks = c(0, 0.5, 1), expand = c(0, 0)) +
  labs(x = NULL, y = "Predicted P(CH)") +
  theme_mlba(12)
```
:::
:::
:::
::::

::: {.fragment}
<p class="q-prompt">In the first two rows, which lever moved — loyalty, or the price gap?</p>
:::

::: {.notes}
Claim lists the three measured chances: 0.63, 0.35, 0.92. First picture: the opposite price gaps at loyalty 0.5. Second picture: add the high-loyalty row, about 0.92 when LoyalCH is 0.9 and the gap is 0.

Answer: the first two rows hold loyalty at 0.5. The price gap moves the chance from about 0.63 to about 0.35. The third row is loyalty.

Do not turn 0.63 into a yes. No cutoff.
:::

## What to remember

- CH `r n_ch` / MM `r n_mm`. Mean loyalty **`r r2(loyal_means["CH"])`** vs **`r r2(loyal_means["MM"])`**.
- A line on 0/1 ran about **`r r3(lin_min)`** to **`r r3(lin_max)`**. Bin CH rates: **0.12, 0.28, 0.53, 0.77, 0.96**.
- Loyalty alone: **−2.76 + 6.09 LoyalCH**. A +0.1 step multiplies the odds by about **`r r2(odds_loyal_step)`**.
- With the price gap: **−3.25 + 6.40 LoyalCH + 2.86 PriceDiff**. Ten cents multiplies the odds by about **`r r2(odds_price_step)`**.
- LoyalCH 0.5 and PriceDiff +0.2 → about **0.63**. The opposite gap → about **0.35**. LoyalCH 0.9 and a flat gap → about **0.92**.

::: {.fragment}
<p class="q-prompt">Can we predict which OJ brand a shopper buys?</p>
:::

::: {.fragment}
<p class="today-note">A probability for <strong>this</strong> visit, from loyalty and the price gap, between 0 and 1. An association. Not a cause.</p>
:::

::: {.notes}
The numbers are the ones already measured. Then return the question from the bar chart.

Answer in one sentence, then stop. Do not add a cutoff, a hit rate, or another dataset.
:::

## Before next class {.next-slide}

<ol class="today-list">
<li>Midterm <strong>review</strong> is the next methods meeting. The <strong>midterm</strong> is the meeting after that. Closed notes. Weeks 1–5, including logistic.</li>
<li>Finish Homework 2 if it is still open, on the schedule.</li>
<li>Reading already named: ISLR Chapter 4, the logistic sections.</li>
<li>The logistic <strong>R lab</strong> is after fall break. Bring a laptop that day, once it is posted.</li>
</ol>

::: {.notes}
Your eyes only. Slot for this deck: Wed 30 Sep 2026 discussion, continuing the Monday pair. Midterm review Mon 5 Oct. Midterm Wed 7 Oct. Fall break 11–15 Oct. Do not invent a logistic lab date. The lab is after fall break.

Homework 2: the outline lock was Sun 4 Oct, 11:59 pm. The materials-site schedule lists Sun 28 Sep. Check Canvas before you say a due date aloud. Do not put either date on a slide.

One deferral, this note only — do not put it on the glass: hits, misses, and decision cutoffs return with the classification-metrics week after the break. If someone asks, one sentence, then stop. No confusion matrix.
:::
