---
title: "Logistic regression"
author: "Andrés Castaño Zuluaga"
format:
  revealjs:
    theme: simple
    css:
      - ../_theme/mlba-reveal.css
      - session05.css
    slide-number: true
    embed-resources: true
    html-math-method: katex
    width: 1280
    height: 720
    margin: 0.04
    center: false
    auto-stretch: false
    min-scale: 0.2
    max-scale: 1.4
    transition: none
    navigation-mode: linear
    controls-tutorial: false
    hash-type: number
    highlight-style: arrow
    code-copy: false
    code-overflow: wrap
    title-slide-attributes:
      data-background-color: "#182C4B"
      class: "quarto-title-block"
    footer: "70-467 · Logistic regression"
    template-partials:
      - title-slide.html
execute:
  echo: false
  warning: false
  message: false
fig-dpi: 140
---

```{r}
#| label: setup
#| include: false
library(ggplot2)

# ---- File -------------------------------------------------------------------
oj_path_candidates <- c(
  "OJ.csv",
  "../../_project/data_pack/teaching/w05_slides/OJ.csv",
  "../../../_project/data_pack/teaching/w05_slides/OJ.csv",
  "../../_project/data_pack/teaching/backups/OJ.csv",
  "../../../_project/data_pack/teaching/backups/OJ.csv"
)
oj_path <- oj_path_candidates[file.exists(oj_path_candidates)][1]
if (is.na(oj_path)) stop("OJ.csv not found next to the deck or in the data pack")

oj <- read.csv(oj_path, stringsAsFactors = FALSE)
# Some exports keep an unnamed row-id in the first column. Drop it.
if (!names(oj)[1] %in% c("Purchase", "LoyalCH", "PriceDiff")) {
  oj <- oj[, -1, drop = FALSE]
}
oj$Purchase <- factor(oj$Purchase, levels = c("CH", "MM"))
oj$y <- as.integer(oj$Purchase == "CH")

n <- nrow(oj)
n_ch <- sum(oj$y)
n_mm <- n - n_ch
share_ch <- mean(oj$y)

# ---- Plot theme (same ink as Session 04) ------------------------------------
weaver <- "#182C4B"; red <- "#C41230"; iron <- "#6D6E71"; paper <- "#F4F6F8"; teal <- "#008F91"
theme_mlba <- function(base = 15) {
  theme_minimal(base_family = "sans", base_size = base) +
    theme(
      plot.background = element_rect(fill = paper, color = NA),
      panel.background = element_rect(fill = paper, color = NA),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "#E0E0E0", linewidth = 0.3),
      axis.title = element_text(color = iron, size = base - 2),
      axis.text = element_text(color = weaver, size = base - 3),
      legend.position = "none",
      plot.margin = margin(6, 10, 4, 4),
      plot.caption = element_text(color = iron, size = base - 5, hjust = 0),
      strip.text = element_text(color = weaver, face = "bold", size = base - 2),
      strip.background = element_rect(fill = "#E0E0E0", color = NA)
    )
}
brand_fill <- c(CH = "#D6DEE8", MM = "#E4E5E7")

# ---- Locked teaching fits: CH = 1, MM = 0 -----------------------------------
fit_lin <- lm(y ~ LoyalCH, data = oj)
lin_hat <- predict(fit_lin)
lin_min <- min(lin_hat)
lin_max <- max(lin_hat)

fit1 <- glm(y ~ LoyalCH, data = oj, family = binomial)
fit2 <- glm(y ~ LoyalCH + PriceDiff, data = oj, family = binomial)

oj$loyal_bin <- cut(
  oj$LoyalCH,
  breaks = c(0, 0.2, 0.4, 0.6, 0.8, 1),
  include.lowest = TRUE,
  right = FALSE
)
bin_df <- data.frame(
  bin = levels(oj$loyal_bin),
  n = as.integer(tapply(oj$y, oj$loyal_bin, length)),
  rate = as.numeric(tapply(oj$y, oj$loyal_bin, mean)),
  stringsAsFactors = FALSE
)
bin_df$label <- sprintf("%.2f", bin_df$rate)

loyal_means <- tapply(oj$LoyalCH, oj$Purchase, mean)
grid_loyal <- data.frame(LoyalCH = seq(0, 1, length.out = 240))
grid_loyal$p_hat <- as.numeric(predict(fit1, newdata = grid_loyal, type = "response"))
grid_loyal$lin <- as.numeric(predict(fit_lin, newdata = grid_loyal))

new3 <- data.frame(
  LoyalCH = c(0.5, 0.5, 0.9),
  PriceDiff = c(0.2, -0.2, 0)
)
new3$p_hat <- as.numeric(predict(fit2, newdata = new3, type = "response"))
new3$who <- factor(
  c("LoyalCH 0.5\nPriceDiff +0.2", "LoyalCH 0.5\nPriceDiff −0.2", "LoyalCH 0.9\nPriceDiff 0"),
  levels = c("LoyalCH 0.5\nPriceDiff +0.2", "LoyalCH 0.5\nPriceDiff −0.2", "LoyalCH 0.9\nPriceDiff 0")
)
new1 <- new3[1, ]

# Odds multipliers from the rounded glass estimates, so the table matches the coefs.
b2_glass <- round(coef(fit2), 2)
odds_mult <- exp(b2_glass)

r2 <- function(x) sprintf("%.2f", x)
r3 <- function(x) sprintf("%.3f", x)

# Hard lock. If a future OJ export drifts, stop rather than teach a new number.
stopifnot(n == 1070L, n_ch == 653L, n_mm == 417L)
stopifnot(r2(share_ch) == "0.61")
stopifnot(r2(loyal_means["CH"]) == "0.72", r2(loyal_means["MM"]) == "0.32")
stopifnot(r3(lin_min) == "0.036", r3(lin_max) == "1.051")
stopifnot(all(bin_df$label == c("0.12", "0.28", "0.53", "0.77", "0.96")))
stopifnot(r2(coef(fit1)[1]) == "-2.76", r2(coef(fit1)[2]) == "6.09")
stopifnot(r2(coef(fit2)[1]) == "-3.25", r2(coef(fit2)[2]) == "6.40", r2(coef(fit2)[3]) == "2.86")
stopifnot(all(r2(new3$p_hat) == c("0.63", "0.35", "0.92")))
```

## Where Cas and Abdullah left us

```{=html}
<div class="pair-card">
  <p><strong>Cas</strong> and <strong>Abdullah</strong> opened logistic regression on Monday.</p>
  <p>Today we continue — same topic, one store file, one decision.</p>
  <p>We will not re-teach their intro from a cold start.</p>
</div>
```

::: {.notes}
Bridge only. Thank them by name. One sentence on what they opened, then move to the grid and the OJ file.

Do not steal their examples onto later slides. Do not put the Monday date on the title. Book pages stay in these notes: ISLR 2nd ed. §4.1–4.3 (~pp. 129–141) is the spine for the hour. Logistics (this discussion; R lab after fall break; midterm week) stay in your mouth or the final next-slide — not here.
:::

## Where are we on the grid?

<p class="q-prompt">Last week we shrank slopes for a dollar balance. What changes when the thing we want to predict is a <strong>yes/no</strong> at the shelf?</p>

```{=html}
<div class="grid-wrap">
<table class="mlba-grid">
  <colgroup>
    <col class="topic" /><col class="desc" /><col class="tools" />
  </colgroup>
  <thead>
    <tr><th>Topic</th><th>Description</th><th>Tools</th></tr>
  </thead>
  <tbody>
    <tr class="band"><td colspan="3">Fit &amp; evaluate</td></tr>
    <tr class="done"><td>Linear regression</td><td>Predict a number</td><td>OLS</td></tr>
    <tr class="done"><td>Metrics</td><td>Size of the errors</td><td>RMSE, R²</td></tr>
    <tr class="done"><td>Train/test, leakage, cross-validation</td><td>Error on new data</td><td>Holdout, k-fold</td></tr>
    <tr class="done"><td>Selection + regularization</td><td>Select variables or shrink coefficients</td><td>Best subset, ridge, lasso</td></tr>
    <tr class="band"><td colspan="3">Classify &amp; decide</td></tr>
    <tr class="here"><td>Logistic</td><td>Predict a yes/no outcome</td><td>Logistic regression</td></tr>
    <tr><td>Thresholds</td><td>Turn a probability into a decision</td><td>ROC, cutoff</td></tr>
    <tr class="band"><td colspan="3">Flexible &amp; unsupervised</td></tr>
    <tr><td>Trees</td><td>Split the predictors in sequence</td><td>Decision trees</td></tr>
    <tr><td>Forests / boosting</td><td>Combine many trees</td><td>Random forest, boosting</td></tr>
    <tr><td>PCA</td><td>Fewer combinations of the predictors</td><td>Principal components</td></tr>
    <tr><td>Clustering</td><td>Group similar observations</td><td>K-means</td></tr>
  </tbody>
</table>
</div>
```

<p class="caption">The marked cell is this hour: a probability of a brand choice. Thresholds stay later. Do not reopen ridge or lasso.</p>

::: {.notes}
Open block — no click. The Logistic row is already marked. You name “probability of a brand choice.”

Thresholds / ROC stay with the classification-metrics week after fall break. Do not open a confusion matrix. ISLR Ch. 4 opens classification here; we stop at $\hat p$.
:::

## Can we predict which OJ brand a shopper buys?

:::: {.columns}
::: {.column .spaced-list width="52%"}
- One row = one store purchase of orange juice ($n$ = `r format(n, big.mark = ",")`).
- $Y$: `Purchase` — **CH** (Citrus Hill) or **MM** (Minute Maid).
- Measured mix: CH `r n_ch` (`r sprintf("%.1f%%", 100 * n_ch / n)`) / MM `r n_mm` (`r sprintf("%.1f%%", 100 * n_mm / n)`). No missing values.
- Decision this hour: a **new** visit is about to happen. What is the chance they pick Citrus Hill?
:::
::: {.column width="48%"}
```{r}
#| fig-width: 5.6
#| fig-height: 3.55
#| fig-alt: "Bar chart of 653 Citrus Hill purchases and 417 Minute Maid purchases."
brand_n <- data.frame(
  Purchase = factor(c("CH", "MM"), levels = c("CH", "MM")),
  n = c(n_ch, n_mm)
)
ggplot(brand_n, aes(Purchase, n, fill = Purchase)) +
  geom_col(width = 0.62) +
  geom_text(aes(label = format(n, big.mark = ",")), vjust = -0.35,
            fontface = "bold", color = weaver, size = 5) +
  scale_fill_manual(values = c(CH = weaver, MM = iron)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.14)), limits = c(0, 820)) +
  labs(x = NULL, y = "Purchases") +
  theme_mlba(15)
```
:::
::::

<p class="q-prompt">If headquarters only said “most people buy CH,” is that enough for <strong>this</strong> shopper, with a loyalty score and a price gap?</p>

::: {.notes}
Sit with the question. The answer comes across the logistic block. Do not say the file name on the title.

Majority share ≈ 0.61 is the dumb baseline — name it later if useful, not as a classification-metrics lesson. ISLR 4.1: qualitative response. Student story is OJ / brand choice only.
:::

## What is in the file?

```{=html}
<table class="cb">
  <thead>
    <tr><th>Variable</th><th>Type</th><th>Unit</th><th>Description</th></tr>
  </thead>
  <tbody>
    <tr><td>Purchase</td><td>CH / MM</td><td>brand</td><td>Which OJ they bought (Y)</td></tr>
    <tr><td>LoyalCH</td><td>numeric</td><td>0–1 score</td><td>Customer loyalty toward Citrus Hill</td></tr>
    <tr><td>PriceDiff</td><td>numeric</td><td>dollars</td><td>Sale price MM − sale price CH</td></tr>
    <tr><td>SalePriceCH</td><td>numeric</td><td>dollars</td><td>CH shelf / sale price</td></tr>
    <tr><td>SalePriceMM</td><td>numeric</td><td>dollars</td><td>MM shelf / sale price</td></tr>
  </tbody>
</table>
```

<p class="caption">Workhorses this hour: <strong>LoyalCH</strong> and <strong>PriceDiff</strong>. List prices, discounts, special-display flags, and store ids are in the file. We will not use them today.</p>

::: {.notes}
Point at LoyalCH and PriceDiff. The rest, if asked: PriceCH, PriceMM (list prices), DiscCH, DiscMM (discounts), SpecialCH, SpecialMM (0/1 display flags), StoreID, STORE, Store7, PctDiscCH, PctDiscMM, ListPriceDiff. WeekofPurchase stays off the main story unless asked.

Do not put every column on the glass.
:::

## Outline of today

<ol class="today-list">
<li>Why a straight line on 0/1 purchases fails (on this file)</li>
<li>Odds and the logit — the language logistic uses</li>
<li>Fit on OJ; read loyalty and price gap carefully</li>
<li>Score <strong>one</strong> new visit (a probability) — and stop there</li>
<li>What we can and cannot claim</li>
</ol>

::: {.notes}
Keep confusion matrices, cutoffs, and ROC out of the bullets. No book pages on the glass. Spine = ISLR through “making predictions,” not into LDA, QDA, or KNN. Midterm week is named only on the closing slide.
:::

## Why not a straight line on brand choice? {.section-start}

<p class="act-kicker">This file</p>

<p class="setup-line">Brand choice is a yes or a no. A straight line was built for a number that can sit anywhere — a dollar balance, not a brand.</p>

::: {.notes}
Section start: weaver band, not a blank dark slide. One sentence, then the binary outcome.

ISLR ~pp. 129–133: binary response, then why not least squares on 0/1. No other dataset.
:::

## Brand choice is not a dollar balance

:::: {.columns}
::: {.column width="48%"}
- Code for this hour: $Y_i = 1$ if `Purchase` = CH, $Y_i = 0$ if MM.
- Mean of $Y$ = share who bought CH ≈ **`r r2(share_ch)`** (`r n_ch` / `r format(n, big.mark = ",")`).
- Linear regression was built for a number that can sit anywhere on the line. Brand choice is not that.

::: {.formula}
$$
Y_i =
\begin{cases}
1 & \text{if Purchase}_i = \text{CH},\\
0 & \text{if Purchase}_i = \text{MM}.
\end{cases}
$$
:::

```{r}
#| echo: true
oj$y <- as.integer(oj$Purchase == "CH")
format(round(mean(oj$y), 3), nsmall = 3)
```
:::
::: {.column width="52%"}
```{r}
#| fig-width: 6.2
#| fig-height: 4.15
#| fig-alt: "Purchases jittered into two bands at 0 and 1 along LoyalCH."
ggplot(oj, aes(LoyalCH, y)) +
  geom_jitter(height = 0.045, width = 0, alpha = 0.28, color = weaver, size = 1.45) +
  scale_y_continuous(
    breaks = c(0, 1),
    labels = c("0   MM", "1   CH"),
    limits = c(-0.18, 1.18)
  ) +
  labs(x = "LoyalCH", y = NULL) +
  theme_mlba(14)
```
:::
::::

<p class="q-prompt">If we say “the average purchase is `r r2(share_ch)`,” what does that mean for the next shopper?</p>

::: {.notes}
You say: it is the share who bought CH — not yet a usable probability for this loyalty score.

ISLR 4.1 qualitative / binary response. Measured only. CH is hard-coded as 1.
:::

## Example on OJ — do loyal shoppers buy CH more often?

:::: {.columns}
::: {.column .spaced-list width="42%"}
- Mean `LoyalCH` ≈ **`r r2(loyal_means["CH"])`** among CH buyers.
- Mean `LoyalCH` ≈ **`r r2(loyal_means["MM"])`** among MM buyers.
- PriceDiff also runs higher among CH buyers (MM relatively more expensive).
- The predictors are not noise. We still need a probability model.

```{r}
#| echo: true
round(tapply(oj$LoyalCH, oj$Purchase, mean), 2)
```
:::
::: {.column width="58%"}
```{r}
#| fig-width: 6.4
#| fig-height: 4.05
#| fig-alt: "Boxplots of LoyalCH for CH and MM buyers, with means marked."
means_df <- data.frame(
  Purchase = factor(c("CH", "MM"), levels = c("CH", "MM")),
  LoyalCH = as.numeric(loyal_means[c("CH", "MM")])
)
ggplot(oj, aes(Purchase, LoyalCH)) +
  geom_boxplot(aes(fill = Purchase), width = 0.55, color = weaver,
               outlier.size = 0.8, outlier.alpha = 0.35) +
  geom_point(data = means_df, color = red, size = 3.2, shape = 18) +
  geom_text(
    data = means_df,
    aes(label = sprintf("mean %.2f", LoyalCH)),
    color = red, fontface = "bold", size = 4.2, nudge_x = 0.28, nudge_y = 0.04
  ) +
  scale_fill_manual(values = brand_fill) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(x = NULL, y = "LoyalCH") +
  theme_mlba(14)
```
:::
::::

<p class="q-prompt">Before any formula: which shopper would you bet on for CH — loyalty 0.3 or 0.7?</p>

::: {.notes}
Application slide, same job as Session 04’s “Example: Credit.” The two means are the click in your mouth: 0.72 versus 0.32. They are already on the figure so the room can see them without a second chart.

Association only. PriceDiff is higher among CH buyers — do not quote a new mean unless someone asks. No causal claim.
:::

## What if we treat CH/MM like a dollar amount?

:::: {.columns}
::: {.column width="44%"}
- Temptation: treat $Y$ like a dollar balance and run least squares on `LoyalCH`.
- The fitted value is then “expected brand” as a straight line in loyalty.
- That number is **not** forced to stay between 0 and 1.

::: {.formula}
$$
\hat Y_i = \hat\beta_0 + \hat\beta_1\,\mathrm{LoyalCH}_i
$$
:::

<p class="caption">On this file, $\hat Y_i$ is not guaranteed to be a probability.</p>

```{r}
#| echo: true
fit_lin <- lm(y ~ LoyalCH, data = oj)
range(round(predict(fit_lin), 3))
```
:::
::: {.column width="56%"}
```{r}
#| fig-width: 6.3
#| fig-height: 3.9
#| fig-alt: "Least-squares line through 0/1 purchases against LoyalCH."
ggplot(oj, aes(LoyalCH, y)) +
  geom_jitter(height = 0.04, width = 0, alpha = 0.25, color = weaver, size = 1.3) +
  geom_line(data = grid_loyal, aes(LoyalCH, lin), color = red, linewidth = 1.05) +
  scale_y_continuous(breaks = c(0, 0.5, 1), limits = c(-0.12, 1.18)) +
  labs(x = "LoyalCH", y = "Fitted line on Y") +
  theme_mlba(13)
```
:::
::::

<p class="q-prompt">Point at a very loyal shopper, or a very disloyal one. Can the line’s prediction leave (0, 1)?</p>

::: {.notes}
Hold the question. The next slide shows the measured range. ISLR why-not-linear beat. Glass stays OJ language. The `range()` line is the same fit as the picture.
:::

## Example on OJ — a “chance” of 1.05?

:::: {.columns}
::: {.column .spaced-list width="42%"}
- On **this** file, `lm(y ~ LoyalCH)` predictions run from about **`r r3(lin_min)`** to about **`r r3(lin_max)`**.
- Values under 0 or over 1 are not chances.
- A brand decision needs a number you can read as a **probability**.
:::
::: {.column width="58%"}
```{r}
#| fig-width: 6.5
#| fig-height: 4.15
#| fig-alt: "Least-squares line crossing the fences at 0 and 1, annotated with the minimum and maximum fitted values."
fence_lab <- data.frame(
  LoyalCH = c(0.18, 0.72),
  y = c(lin_min + 0.10, 1.08),
  lab = c(sprintf("min ≈ %.3f", lin_min), sprintf("max ≈ %.3f", lin_max))
)
ggplot(oj, aes(LoyalCH, y)) +
  geom_jitter(height = 0.035, width = 0, alpha = 0.22, color = weaver, size = 1.25) +
  geom_line(data = grid_loyal, aes(LoyalCH, lin), color = red, linewidth = 1.05) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed", color = iron, linewidth = 0.4) +
  geom_text(data = fence_lab, aes(label = lab), color = red, fontface = "bold", size = 4) +
  scale_y_continuous(breaks = c(0, 0.5, 1), limits = c(-0.08, 1.2)) +
  labs(x = "LoyalCH", y = "Least-squares fit") +
  theme_mlba(13)
```
:::
::::

<p class="q-prompt">Would you trust a forecast that says “`r r3(lin_max)` chance of buying CH”?</p>

::: {.notes}
The class says no. These are the measured bounds on this file (about 0.036 to 1.051). Hard why-not-linear beat from ISLR, spoken in OJ. The code already ran on the previous slide.
:::

## Example on OJ — CH share by loyalty bins

:::: {.columns}
::: {.column width="40%"}
Slice shoppers by `LoyalCH` and count the CH rate in each slice.

```{=html}
<table class="cb odds-mini">
  <thead><tr><th>LoyalCH</th><th>CH rate</th></tr></thead>
  <tbody>
```
```{r}
#| results: asis
cat(paste0(
  "<tr><td>", bin_df$bin, "</td><td>", bin_df$label, "</td></tr>",
  collapse = "\n"
))
```
```{=html}
  </tbody>
</table>
```

```{r}
#| echo: true
round(tapply(oj$y, oj$loyal_bin, mean), 2)
```
:::
::: {.column width="60%"}
```{r}
#| fig-width: 6.6
#| fig-height: 4.2
#| fig-alt: "Bar chart of the Citrus Hill purchase rate in five LoyalCH bins, rising from about 0.12 to about 0.96."
ggplot(bin_df, aes(bin, rate)) +
  geom_col(width = 0.72, fill = weaver) +
  geom_text(aes(label = label), vjust = -0.35, fontface = "bold", color = weaver, size = 4.4) +
  scale_y_continuous(limits = c(0, 1.12), breaks = seq(0, 1, 0.25), expand = c(0, 0)) +
  labs(x = "LoyalCH bin", y = "Share who bought CH") +
  theme_mlba(13)
```
:::
::::

<p class="q-prompt">Is the jump from 0.2 to 0.4 the same size as the jump from 0.8 to 1.0?</p>

::: {.notes}
You say: a straight line cannot tell this story cleanly. The climb is steep in the middle and flatter near the ends — an S, before we fit one.

Exact bin rates, locked from this file: 0.12, 0.28, 0.53, 0.77, 0.96. This is the OJ stand-in for ISLR’s “probability is not linear in X” picture.
:::

## Stay between 0 and 1 — and still use loyalty

:::: {.columns}
::: {.column .spaced-list width="44%"}
- Goal: for each shopper, $p = \mathrm{P}(\mathrm{Purchase}=\mathrm{CH} \mid \text{their features})$.
- $p$ must live in (0, 1).
- Logistic regression builds that $p$ from a linear score, through a special curve.
:::
::: {.column width="56%"}
```{r}
#| fig-width: 6.2
#| fig-height: 3.7
#| fig-alt: "A smooth S-shaped curve from near 0 to near 1 as loyalty rises. Labeled as a teaching sketch, not the fit."
sketch <- data.frame(LoyalCH = seq(0, 1, length.out = 200))
sketch$p <- plogis(-2.2 + 4.8 * sketch$LoyalCH)
ggplot(sketch, aes(LoyalCH, p)) +
  geom_line(color = weaver, linewidth = 1.25) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed", color = iron, linewidth = 0.35) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
  labs(
    x = "LoyalCH",
    y = "p",
    caption = "Teaching sketch. Not the fit on this file."
  ) +
  theme_mlba(14)
```
:::
::::

<p class="q-prompt">If loyalty rises, should $p$ rise smoothly, without jumping the fence?</p>

::: {.notes}
Click the shape with your hand: that is where we are going. The sketch uses no coefficient from this file — do not read a number off it.

Bridge into odds and the logit. If Cas or Abdullah already showed an S-curve, nod and do not re-lecture their talk.
:::

## From chance to odds to a curve {.section-start}

<p class="act-kicker">Language</p>

<p class="setup-line">A probability stays between 0 and 1. Odds, and then the log of the odds, are how logistic regression puts a straight line under that curve.</p>

::: {.notes}
Section start, band only. ISLR ~pp. 133–137: logistic function, odds, logit. Glass = OJ moves.
:::

## Odds are “how many to one,” not the chance itself

:::: {.columns}
::: {.column width="48%"}
- If $p = \mathrm{P}(\mathrm{CH})$, the odds of CH are $p / (1-p)$.
- If $p = 0.75$, the odds are 3. Three to one for CH.
- If $p = `r r2(share_ch)`$ (the CH share in this file), the odds are about **`r sprintf("%.2f", round(share_ch, 2) / (1 - round(share_ch, 2)))`**.

::: {.formula}
$$
\mathrm{odds}(p) = \frac{p}{1-p}.
$$
:::

```{r}
#| echo: true
p <- round(mean(oj$y), 2)
round(p / (1 - p), 2)
```
:::
::: {.column width="52%"}
```{=html}
<table class="cb">
  <thead>
    <tr><th>p = P(CH)</th><th>Odds of CH</th><th></th></tr>
  </thead>
  <tbody>
    <tr><td>0.25</td><td>1/3</td><td>one to three against</td></tr>
    <tr><td>0.50</td><td>1</td><td>even</td></tr>
    <tr><td>0.75</td><td>3</td><td>three to one for CH</td></tr>
    <tr><td>0.61</td><td>1.56</td><td>CH share in this file</td></tr>
  </tbody>
</table>
```
:::
::::

<p class="q-prompt">If the odds of CH are 1, what is $p$?</p>

::: {.fragment}
<p class="today-note">$p = 0.5$. Even chance of Citrus Hill and Minute Maid.</p>
:::

::: {.notes}
Keep the examples in CH language. No horse-racing digression. ISLR odds beat.

The 0.25 / 0.50 / 0.75 rows are the language. The 0.61 row is this file. Click reveals $p = 0.5$ when the odds are 1.
:::

## Why take the log of the odds?

- $\mathrm{logit}(p) = \log\bigl(p/(1-p)\bigr)$.
- The logit lives on the whole real line. The probability stays in (0, 1).
- Logistic regression puts a **linear** model on the logit — not on $p$ itself.

::: {.formula}
$$
\mathrm{logit}(p) = \log\left(\frac{p}{1-p}\right),
\qquad
p = \mathrm{P}(\mathrm{Purchase}=\mathrm{CH}\mid \text{predictors}).
$$
:::

<p class="q-prompt">If the logit rises by a fixed amount when loyalty rises, does $p$ rise by that same fixed amount?</p>

::: {.fragment}
<p class="today-note">No. The change in $p$ depends on where you start on the S-curve. The bins already showed that.</p>
:::

::: {.notes}
Spoken title is the move (“why take the log”), not a textbook label alone. $p$ on this slide is P(Purchase = CH | predictors) in this file.

You say: the probability change is not a fixed step. ISLR logit beat, ~pp. 133–137.
:::

## A linear model for the logit — then back to a chance {.model-slide}

- Score loyalty and the price gap on the log-odds scale. Then turn the score into a probability.
- The map from score to probability is the logistic curve. It stays in (0, 1).

::: {.formula}
$$
\log\left(\frac{p_i}{1-p_i}\right)
=
\beta_0 + \beta_1\,\mathrm{LoyalCH}_i + \beta_2\,\mathrm{PriceDiff}_i
$$
$$
p_i
=
\frac{e^{\eta_i}}{1+e^{\eta_i}},
\qquad
\eta_i
=
\beta_0 + \beta_1\,\mathrm{LoyalCH}_i + \beta_2\,\mathrm{PriceDiff}_i
$$
:::

<p class="caption">$p_i = \mathrm{P}(\mathrm{Purchase}_i = \mathrm{CH} \mid \mathrm{LoyalCH}_i,\ \mathrm{PriceDiff}_i)$.</p>

<p class="q-prompt">If $\beta_1 > 0$, what happens to the <strong>odds</strong> of CH when loyalty rises, holding the price gap fixed?</p>

::: {.fragment}
<p class="today-note">The odds are multiplied by $e^{\Delta \cdot \beta_1}$. That is an association in the fit, not a lever that causes the purchase.</p>
:::

::: {.notes}
ISLR logistic model and the sigmoid. Prediction-first. The one-predictor fit is the warm-up; this is the workhorse equation for the hour. The fit itself is the next act — do not read numbers yet.

Click: odds multiply. Do not say “raise loyalty and you cause CH.”
:::

## Example on OJ — probability vs loyalty from this fit

:::: {.columns}
::: {.column width="44%"}
- `glm(y ~ LoyalCH)` on this file: intercept ≈ **`r r2(coef(fit1)[1])`**, LoyalCH ≈ **`r r2(coef(fit1)[2])`**.
- As `LoyalCH` rises, $\hat p = \mathrm{P}(\mathrm{CH})$ rises smoothly between 0 and 1.
- It matches the bin climb you already saw — now as one curve.

::: {.formula}
$$
\hat p = \frac{e^{\hat\eta}}{1+e^{\hat\eta}},
\quad
\hat\eta = `r r2(coef(fit1)[1])` + `r r2(coef(fit1)[2])`\,\mathrm{LoyalCH}
$$
:::

```{r}
#| echo: true
fit1 <- glm(y ~ LoyalCH, oj, family = binomial)
round(coef(fit1), 2)
```
:::
::: {.column width="56%"}
```{r}
#| fig-width: 6.3
#| fig-height: 4.05
#| fig-alt: "Fitted logistic curve of the probability of CH against LoyalCH, with the 0/1 purchases jittered behind it."
marks <- data.frame(LoyalCH = c(0.3, 0.7))
marks$p_hat <- as.numeric(predict(fit1, newdata = marks, type = "response"))
ggplot(oj, aes(LoyalCH, y)) +
  geom_jitter(height = 0.035, width = 0, alpha = 0.18, color = weaver, size = 1.15) +
  geom_line(data = grid_loyal, aes(LoyalCH, p_hat), color = red, linewidth = 1.15) +
  geom_point(data = marks, aes(LoyalCH, p_hat), color = red, size = 3.2) +
  scale_y_continuous(breaks = c(0, 0.5, 1), limits = c(-0.08, 1.08)) +
  labs(x = "LoyalCH", y = "Probability of CH") +
  theme_mlba(13)
```
:::
::::

<p class="q-prompt">Loyalty near 0.3, or near 0.7. Which shopper gets the higher chance of CH?</p>

::: {.notes}
The two dots on the curve are the click: about 0.3 versus about 0.7. The higher loyalty sits higher on the curve. Do not add a new printed probability beyond what the figure shows.

Exact coefficients, locked: intercept ≈ −2.76, LoyalCH ≈ 6.09. Application slide after the model idea.
:::

## Do not read β like a Balance slope

:::: {.columns}
::: {.column .spaced-list width="46%"}
- In linear regression, a slope was “dollars of balance per unit of $X$.”
- Here the LoyalCH coefficient is a change in the **log-odds** of CH — not “+6 points of probability.”
- How much $p$ moves depends on where you are on the S-curve. The bins already showed that.
:::
::: {.column width="54%"}
```{r}
#| fig-width: 6.1
#| fig-height: 3.85
#| fig-alt: "Least-squares line and logistic S-curve on the same LoyalCH axis, with fences at 0 and 1."
jump <- data.frame(
  x = c(0.2, 0.8),
  xend = c(0.3, 0.9),
  y = plogis(coef(fit1)[1] + coef(fit1)[2] * c(0.2, 0.8)),
  yend = plogis(coef(fit1)[1] + coef(fit1)[2] * c(0.3, 0.9)),
  ink = c(red, teal)
)
ggplot(grid_loyal, aes(LoyalCH)) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed", color = iron, linewidth = 0.35) +
  geom_line(aes(y = lin, color = "Least squares"), linewidth = 0.9) +
  geom_line(aes(y = p_hat, color = "Logistic"), linewidth = 1.15) +
  geom_segment(
    data = jump,
    aes(x = x, xend = xend, y = y, yend = yend),
    inherit.aes = FALSE,
    color = jump$ink,
    linewidth = 1.15
  ) +
  scale_color_manual(values = c("Least squares" = iron, "Logistic" = weaver)) +
  scale_y_continuous(limits = c(-0.05, 1.12), breaks = c(0, 0.5, 1)) +
  labs(x = "LoyalCH", y = NULL, color = NULL) +
  theme_mlba(13) +
  theme(legend.position = "bottom", legend.text = element_text(size = 11, color = weaver))
```
:::
::::

::: {.fragment}
<p class="today-note">A +0.1 step in loyalty is a taller step on the curve near 0.2 (red) than near 0.8 (teal). Same coefficient. Different change in $p$.</p>
:::

::: {.notes}
ISLR interpreting coefficients — plant this before the formal odds-multiplier slide. Association, not cause.

The two segments are the click. Do not quote a new decimal for the size of the jump unless someone asks you to read it off the curve. The point is that the jumps differ.
:::

## What to remember so far

:::: {.two-box}
::: {.fragment}
::: {.box-card}
<p class="card-h">Remember</p>
<ul>
<li>Brand choice is a yes/no. The mean of $Y$ is a share, not a shopper-level chance.</li>
<li>A straight line on 0/1 can leave (0, 1). On OJ, about <strong>0.036 to 1.051</strong>.</li>
<li>Odds and log-odds are the language. Logistic maps a linear score back to a probability in (0, 1).</li>
<li>A logistic coefficient is not a linear-probability slope.</li>
</ul>
:::
:::
::: {.fragment}
::: {.box-card .alt}
<p class="card-h">Next</p>
<ul>
<li>Fit <code>glm</code> on OJ.</li>
<li>Read loyalty and the price gap as odds multipliers. Association, not a cause.</li>
<li>Score <strong>one</strong> new visit with $\hat p$. No cutoff today.</li>
</ul>
:::
:::
::::

::: {.notes}
Mid-hour checkpoint, same job as Session 04’s “what to remember” before the next act. Click 1: remember. Click 2: next. Then the band into the fit. Keep it short.
:::

## Fit on OJ — then read carefully {.section-start}

<p class="act-kicker">The fit</p>

<p class="setup-line">The curve is no longer a sketch. We fit it on this file, then we read loyalty and the price gap as association — not as a cause.</p>

::: {.notes}
Section start. Short R from here. No confusion matrix. No other dataset. ISLR estimation, light.
:::

## One call — glm with a binomial family

- Same grammar as `lm`. Different family: binary CH versus MM.
- Start with LoyalCH. Then add PriceDiff.
- The fitted object stores coefficients on the **logit** scale.

::: {.formula}
$$
\log\left(\frac{p_i}{1-p_i}\right)
=
\beta_0 + \beta_1\,\mathrm{LoyalCH}_i + \beta_2\,\mathrm{PriceDiff}_i
$$
:::

```{r}
#| echo: true
fit1 <- glm(y ~ LoyalCH, data = oj, family = binomial)
fit2 <- glm(y ~ LoyalCH + PriceDiff, data = oj, family = binomial)
round(coef(fit2), 2)
```

<p class="q-prompt">After we fit, what do we ask the model for a new row — a class label, or a probability?</p>

::: {.fragment}
<p class="today-note"><strong>Probability first</strong> this hour. A yes/no label is a later business choice.</p>
:::

::: {.notes}
`y` = 1 for CH. Prediction-first. Do not call `table()` on a predicted class. ISLR estimation setup without a theory dump.

You say: probability first. Labels and cutoffs wait.
:::

## Example on OJ — one predictor, measured coefficients

:::: {.columns}
::: {.column width="46%"}
```{=html}
<table class="cb">
  <thead><tr><th>Term</th><th>Estimate</th><th>Scale</th></tr></thead>
  <tbody>
    <tr><td>Intercept</td><td>`r r2(coef(fit1)[1])`</td><td>logit</td></tr>
    <tr><td>LoyalCH</td><td>`r r2(coef(fit1)[2])`</td><td>logit</td></tr>
  </tbody>
</table>
```

<p class="caption">On the logit scale: $\hat\eta = `r r2(coef(fit1)[1])` + `r r2(coef(fit1)[2])`\,\mathrm{LoyalCH}$.</p>

- Positive LoyalCH coefficient: a higher loyalty score goes with higher log-odds of CH.
- You already saw the S-curve these two numbers draw.
:::
::: {.column width="54%"}
```{r}
#| fig-width: 5.8
#| fig-height: 3.5
#| fig-alt: "Small fitted S-curve of P(CH) against LoyalCH from the one-predictor model."
ggplot(grid_loyal, aes(LoyalCH, p_hat)) +
  geom_line(color = red, linewidth = 1.15) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed", color = iron, linewidth = 0.3) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
  labs(x = "LoyalCH", y = "Fitted P(CH)") +
  theme_mlba(13)
```
:::
::::

<p class="q-prompt">At LoyalCH = 0, is $\hat p$ near 0, near 0.5, or near 1?</p>

::: {.fragment}
<p class="today-note">Near 0. The intercept is about −2.76 on the logit scale, so loyalty has to rise before CH becomes likely.</p>
:::

::: {.notes}
Exact numbers. $\hat p = e^{-2.76}/(1+e^{-2.76})$ is low (about 0.06 — you may say “near 0”; do not build a lesson on the third digit). Do not invent standard errors on the glass. Optional only if a student asks, and then from `summary(fit1)`, not from memory.

`coef(fit1)` already ran.
:::

## What sentence is a coefficient allowed to support?

- The LoyalCH coefficient: associated change in the **log-odds** of CH for a one-unit rise in loyalty, holding PriceDiff fixed — once PriceDiff is in the model.
- Exponentiate for an odds multiplier: $e^{\hat\beta_j}$.
- Still an **association** in a fitted predictor. Not “raise loyalty and you cause CH.”

::: {.formula}
$$
\frac{\mathrm{odds}(\mathrm{CH}\mid x_j+1)}{\mathrm{odds}(\mathrm{CH}\mid x_j)}
=
e^{\beta_j}
\quad\text{(other predictors fixed)}.
$$
:::

```{r}
#| echo: true
round(exp(round(coef(fit2), 2)), 3)
```

<p class="q-prompt">If $e^{\hat\beta_1} > 1$, what sentence are you allowed to say in a meeting?</p>

::: {.fragment}
<p class="today-note">Allowed: “In this fit, higher loyalty goes with higher odds of CH, holding the price gap fixed.” Not allowed: “If we raise loyalty, we cause CH.”</p>
:::

::: {.notes}
ISLR interpreting coefficients. Lock the honesty. Prediction-first.

`exp(coef(fit2))` is large for LoyalCH because a one-unit move in LoyalCH runs the whole 0–1 score. Say that if someone flinches at the size. Do not turn it into a causal lift.
:::

## Example on OJ — loyalty and price gap together

:::: {.columns}
::: {.column width="52%"}
```{=html}
<table class="cb">
  <thead>
    <tr><th>Term</th><th>Estimate</th><th>exp(estimate)</th></tr>
  </thead>
  <tbody>
    <tr><td>Intercept</td><td>`r r2(b2_glass[1])`</td><td>`r sprintf("%.3f", odds_mult[1])`</td></tr>
    <tr><td>LoyalCH</td><td>`r r2(b2_glass[2])`</td><td>`r sprintf("%.3f", odds_mult[2])`</td></tr>
    <tr><td>PriceDiff</td><td>`r r2(b2_glass[3])`</td><td>`r sprintf("%.3f", odds_mult[3])`</td></tr>
  </tbody>
</table>
```

<p class="caption">$\hat\eta = `r r2(coef(fit2)[1])` + `r r2(coef(fit2)[2])`\,\mathrm{LoyalCH} + `r r2(coef(fit2)[3])`\,\mathrm{PriceDiff}$. Estimates on the logit scale. $e^{\hat\beta}$ uses those rounded estimates.</p>

- PriceDiff = sale price MM − sale price CH. When MM is more expensive relative to CH, the log-odds of CH rise.
- The LoyalCH coefficient stays large after the price gap is in the model.
:::
::: {.column width="48%"}
```{r}
#| fig-width: 5.4
#| fig-height: 3.55
#| fig-alt: "Boxplots of PriceDiff for CH buyers and MM buyers."
ggplot(oj, aes(Purchase, PriceDiff, fill = Purchase)) +
  geom_boxplot(width = 0.55, color = weaver, outlier.size = 0.7, outlier.alpha = 0.35) +
  geom_hline(yintercept = 0, linetype = "dashed", color = iron, linewidth = 0.35) +
  scale_fill_manual(values = brand_fill) +
  labs(x = NULL, y = "PriceDiff (MM − CH), dollars",
       caption = "Higher among CH buyers. No cutoff on this chart.") +
  theme_mlba(13)
```
:::
::::

<p class="q-prompt">Holding loyalty fixed, does a bigger MM–CH price gap push $\hat p$ for CH up or down?</p>

::: {.fragment}
<p class="today-note">Up. The PriceDiff estimate is about +`r r2(coef(fit2)[3])` on the logit scale.</p>
:::

::: {.notes}
Exact measured coefficients: intercept ≈ −3.25, LoyalCH ≈ 6.40, PriceDiff ≈ 2.86. No third predictor this hour.

The boxplot is the picture that PriceDiff runs higher for CH. Do not annotate a new mean on the glass. Association, not “raise the gap and you cause the switch.”
:::

## How does R pick those numbers?

- Not least squares. Logistic regression uses **maximum likelihood**: choose $\beta$ so the observed CH and MM purchases are as likely as possible under the model.
- You need the name and the idea. You do not need a derivation.
- `glm(..., family = binomial)` does this for you.

::: {.formula}
Choose $\beta$ to make the purchases in this file as likely as the model allows.
:::

<p class="q-prompt">If two coefficient vectors gave the same in-sample probabilities, would we care which one `glm` prints?</p>

::: {.fragment}
<p class="today-note">Not for this hour. Trust `glm`. What we care about next is $\hat p$ for a new visit.</p>
:::

::: {.notes}
ISLR maximum likelihood — one slide, not a theory dump. Do not open iteratively reweighted least squares or Fisher scoring. No likelihood formula beyond the one English line.
:::

## Example on OJ — loyalty and price, side by side

:::: {.columns}
::: {.column width="50%"}
```{r}
#| fig-width: 5.6
#| fig-height: 3.7
#| fig-alt: "Boxplot of LoyalCH by brand, CH higher than MM."
ggplot(oj, aes(Purchase, LoyalCH, fill = Purchase)) +
  geom_boxplot(width = 0.55, color = weaver, outlier.size = 0.7, outlier.alpha = 0.3) +
  scale_fill_manual(values = brand_fill) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(x = NULL, y = "LoyalCH", title = NULL,
       caption = "Means about 0.72 (CH) and 0.32 (MM).") +
  theme_mlba(13)
```
:::
::: {.column width="50%"}
```{r}
#| fig-width: 5.6
#| fig-height: 3.7
#| fig-alt: "Boxplot of PriceDiff by brand, higher for CH buyers."
ggplot(oj, aes(Purchase, PriceDiff, fill = Purchase)) +
  geom_boxplot(width = 0.55, color = weaver, outlier.size = 0.7, outlier.alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", color = iron, linewidth = 0.35) +
  scale_fill_manual(values = brand_fill) +
  labs(x = NULL, y = "PriceDiff (dollars)",
       caption = "Higher among CH buyers.") +
  theme_mlba(13)
```
:::
::::

- Together, loyalty and the price gap are why the two-predictor fit is the workhorse for a new visit today.

<p class="q-prompt">If you could keep only one predictor for a quick shelf check, which one — and why is “both” better for $\hat p$?</p>

::: {.notes}
Sit briefly, then move to scoring. LoyalCH separates the brands more sharply in the picture; PriceDiff still moves $\hat p$ once loyalty is held fixed. Both is the scoring model today.

No ROC, no AUC, no confusion matrix. Separation on a boxplot is not a classification-metrics week.
:::

## One new shopper {.section-start}

<p class="act-kicker">A prediction</p>

<p class="setup-line">One new visit. A loyalty score and a price gap go in. A probability comes out. Then we stop.</p>

::: {.notes}
Prediction-first close of the ISLR spine. Do not open LDA, QDA, or KNN. Do not threshold.
:::

## What is this shopper’s chance of buying CH?

:::: {.columns}
::: {.column width="46%"}
- A new row: a loyalty score and a price gap.
- $\hat\eta = \hat\beta_0 + \hat\beta_1\,\mathrm{LoyalCH} + \hat\beta_2\,\mathrm{PriceDiff}$.
- $\hat p = e^{\hat\eta}/(1+e^{\hat\eta})$.

<p class="p-hero">`r r2(new1$p_hat)`</p>
<p class="caption">LoyalCH = 0.5 and PriceDiff = +0.2, from the two-predictor fit.</p>

```{r}
#| echo: true
new <- data.frame(LoyalCH = 0.5, PriceDiff = 0.2)
round(predict(fit2, new, type = "response"), 2)
```
:::
::: {.column width="54%"}
```{r}
#| fig-width: 6.2
#| fig-height: 3.9
#| fig-alt: "Predicted probability of CH against PriceDiff when LoyalCH is held at 0.5, with the +0.2 shopper marked."
price_grid <- data.frame(
  PriceDiff = seq(-0.4, 0.5, length.out = 200),
  LoyalCH = 0.5
)
price_grid$p_hat <- as.numeric(predict(fit2, newdata = price_grid, type = "response"))
ggplot(price_grid, aes(PriceDiff, p_hat)) +
  geom_line(color = weaver, linewidth = 1.15) +
  geom_point(data = new1, aes(PriceDiff, p_hat), color = red, size = 3.4) +
  geom_hline(yintercept = c(0, 1), linetype = "dashed", color = iron, linewidth = 0.3) +
  scale_y_continuous(limits = c(0, 1), breaks = c(0, 0.5, 1)) +
  labs(x = "PriceDiff (dollars)", y = "Predicted P(CH)",
       caption = "LoyalCH held at 0.5. The dot is PriceDiff = +0.2.") +
  theme_mlba(13)
```
:::
::::

<p class="q-prompt">Before the number: loyalty 0.5 and a price gap of +0.2. Closer to 0.3, or closer to 0.7?</p>

::: {.notes}
Reveal `predict(..., type = "response")` ≈ 0.63. Closer to 0.7 than to 0.3. ISLR “making predictions.” Do not threshold. This first triple is locked.

The curve holds loyalty at 0.5 so the room sees the price gap move $\hat p$. It is the two-predictor fit, not a new model.
:::

## Example on OJ — three visits, three chances

:::: {.columns}
::: {.column .spaced-list width="42%"}
- LoyalCH = 0.5, PriceDiff = **+0.2** → $\hat p$ ≈ **`r r2(new3$p_hat[1])`**
- LoyalCH = 0.5, PriceDiff = **−0.2** → $\hat p$ ≈ **`r r2(new3$p_hat[2])`**
- LoyalCH = 0.9, PriceDiff = **0** → $\hat p$ ≈ **`r r2(new3$p_hat[3])`**
- Same loyalty, opposite price gap: different chances. High loyalty dominates when the price gap is flat.

```{r}
#| echo: true
visits <- data.frame(
  LoyalCH   = c(0.5, 0.5, 0.9),
  PriceDiff = c(0.2, -0.2, 0)
)
round(predict(fit2, visits, type = "response"), 2)
```
:::
::: {.column width="58%"}
```{r}
#| fig-width: 6.4
#| fig-height: 4.0
#| fig-alt: "Three bars for predicted P(CH): about 0.63, 0.35, and 0.92."
shop_cols <- c(weaver, iron, teal)
ggplot(new3, aes(who, p_hat, fill = who)) +
  geom_col(width = 0.68) +
  geom_text(aes(label = sprintf("%.2f", p_hat)), vjust = -0.35,
            fontface = "bold", color = weaver, size = 5) +
  scale_fill_manual(values = shop_cols) +
  scale_y_continuous(limits = c(0, 1.12), breaks = seq(0, 1, 0.25), expand = c(0, 0)) +
  labs(x = NULL, y = "Predicted P(CH)") +
  theme_mlba(12)
```
:::
::::

<p class="q-prompt">In the first two rows, which lever moved $\hat p$ — loyalty, or the price gap?</p>

::: {.fragment}
<p class="today-note">The first two rows hold loyalty at 0.5. The price gap moves $\hat p$ from about 0.63 to about 0.35. The third row is loyalty: about 0.92 when LoyalCH is 0.9 and the gap is 0.</p>
:::

::: {.notes}
Exact three triples, locked. No class labels. No 0.5 cutoff. Click: 0.63 versus 0.35 is the price gap; 0.92 is high loyalty with a flat gap.
:::

## A chance — not a shelf decision yet

- $\hat p$ is our estimate of P(CH | this visit’s features).
- It is not yet “always predict CH if $\hat p > 0.5$.” That cutoff is a **business** choice. It depends on the costs of being wrong.
- Scoring hits and misses, and decision cutoffs, return **after fall break**.

<p class="q-prompt">If a false CH call costs the chain differently than a false MM call, should 0.5 still be automatic?</p>

::: {.fragment}
<p class="today-note">Park it. 0.5 is not the statistical answer. We come back to costs when we score hits and misses — not today.</p>
:::

::: {.notes}
Allowed deferral: one line, then stop. Forbidden on this slide: a confusion-matrix table, sensitivity, specificity, cutoff optimization, or a ROC drawing.

Week 7 on the course grid is thresholds. The lab for logistic R is after fall break. Do not invent a lab date.
:::

## Can / cannot — after today’s OJ hour {.can-cant}

:::: {.two-box}
::: {.box-card}
<p class="card-h">Can</p>
<ul>
<li>Turn loyalty and the price gap into a <strong>probability</strong> of CH that stays in (0, 1).</li>
<li>Read coefficients as log-odds, and $e^{\hat\beta}$ as an odds multiplier. Association.</li>
<li>Score a <strong>new</strong> visit with <code>predict(..., type = "response")</code>.</li>
</ul>
:::
::: {.box-card .alt}
<p class="card-h">Cannot, from this fit alone</p>
<ul>
<li>Prove that raising loyalty <strong>causes</strong> CH purchases.</li>
<li>Pick the “right” yes/no cutoff without costs. That waits.</li>
<li>Replace a holdout with a training-accuracy story.</li>
</ul>
:::
::::

::: {.notes}
Same honesty beat as Session 04. Prediction-first. No new chart.
:::

## When you use logistic for a real brand decision {.practice}

:::: {.do-dont}
::: {}
<h3>Dos</h3>
<ul>
<li>Code the yes/no outcome in one line. Here, CH = 1. Say what $p$ means in one sentence.</li>
<li>Use <code>type = "response"</code> when you want a probability for a new visit.</li>
<li>Judge the fit on visits you did not use to estimate the coefficients.</li>
<li>Say <strong>prediction</strong> and <strong>association</strong> when you read $e^{\hat\beta}$.</li>
<li>Use features you would know <strong>before</strong> the purchase. Loyalty and the price gap. Not an after-the-fact label.</li>
</ul>
:::
::: {}
<h3>Don’ts</h3>
<ul>
<li>Do not treat a straight line on 0/1 as safe when predictions leave (0, 1).</li>
<li>Do not ship 0.5 as “the statistical answer” without costs.</li>
<li>Do not paste a hit/miss table into a meeting that only asked for a probability.</li>
<li>Do not invent lift, accuracy, or other scores you did not measure on this file.</li>
<li>Do not swap in a different dataset mid-meeting. This hour is OJ.</li>
</ul>
:::
::::

::: {.notes}
Practitioner hat only. No MLOps. The “different dataset” line is the don’t: stay on OJ. Do not name a tour of other textbook files on the glass.
:::

## What you can do after today {.takeaways}

<ol class="today-list">
<li>State the OJ job: predict P(Purchase = CH) for a <strong>new</strong> store visit.</li>
<li>Explain why least squares on 0/1 can break the probability fence. On OJ: about 0.036 to 1.051.</li>
<li>Move from probability to odds to the logit, and write the logistic model in this file’s units.</li>
<li>Fit with <code>glm(..., family = binomial)</code>. Read $e^{\hat\beta}$ as an odds multiplier — association, not a cause.</li>
<li>Produce one visit’s $\hat p$ with <code>predict(..., type = "response")</code>, and refuse a cutoff until the costs show up.</li>
</ol>

::: {.notes}
In this file’s language. No topic chips required. The takeaway sentence, if you need one aloud: for a new store visit, logistic regression turns loyalty and the price gap into a probability of Citrus Hill that stays between 0 and 1.
:::

## Before next class {.next-slide}

<ol class="today-list">
<li>Midterm <strong>review</strong> is the next methods meeting. The <strong>midterm</strong> is the meeting after that. Closed notes. Weeks 1–5, including logistic.</li>
<li>Finish Homework 2 if it is still open. It is due at the end of the Sunday before midterm week, as on the schedule.</li>
<li>Reading already named: ISLR Chapter 4, the logistic pages (about 129–141).</li>
<li>The logistic <strong>R lab</strong> is after fall break. Bring a laptop that day, once it is posted.</li>
<li>Hits, misses, and decision cutoffs return with the classification-metrics week after the break.</li>
</ol>

::: {.notes}
Your eyes only. Slot for this deck: Wed 30 Sep 2026 discussion, continuing the Monday pair. Midterm review Mon 5 Oct. Midterm Wed 7 Oct. Fall break 11–15 Oct. Do not invent a logistic lab calendar date. The schedule says the lab is after fall break.

Homework 2: this outline’s lock is Sun 4 Oct, 11:59 pm. The materials-site schedule page currently lists Sun 28 Sep. Check Canvas before you say the due date aloud. Do not put either date on a slide title.

Do not open a confusion matrix here. One deferral line only. No book-page tour beyond the chapter already named on the glass.
:::
