---
title: "Lab 4: Predicting yearly spending from browsing"
subtitle: "70-467 · Wednesday 23 September"
author: "Andrés Castaño Zuluaga"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: right
    body-classes: "mlba-page"
    code-fold: false
    code-tools: true
execute:
  echo: true
  warning: false
  message: false
---

<!--
Student starter for the hour. Code is filled on purpose. Run each chunk
and read the comments. We will talk through only some of the Question
and Discuss prompts in the room. You choose one method and one lambda,
then score that choice once.
-->

# Introduction to the Business Problem

An advertising platform has yearly online spending and web browsing for
10,000 anonymous households. Spending is one row per household. Browsing
is a long file: a household, a site code, and a visit count. The names
are a third file, one domain per line. These are ordinary websites, not
pages on one store.

The outcome is yearly spending, in dollars. The median is $510. The
largest value is $401,338. Every household spent at least $1. Age,
income, and purchases are not in these files. The 1,000 site columns
are not in them either. You build those columns.

There are many sites, so there are many possible models. If we put
every site into one least-squares fit, the fit on these 10,000
households can look better as we add sites. That can be overfitting.
It does not mean the model will predict well for a new household.

The question for the hour is the one from class: how can we build a
good predictive model for spending when there are 1,000 sites, and how
can we avoid overfitting?

We will use ridge and the lasso. For each method we will read two
values of $\lambda$ from the cross-validation curve: the $\lambda$ with
the lowest cross-validated error, and the one-standard-error $\lambda$.
Then you will commit to one method and one $\lambda$. The number that
says whether you chose well is that fit's cross-validated error. An
in-sample $R^2$ is not that number.

Keep the case PDF open. The codebook is page 2.

**Before you run anything.** A count of 10 page views is a large share
for a household with 25 page views, and a very small share for a
household with about 48,000. Talk this through with the person next to
you. Why would a model on raw visit counts treat those two households
as if the site mattered the same amount?

# Outline of the analysis

1. **Load the three files.** Match each file to the codebook. Count
   households, browsing rows, and site names before you build anything.
2. **Build a share for each site.** A visit count is not comparable
   across households. The column we fit is that site's percent of the
   household's page views. The shares in a row sum to 100.
3. **See why the response is log(spend).** Spending in dollars has a
   long right tail. A few households would dominate a fit on that scale.
4. **Fit ridge and the lasso on the same folds.** Ridge shrinks every
   coefficient and keeps every site. The lasso can set coefficients to
   exactly zero. Seed 369 so the room is looking at the same curves.
5. **Read the two values of $\lambda$.** `lambda.min` is the lowest
   cross-validated error. `lambda.1se` is the largest $\lambda$ within
   one standard error of that minimum. The two rules will not keep the
   same websites.
6. **Read one coefficient at a share a household actually has.** A one
   percentage point share is large on this file. The median share among
   visitors is much smaller.
7. **Commit to one method and one $\lambda$, then score that choice
   once.**
8. **After class.** Elastic net, only after the choice in class is
   already written down.

# Load the three files

`Matrix` builds the household-by-site table without storing a zero for
every site a household did not open. `glmnet` fits ridge and the lasso
and chooses $\lambda$ by cross-validation.

```{r}
#| label: load
# Matrix::sparseMatrix stores only the nonzero shares.
# glmnet::cv.glmnet fits the penalized regression and the 10 folds.
library(Matrix)
library(glmnet)

# Spending: one row per household. Columns are id and spend (dollars).
spend <- read.csv("data/browser-totalspend.csv")

# Browsing, long form: one row is one household and one site they opened.
# Columns are id, site (1 to 1,000), and visits (page views).
web <- read.csv("data/browser-domains.csv")

# Names: no header. Line 1 is site code 1. scan() reads one name per line.
sites <- scan("data/browser-sites.txt", what = "character", quiet = TRUE)

# Question: How many households? How many rows in the browsing file?
# How many site names? Do the names line up with site codes 1 to 1,000?
nrow(spend)
nrow(web)
length(sites)
head(spend)
head(web)
head(sites, 5)

# Interpret: 10,000 households, about 2.27 million browsing rows, and
# 1,000 names. Each household opened many sites, so the browsing file
# has many rows per household.
#
# Question: What would go wrong if you treated each browsing row as a
# separate household when you predict yearly spending?
```

# Build a share for each site

A visit count is not comparable across households. Some households have
25 page views. Some have about 48,000. The column we fit is that site's
share of the household's page views. After this step, every household's
shares sum to 100.

```{r}
#| label: shares
# A count of 10, as a percent, at the two ends of the page-view range
# named in the case. Run these two lines before you build the matrix.
100 * 10 / 25
100 * 10 / 48786

# Total page views for each household. The name of each total is the id.
tot <- tapply(web$visits, web$id, sum)

# Share, in percent, of that household's page views. as.numeric() keeps
# the value a plain number, which sparseMatrix requires.
web$pct <- as.numeric(100 * web$visits / tot[as.character(web$id)])

# Row number in the spending file for each browsing row. Household ids
# are not 1, 2, 3, ..., so we look them up.
id_index <- setNames(seq_len(nrow(spend)), as.character(spend$id))
i <- as.integer(id_index[as.character(web$id)])

# i is the household row. j is the site code. x is the share, in percent.
# The table has 10,000 rows and 1,000 columns. A site the household
# did not open stays zero. Column names are the domains, in site-code order.
X <- sparseMatrix(
  i = i,
  j = as.integer(web$site),
  x = web$pct,
  dims = c(nrow(spend), length(sites)),
  dimnames = list(NULL, sites)
)

dim(X)
range(Matrix::rowSums(X))
summary(Matrix::rowSums(X > 0))

# Why? glmnet can take this sparse matrix. A dense table of 10,000
# rows and 1,000 columns would store a zero for every site a household
# did not visit. Most of those cells are zero. The median household
# opened about 219 of the 1,000 sites.
#
# Interpret: Each row sums to 100. That is the household's page views,
# split across the sites they visited.
#
# Question: Why is a raw count of 10 a poor column to put in the model?
#
# Discuss: Two households can have the same share on a site and very
# different visit counts. Which fact do we want the coefficient to use?
```

# Why the response is log(spend)

In dollars, the median is $510 and the maximum is $401,338. Least
squares squares the errors, so a few very large households would
dominate a fit on the dollar scale. Every household spent at least $1,
so $\log(\text{spend})$ is defined. The smallest log is 0, the household
that spent $1.

```{r}
#| label: response
#| fig-width: 8
#| fig-height: 3.5
y <- log(spend$spend)

par(mfrow = c(1, 2))
hist(spend$spend, breaks = 40, main = "Spending, dollars", xlab = "dollars")
hist(y, breaks = 40, main = "log(spend)", xlab = "log dollars")

# The standard deviation of log(spend) is the scale of the outcome.
# Predicting the mean for every household has RMSE equal to that
# standard deviation, up to a small difference between sd() and the
# RMSE formula. sd() divides by n - 1. The RMSE line divides by n.
sd(y)
sqrt(mean((y - mean(y))^2))

# Interpret: The dollar histogram has a long right tail. The log
# histogram does not. log(spend) is the response we fit.
#
# Question: The standard deviation of log(spend) is about 1.67. What
# does a cross-validated RMSE near that number say about the fit?
#
# Discuss: Predicting the mean of log(spend) misses by about 1.67.
# How much smaller than 1.67 should the cross-validated RMSE be
# before you would use the sites?
```

# Fit ridge and the lasso

Ridge (`alpha = 0`) shrinks coefficients toward zero and keeps every
site. The lasso (`alpha = 1`) can set coefficients to exactly zero, so
some sites drop out. Both fits use 10 folds. The same seed is set
before each fit, so both use the same folds.

`standardize = TRUE` scales the shares before the penalty. The
coefficients are returned on the original scale: the change in
log(spend) for a one percentage point share, other shares held fixed.

```{r}
#| label: cv-fits
# Why a fixed seed? The fold assignment is random. Fixing it means
# everyone in the room is looking at the same curves. Set it again
# before the lasso so that fit does not inherit the random draws
# already used inside the ridge fit.
set.seed(369)
cv_ridge <- cv.glmnet(X, y, alpha = 0, nfolds = 10, standardize = TRUE)

set.seed(369)
cv_lasso <- cv.glmnet(X, y, alpha = 1, nfolds = 10, standardize = TRUE)
```

The vertical axis of each plot is cross-validated mean squared error
on log(spend). The left dotted line is `lambda.min`. The right dotted
line is `lambda.1se`. The numbers along the top of the lasso plot are
how many sites still have a nonzero coefficient at that $\lambda$.

```{r}
#| label: cv-plots
#| fig-width: 8
#| fig-height: 3.6
par(mfrow = c(1, 2))
plot(cv_ridge, main = "Ridge")
plot(cv_lasso, main = "Lasso")

# Interpret: As lambda increases, the penalty gets stronger and the
# cross-validated error eventually rises. The left line is the bottom
# of that curve. The right line is a stronger penalty that is still
# close to that bottom, within one standard error.
#
# Question: As lambda increases, what happens to the number of sites
# at the top of the lasso plot? What happens on the ridge plot?
#
# Discuss: Ridge does not print a shrinking site count along the top.
# Why not?
```

# Two values of lambda

`lambda.min` is the $\lambda$ with the lowest cross-validated mean
squared error. `lambda.1se` is the largest $\lambda$ still within one
standard error of that minimum.

The RMSE below is the square root of that mean squared error, on the
scale of log(spend). The site count leaves out the intercept. A
coefficient of zero means that site is out.

```{r}
#| label: lambda-table
# How many site coefficients are not zero at a chosen lambda.
n_sites <- function(fit, s) {
  b <- as.matrix(coef(fit, s = s))
  sum(b[rownames(b) != "(Intercept)", 1] != 0)
}

# Cross-validated RMSE at lambda.min or lambda.1se.
# cvm is mean squared error. The square root puts it next to the
# standard deviation of log(spend).
cv_rmse <- function(fit, s) {
  ix <- which.min(abs(fit$lambda - fit[[s]]))
  sqrt(fit$cvm[ix])
}

ridge_min_rmse <- cv_rmse(cv_ridge, "lambda.min")
ridge_1se_rmse <- cv_rmse(cv_ridge, "lambda.1se")
lasso_min_rmse <- cv_rmse(cv_lasso, "lambda.min")
lasso_1se_rmse <- cv_rmse(cv_lasso, "lambda.1se")

data.frame(
  method = c("ridge", "ridge", "lasso", "lasso"),
  rule = c("lambda.min", "lambda.1se", "lambda.min", "lambda.1se"),
  lambda = c(cv_ridge$lambda.min, cv_ridge$lambda.1se,
             cv_lasso$lambda.min, cv_lasso$lambda.1se),
  cv_rmse = c(ridge_min_rmse, ridge_1se_rmse, lasso_min_rmse, lasso_1se_rmse),
  sites = c(
    n_sites(cv_ridge, "lambda.min"), n_sites(cv_ridge, "lambda.1se"),
    n_sites(cv_lasso, "lambda.min"), n_sites(cv_lasso, "lambda.1se")
  )
)

# Null RMSE: the typical miss if we ignore every site and predict the
# mean of log(spend).
sqrt(mean((y - mean(y))^2))

# Interpret: On this seed, ridge at lambda.min and the lasso at
# lambda.min have cross-validated RMSE values that agree to about
# three decimals, near 1.575. Predicting the mean is about 1.668.
# Ridge keeps all 1,000 sites. The lasso at lambda.min keeps 206
# sites. The lasso at lambda.1se keeps 86. Its RMSE is about 1.590.
# The gap in mean squared error from the lasso minimum sits inside
# one standard error of that minimum.
#
# Question: If the job is the score, which of these four fits can you
# defend? If the job is a short list of sites, which fits are still
# available?
#
# Discuss: The one-standard-error lambda is not a better score. What
# is it for?
#
# Business implication: A short list of sites and a score for a new
# household are two uses of these fits. Name which one you are using
# when you pick.
```

# Read a coefficient at a share a household actually has

A coefficient from `glmnet` is the change in log(spend) for a one
percentage point share on that site, other shares held fixed.
$\exp(\text{coefficient})$ is the factor on spending for that one-point
share. Households do not put one percent of their page views on these
sites. Read the coefficient at a share you can see in the matrix, not
only at one percentage point.

We read the largest coefficient on the lasso at `lambda.1se`, the
short list.

```{r}
#| label: one-coefficient
# Coefficients at the short list. Drop the intercept. Drop the zeros.
b <- as.matrix(coef(cv_lasso, s = "lambda.1se"))
b <- b[rownames(b) != "(Intercept)", 1]
b <- b[b != 0]

# The site whose coefficient is farthest from zero.
site_name <- names(b)[which.max(abs(b))]
site_coef <- b[site_name]
site_name
site_coef

# Shares on that site among households who opened it. Drop the zeros.
# Those households did not open the site.
j <- match(site_name, colnames(X))
visited <- as.numeric(X[, j])
visited <- visited[visited > 0]
median_share <- median(visited)
median_share
quantile(visited, 0.99)

# Change in log(spend) at a one-point share, and at the median share.
# exp() turns that change into a factor on spending.
site_coef * 1
site_coef * median_share
exp(site_coef * 1)
exp(site_coef * median_share)

# Interpret: On this seed the largest coefficient at the lasso
# one-standard-error fit is about 0.83, on victoriassecret.com.
# exp(0.83) is about 2.3, which is the factor for a one percentage
# point share. Among the households who visit that site, the median
# share is about 0.042 percent. The 99th percentile is still under
# 1 percent. The factor at the median share is about 1.035.
#
# Question: Which of those two factors would you quote if someone
# asked what the fit says about a typical visit to that site?
#
# Discuss: The shares in a row sum to 100. What is awkward about
# reading the coefficient as the change from moving one percentage
# point onto that site, holding the other shares fixed?
```

**Before you commit.** The four cross-validated errors are close. The
site counts are not. Talk with the person next to you. Are you choosing
a score for a new household, a short list of sites, or both? Which
method and which $\lambda$ follow from that?

# Commit to one model and score it once

Write your choice in the next chunk, then run the score. Do not change
the choice after you have seen that number.

`alpha = 0` is ridge. `alpha = 1` is the lasso. `chosen_s` is
`"lambda.min"` or `"lambda.1se"`.

The lines below are a starting point so the file runs. Replace them
with the choice you just made.

```{r}
#| label: commit
# Starting point only, so the chunk runs. Replace both lines.
chosen_alpha <- 1
chosen_s <- "lambda.1se"
chosen_fit <- if (chosen_alpha == 1) cv_lasso else cv_ridge

c(alpha = chosen_alpha, rule = chosen_s)
chosen_fit[[chosen_s]]
cv_rmse(chosen_fit, chosen_s)
n_sites(chosen_fit, chosen_s)

# The null RMSE again, so the score sits next to "predict the mean."
sqrt(mean((y - mean(y))^2))

# Interpret: Put your cross-validated RMSE next to 1.67, the RMSE
# from predicting the mean of log(spend), and next to the other three
# fits in the table above. The site count is part of the choice.
#
# Question: Suppose a classmate now tries the other three fits and
# keeps the one with the smallest cross-validated RMSE. What happens
# to the claim that your number is the score of a choice you made
# before looking?
#
# Business implication: One cross-validated error is what you can
# take as the score of the fit you chose. A search over the four
# numbers, after you have seen all four, is still a search.
```

# After class: elastic net

Work this after you have committed. Do not change `chosen_alpha` or
`chosen_s`.

Elastic net uses `alpha = 0.5`, between ridge (`alpha = 0`) and the
lasso (`alpha = 1`). It was the last question in an older lab on these
files. It is not on the lecture slides.

```{r}
#| label: after-class-enet
# Same folds and the same standardization. alpha is 0.5.
set.seed(369)
cv_enet <- cv.glmnet(X, y, alpha = 0.5, nfolds = 10, standardize = TRUE)

data.frame(
  rule = c("lambda.min", "lambda.1se"),
  lambda = c(cv_enet$lambda.min, cv_enet$lambda.1se),
  cv_rmse = c(cv_rmse(cv_enet, "lambda.min"), cv_rmse(cv_enet, "lambda.1se")),
  sites = c(n_sites(cv_enet, "lambda.min"), n_sites(cv_enet, "lambda.1se"))
)

# Question: How does the site count at lambda.min, and at lambda.1se,
# sit between the ridge fit and the lasso fit you already ran?
#
# Question: Would you replace the choice you committed to in class
# with one of these two? Why not?
```
