---
title: "Lab 4: Predicting yearly spending from browsing"
subtitle: "70-467 · Wednesday 23 September"
author: "Andrés Castaño Zuluaga"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: right
    body-classes: "mlba-page"
    code-fold: false
    code-tools: true
execute:
  echo: true
  warning: false
  message: false
---

<!--
Student starter for the hour. Code is filled on purpose. Run each chunk
and read the comments. The choice of method and of lambda is yours.
Score that choice once.
-->

# Introduction to the Business Problem

An advertising platform has yearly online spending and web browsing for
10,000 anonymous households. Spending is one row per household. Browsing
is a long file: a household, a site code, and a visit count. The names
are a third file, one domain per line. These are ordinary websites, not
pages on one store.

The outcome is yearly spending, in dollars. The median is $510. The
largest value is $401,338. Every household spent at least $1. Age,
income, and purchases are not in these files. The 1,000 site columns
are not in them either. You build those columns.

There are many sites, so there are many possible models. If we put
every site into one least-squares fit, the fit on these 10,000
households can look better as we add sites. That can be overfitting.
It does not mean the model will predict well for a new household.

The question for the hour is the one from class: how can we build a
good predictive model for spending when there are 1,000 sites, and how
can we avoid overfitting?

We will use ridge and the lasso. For each method we will read two
values of $\lambda$ from the cross-validation curve: the $\lambda$ with
the lowest cross-validated error, and the one-standard-error $\lambda$.
Then you will commit to one method and one $\lambda$. The number that
says whether you chose well is that fit's cross-validated error. An
in-sample $R^2$ is not that number.

Keep the case PDF open. The codebook is page 2.

**Before you run anything.** A count of 10 page views is a large share
for a household with 25 page views, and a very small share for a
household with about 48,000. Talk this through with the person next to
you. Why would a model on raw visit counts treat those two households
as if the site mattered the same amount?

# Roadmap

1. **Load the three files.** Match the codebook.
2. **Build a share for each site.** Percent of that household's page
   views. The shares sum to 100.
3. **See why the response is log(spend).** The dollar distribution has
   a long right tail.
4. **Fit ridge and the lasso.** Same folds, seed 369.
5. **Read the two values of $\lambda$.** They will not keep the same
   websites.
6. **Read a coefficient at a share a household actually has.** One
   percentage point is a large share on this file.
7. **Commit to one model and score it once.**
8. **After class.** Elastic net.

# Load the three files

```{r}
#| label: load
library(Matrix)
library(glmnet)

spend <- read.csv("data/browser-totalspend.csv")
web <- read.csv("data/browser-domains.csv")
sites <- scan("data/browser-sites.txt", what = "character", quiet = TRUE)

# Question: How many households? How many rows in the browsing file?
# How many site names? Do the names line up with site codes 1 to 1,000?
nrow(spend)
nrow(web)
length(sites)
head(spend)
head(web)
head(sites, 5)
```

# Build a share for each site

A visit count is not comparable across households. Some households have
25 page views. Some have about 48,000. The column we fit is that site's
share of the household's page views.

```{r}
#| label: shares
# A count of 10, as a percent, at the two ends of the page-view range
# named in the case.
100 * 10 / 25
100 * 10 / 48786

tot <- tapply(web$visits, web$id, sum)
web$pct <- as.numeric(100 * web$visits / tot[as.character(web$id)])

id_index <- setNames(seq_len(nrow(spend)), as.character(spend$id))
i <- as.integer(id_index[as.character(web$id)])

X <- sparseMatrix(
  i = i,
  j = as.integer(web$site),
  x = web$pct,
  dims = c(nrow(spend), length(sites)),
  dimnames = list(NULL, sites)
)

# Why? glmnet can take this sparse matrix. A dense table of 10,000
# rows and 1,000 columns would store a zero for every site a household
# did not visit.
dim(X)
range(Matrix::rowSums(X))
summary(Matrix::rowSums(X > 0))

# Interpret: Each row sums to 100. That is the household's page views,
# split across the sites they visited. The median household has a
# positive share on about 219 of the 1,000 sites.
#
# Question: Why is a raw count of 10 a poor column to put in the model?
#
# Discuss: Two households can have the same share on a site and very
# different visit counts. Which fact do we want the coefficient to use?
```

# Why the response is log(spend)

```{r}
#| label: response
#| fig-width: 8
#| fig-height: 3.5
y <- log(spend$spend)

par(mfrow = c(1, 2))
hist(spend$spend, breaks = 40, main = "Spending, dollars", xlab = "dollars")
hist(y, breaks = 40, main = "log(spend)", xlab = "log dollars")

# The standard deviation of log(spend) is the scale of the outcome.
# Predicting the mean for every household has RMSE equal to that
# standard deviation, up to a small difference between sd() and the
# RMSE formula.
sd(y)
sqrt(mean((y - mean(y))^2))

# Interpret: In dollars, the median is $510 and the maximum is
# $401,338. A few households would dominate a least-squares fit on
# the dollar scale. log(spend) is the response we fit. Every household
# spent at least $1, so the log is defined. On this file the smallest
# log is 0.
#
# Question: The standard deviation of log(spend) is about 1.67. What
# does a cross-validated RMSE near that number say about the fit?
```

# Fit ridge and the lasso

Ridge (`alpha = 0`) shrinks coefficients toward zero and keeps every
site. The lasso (`alpha = 1`) can set coefficients to zero. We use the
same 10 folds and the same seed for both, and we standardize the
shares before the penalty. `glmnet` does that when `standardize = TRUE`.

```{r}
#| label: cv-fits
set.seed(369)
cv_ridge <- cv.glmnet(X, y, alpha = 0, nfolds = 10, standardize = TRUE)

set.seed(369)
cv_lasso <- cv.glmnet(X, y, alpha = 1, nfolds = 10, standardize = TRUE)

# Why a fixed seed? The fold assignment is random. Fixing it means
# everyone in the room is looking at the same curves.
```

```{r}
#| label: cv-plots
#| fig-width: 8
#| fig-height: 3.6
par(mfrow = c(1, 2))
plot(cv_ridge, main = "Ridge")
plot(cv_lasso, main = "Lasso")

# Interpret: The vertical axis is cross-validated mean squared error
# on log(spend). The left dotted line is lambda.min, the value with
# the lowest cross-validated error. The right dotted line is
# lambda.1se, the largest lambda within one standard error of that
# minimum.
#
# Question: As lambda increases, what happens to the number of sites
# at the top of the lasso plot? What happens on the ridge plot?
```

# Two values of lambda

```{r}
#| label: lambda-table
n_sites <- function(fit, s) {
  b <- as.matrix(coef(fit, s = s))
  sum(b[rownames(b) != "(Intercept)", 1] != 0)
}
cv_rmse <- function(fit, s) {
  ix <- which.min(abs(fit$lambda - fit[[s]]))
  sqrt(fit$cvm[ix])
}

ridge_min_rmse <- cv_rmse(cv_ridge, "lambda.min")
ridge_1se_rmse <- cv_rmse(cv_ridge, "lambda.1se")
lasso_min_rmse <- cv_rmse(cv_lasso, "lambda.min")
lasso_1se_rmse <- cv_rmse(cv_lasso, "lambda.1se")

data.frame(
  method = c("ridge", "ridge", "lasso", "lasso"),
  rule = c("lambda.min", "lambda.1se", "lambda.min", "lambda.1se"),
  lambda = c(cv_ridge$lambda.min, cv_ridge$lambda.1se,
             cv_lasso$lambda.min, cv_lasso$lambda.1se),
  cv_rmse = c(ridge_min_rmse, ridge_1se_rmse, lasso_min_rmse, lasso_1se_rmse),
  sites = c(
    n_sites(cv_ridge, "lambda.min"), n_sites(cv_ridge, "lambda.1se"),
    n_sites(cv_lasso, "lambda.min"), n_sites(cv_lasso, "lambda.1se")
  )
)

# Null RMSE: the typical miss if we ignore every site and predict the
# mean of log(spend).
sqrt(mean((y - mean(y))^2))

# Interpret: On this seed, ridge at lambda.min and the lasso at
# lambda.min have cross-validated RMSE values that agree to about
# three decimals, near 1.575. Predicting the mean is about 1.668.
# Ridge keeps all 1,000 sites. The lasso at lambda.min keeps 206
# sites. The lasso at lambda.1se keeps 86. Its RMSE is about 1.590.
# The gap in mean squared error from the lasso minimum sits inside
# one standard error of that minimum.
#
# Question: If the job is the score, which of these four fits can you
# defend? If the job is a short list of sites, which fits are still
# available?
#
# Discuss: The one-standard-error lambda is not a better score. What
# is it for?
#
# Business implication: A list of sites and a score for a new
# household are different uses of the same fits. Say which use you
# are serving before you pick.
```

# Read a coefficient at a share a household actually has

A coefficient from `glmnet` is the change in log(spend) for a one
percentage point share on that site, other shares held fixed. Households
do not put one percent of their page views on these sites.

```{r}
#| label: one-coefficient
# Lasso at lambda.1se, the short list. Take the coefficient with the
# largest absolute value and read it at the median share among
# households who visit that site.
b <- as.matrix(coef(cv_lasso, s = "lambda.1se"))
b <- b[rownames(b) != "(Intercept)", 1]
b <- b[b != 0]
site_name <- names(b)[which.max(abs(b))]
site_coef <- b[site_name]
site_name
site_coef

j <- match(site_name, colnames(X))
visited <- as.numeric(X[, j])
visited <- visited[visited > 0]
median_share <- median(visited)
median_share
quantile(visited, 0.99)

# Change in log(spend) at a one-point share, and at the median share.
site_coef * 1
site_coef * median_share
exp(site_coef * 1)
exp(site_coef * median_share)

# Interpret: On this seed the largest coefficient at the lasso
# one-standard-error fit is about 0.83, on victoriassecret.com.
# exp(0.83) is about 2.3, which is the factor for a one percentage
# point share. Among the households who visit that site, the median
# share is about 0.042 percent. The 99th percentile is still under
# 1 percent. The factor at the median share is about 1.035.
#
# Question: Which of those two factors would you quote if someone
# asked what the fit says about a typical visit to that site?
#
# Discuss: The shares in a row sum to 100. What is awkward about
# reading the coefficient as the change from moving one percentage
# point onto that site, holding the other shares fixed?
```

**Before you commit.** The four cross-validated errors are close. The
site counts are not. Talk with the person next to you. Are you choosing
a score for a new household, a short list of sites, or both? Which
method and which $\lambda$ follow from that?

# Commit to one model and score it once

Write your choice in the next chunk, then run the score. Do not go
back and change the choice after you have seen that number.

`alpha = 0` is ridge. `alpha = 1` is the lasso. `chosen_s` is
`"lambda.min"` or `"lambda.1se"`.

The lines below are a starting point so the file runs. Replace them
with the choice you just made.

```{r}
#| label: commit
chosen_alpha <- 1
chosen_s <- "lambda.1se"
chosen_fit <- if (chosen_alpha == 1) cv_lasso else cv_ridge

c(alpha = chosen_alpha, rule = chosen_s)
chosen_fit[[chosen_s]]
cv_rmse(chosen_fit, chosen_s)
n_sites(chosen_fit, chosen_s)
sqrt(mean((y - mean(y))^2))

# Interpret: Put your cross-validated RMSE next to 1.67, the RMSE
# from predicting the mean of log(spend), and next to the other three
# fits in the table above. The site count is part of the choice.
#
# Question: Suppose a classmate now tries the other three fits and
# keeps the one with the smallest cross-validated RMSE. What happens
# to the claim that your number is the score of a choice you made
# before looking?
#
# Business implication: One cross-validated error is what you can
# take as the score of the fit you chose. A search over the four
# numbers, after you have seen all four, is still a search.
```

# After class: elastic net

Work this after you have committed. Do not change `chosen_alpha` or
`chosen_s` because this fit looks a little different.

Elastic net uses `alpha = 0.5`, between ridge and the lasso. It was
the last question in an older lab on these files. It is not on the
lecture slides.

```{r}
#| label: after-class-enet
set.seed(369)
cv_enet <- cv.glmnet(X, y, alpha = 0.5, nfolds = 10, standardize = TRUE)

data.frame(
  rule = c("lambda.min", "lambda.1se"),
  lambda = c(cv_enet$lambda.min, cv_enet$lambda.1se),
  cv_rmse = c(cv_rmse(cv_enet, "lambda.min"), cv_rmse(cv_enet, "lambda.1se")),
  sites = c(n_sites(cv_enet, "lambda.min"), n_sites(cv_enet, "lambda.1se"))
)

# Question: How does the site count at lambda.min, and at lambda.1se,
# sit between the ridge fit and the lasso fit you already ran?
#
# Question: Would you replace the choice you committed to in class
# with one of these two? Why not?
```
