---
title: "Linear regression"
subtitle: "70-467 · Machine Learning for Business Analytics"
author: "Andrés Castaño Zuluaga"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: right
    body-classes: "mlba-page"
    code-fold: false
    code-tools: true
    html-math-method: katex
execute:
  echo: true
  warning: false
  message: false
  dir: file
fig-width: 8.2
fig-height: 4.2
---

```{r}
#| label: setup
#| include: false
library(readr)
library(ggplot2)

qmd_dir <- tryCatch(
  dirname(normalizePath(knitr::current_input(dir = TRUE))),
  error = function(e) getwd()
)
stores_path <- file.path(qmd_dir, "data", "stores.csv")
if (!file.exists(stores_path)) {
  for (p in c(
    "data/stores.csv",
    file.path(getwd(), "data", "stores.csv"),
    file.path(getwd(), "notes", "data", "stores.csv"),
    file.path(getwd(), "site", "notes", "data", "stores.csv")
  )) {
    if (file.exists(p)) {
      stores_path <- normalizePath(p)
      break
    }
  }
}
if (!file.exists(stores_path)) {
  stop(
    "Cannot find data/stores.csv. File > Open Project on site/notes/notes.Rproj, then Render."
  )
}

stores <- read_csv(stores_path, show_col_types = FALSE)
stores$ShelveLoc <- factor(stores$ShelveLoc, levels = c("Bad", "Medium", "Good"))
stores$Urban <- factor(stores$Urban, levels = c("No", "Yes"))
stores$US <- factor(stores$US, levels = c("No", "Yes"))

m1 <- lm(Sales ~ Price, data = stores)
m_ads <- lm(Sales ~ Advertising, data = stores)
m2 <- lm(Sales ~ Price + CompPrice + Income, data = stores)
m3 <- lm(Sales ~ Price + CompPrice + Income + Advertising, data = stores)
m4 <- lm(Sales ~ Price + CompPrice + Income + Advertising + ShelveLoc, data = stores)

coef_tab <- function(m) {
  s <- coef(summary(m))
  term <- rownames(s)
  term <- gsub('relevel\\(Urban, ref = "Yes"\\)', "Urban", term)
  term <- gsub('relevel\\(ShelveLoc, ref = "Good"\\)', "ShelveLoc", term)
  data.frame(
    Term = term,
    Estimate = round(s[, "Estimate"], 4),
    `Std. Error` = round(s[, "Std. Error"], 4),
    t = round(s[, "t value"], 2),
    `p-value` = ifelse(
      s[, "Pr(>|t|)"] < 0.001,
      "< 0.001",
      format.pval(s[, "Pr(>|t|)"], digits = 3, eps = 0.001)
    ),
    check.names = FALSE,
    row.names = NULL
  )
}

theme_mlba <- function() {
  theme_minimal(base_family = "sans", base_size = 13) +
    theme(
      plot.background = element_rect(fill = "#F4F6F8", color = NA),
      panel.background = element_rect(fill = "#F4F6F8", color = NA),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = "#E0E0E0", linewidth = 0.3),
      axis.title = element_text(color = "#6D6E71", size = 11),
      axis.text = element_text(color = "#182C4B", size = 11),
      legend.position = "none"
    )
}
```

## Introduction to Linear Regression

You have already seen linear regression. In Probability and Statistics, and in Regression Analysis, it was a statistical model that allows you to study the relationship between variables and to make inference about the strength of relationships.

We start this course in the same place. Many decisions in a firm — a list price, an advertising budget, a sales number for next quarter — depend on how an outcome lines up with things we can observe. Linear regression is the first tool we will use for that. The slope has units. The leftover has units. You can say what a \$10 gap in list price is associated with, in seats. Later methods in the course are more flexible. They are also easier to treat as a box that only produces a guess.

Machine learning puts that same tool on a map. The first cut is whether we have an outcome we are trying to match.

**Supervised learning.** Each row has $Y$ and $X$. The method encodes how $Y$ lines up with $X$ where $Y$ is known. Later we can produce a number, or a label, for a new row. Linear regression, logistic regression, trees, and forests in this course are all supervised.

**Unsupervised learning.** We have $X$ and no $Y$. Clustering, later in the term, is of this kind: find structure among customers or stores without a pre-specified outcome.

Within supervised learning there is a second split, on the type of $Y$:

**Regression.** $Y$ is quantitative — sales, a price, a return, a wait time.  
**Classification.** $Y$ is a category — buy / not buy, default / perform, churn / stay.

Linear regression is supervised, with a quantitative $Y$. That still does not tell you **what the fit is for**.

In most statistics and machine learning courses, regression is a way to produce $\hat Y$ for a new row. In applied social science and in business, the same fit is often used to read a coefficient: how sales line up with price, with advertising, with a policy. Same `lm`. Different question.

The book organizes this as **prediction** versus **inference** (James, Witten, Hastie, and Tibshirani 2021, pp. 15–22). Inference there means reading how $Y$ lines up with $X$. In this class I will use a slightly finer split — descriptive, associational, causal, and predictive (Shmueli 2010). The same line can do more than one of these jobs. Which one is appropriate depends on the problem in front of you, and on how the data were generated. Please make sure you can say which job you are doing before you interpret a coefficient.

1. **Descriptive.** How the variables in the file move together. A scatter, a slope, a correlation. No claim that we have found $f$ in the world.
2. **Explanatory (associational).** The association of $Y$ with $X_j$ after the other included variables are held fixed. That is what people mean by “controlling for.” It is the *ceteris paribus* language you have used in economics. It is not, by itself, the effect of *changing* $X_j$.
3. **Explanatory (causal).** The coefficient on a treatment is the effect of changing that treatment only if assignment is randomized, or as-if randomized — a credible natural experiment, a discontinuity, an instrument that shifts the treatment and nothing else (Holland 1986; Angrist and Pischke 2009, ch. 2). A file of stores that already chose their own prices and ads does **not** give you that.
4. **Predictive.** $\hat Y$ for a store we have not yet seen, or for next quarter. Accuracy of the guess — leftover size — matters more than a causal story.

A good predictive model can be a poor guide to a price change. A randomized trial can identify an effect and still be a weak forecast if the trial sample is unlike the stores you will operate next quarter. They are different questions.

Which job is the right one depends on the problem in front of you. We will do this with one firm and one file.

## A firm, 400 stores, several decisions

Think of a firm that sells child car seats through 400 stores. The seat is the same product. The store is not. In some locations the seat sits at eye level. In others it is on a bad shelf, next to a cheaper rival. Parents see a sticker price. They also see whatever the store spent on local ads.

Sales are all over the place. Two stores with the same Bad shelf can differ by more than a factor of two. Headquarters has to lock list price, local advertising, and shelf quality for next quarter. They also have to live with a competitor who already posted a price, and with neighborhoods that differ in income and in whether they are urban. Nobody assigned prices at random this quarter. They have a file.

That file can answer more than one question. If the question is how sales line up with price and ads in these rows, that is **descriptive**. If the question is how our list price is associated with sales, holding the competitor still, that is **associational**. It is not the effect of a cut. If the question is what happens if we *change* the price, this file does not answer it. If finance needs a sales number for next quarter for a store like these, that is **prediction**, and the leftover is the risk around that number.

Look at two stores before any model. Both have a **Bad** shelf.

```{r}
#| label: two-stores
#| echo: false
knitr::kable(
  stores[c(1, 5), c("Sales", "Price", "CompPrice", "Advertising", "ShelveLoc", "US")],
  row.names = TRUE
)
```

Store 1: 9.50 thousand units at \$120, competitor at \$138, \$11 thousand in local ads, in the US.  
Store 5: 4.15 thousand units at \$128, competitor at \$141, \$3 thousand in ads, not in the US.

Same shelf quality. More than twice the sales. Price, advertising, geography, and the competitor are all in play. If you have taken marketing, this is price, promotion, and place. If you have taken micro, this is own price, rival price, and demand shifters. If you have taken finance, next quarter’s sales number is an input to a revenue forecast, and the leftover around that number is risk.

## What is in the file

400 stores, 11 columns. Sales in thousands of units. Price and CompPrice in dollars. Advertising in thousands of dollars.

```{r}
#| label: load
stores <- read_csv("data/stores.csv", show_col_types = FALSE)
stores$ShelveLoc <- factor(stores$ShelveLoc, levels = c("Bad", "Medium", "Good"))
stores$Urban <- factor(stores$Urban, levels = c("No", "Yes"))
dim(stores)
head(stores[, c("Sales", "Price", "CompPrice", "Advertising", "ShelveLoc", "Income", "Urban", "US")])
```

| Column | What it is | Where you have seen this kind of variable |
|--------|------------|-------------------------------------------|
| `Sales` | Units sold (thousands) | Outcome. Revenue if you multiply by price. |
| `Price` | Our list price (dollars) | Pricing; own-price in a demand curve |
| `CompPrice` | Competitor’s price at that location | A substitute’s price |
| `Advertising` | Local advertising (thousands of dollars) | Promotion |
| `ShelveLoc` | Bad / Medium / Good | Place / merchandising, not the seat itself |
| `Income`, `Age`, `Population`, `Education` | Neighborhood | Demand shifters; targeting |
| `Urban`, `US` | Yes / No | Geography |

What headquarters can still set: `Price`, `Advertising`, `ShelveLoc`. What they do not set this quarter: `CompPrice`, income in the neighborhood, whether the store is urban or in the US.

## Questions, in order: describe, explain, predict

I want you to ask questions of this file in that order. The method is the same linear regression. The job changes.

**Descriptive**

- How do sales line up with our price, with local ads, with shelf quality, in the raw file?
- Is the own-price association negative? Is the competitor-price association positive?
- How dispersed are sales? That dispersion is what a forecast has to live with.

**Explanatory / associational**

- Holding the competitor’s price and the local market still, how is **our** list price associated with sales? That is the \$10 conversation. It is still not an experiment.
- Do local ads move sales once price and the shelf are in the picture? Does a Good shelf still show a gap after we hold price and ads fixed?
- Causal, only if we had it: if we had randomized list price across stores — or a rule that assigned cuts as-if randomly — the price coefficient could be read as the effect of the cut (Holland 1986; Angrist and Pischke 2009). We do not have that design in this file.

**Predictive**

- For a store with a given price, competitor, ads, and shelf, what sales do we guess for next quarter, and how large is a typical miss?
- You have to assume that next quarter looks like these 400 rows. If headquarters changes price **and** the world responds, that assumption is strained. We will be stricter about scoring a forecast later in the term. This week we stay in-sample and honest about it.

Four questions, rewritten in that order:

1. *Describe.* Stores 1 and 5, both Bad shelf: 9.50 vs 4.15. Which other columns differ?
2. *Explain (association).* Holding the competitor and the market still, what is the association of our price with sales?
3. *Explain (association).* Do ads and the shelf still show up after price is in the model?
4. *Predict.* What sales would you guess for a store like store 1 next quarter, and how far off should you expect to be?

A regression can speak to all four. We start with **one** predictor — our price — because the \$10 cut is a pricing question. We will not stop there. Advertising is a promotion question. It gets a short pass of its own, then it comes back once we have more than one $X$.

## A line for sales and our price

We start with our price because the \$10 cut is a pricing question.

A quantitative outcome is

$$
Y = f(X) + \varepsilon.
$$

$Y$ is sales. $X$ is our price, for this section. $\varepsilon$ is everything $f$ misses: ads, shelf, competitor, luck. We estimate $f$ to **predict** and to **read** an association. Linear regression restricts $f$ so that the coefficient has units (thousands of seats per dollar).

With one predictor, the **population** line is

$$
Y = \beta_0 + \beta_1 X + \varepsilon.
$$

- $\beta_1$: average change in sales associated with a one-dollar higher list price.
- $\beta_0$: $\mathrm{E}(Y \mid X = 0)$. Fitted sales at a price of zero. That is not a store.
- $\varepsilon$: leftover.

We observe 400 pairs and compute $\hat\beta_0$, $\hat\beta_1$:

$$
\hat y_i = \hat\beta_0 + \hat\beta_1 x_i, \qquad e_i = y_i - \hat y_i.
$$

$e_i$ is how far store $i$ sits off the line, in thousands of units. If the job is descriptive, $\hat\beta_1$ is how sales and price line up. If the job is associational, it is the association of our price with sales — still not a causal cut. If the job is prediction, $\hat y_i$ is the guess and $e_i$ is the miss.

Does a line look reasonable here?

::: {.panel-tabset}

#### The cloud

```{r}
#| label: scatter
ggplot(stores, aes(Price, Sales)) +
  geom_point(alpha = 0.55, size = 2, color = "#182C4B") +
  labs(x = "Price (dollars)", y = "Sales (thousands of units)") +
  theme_mlba()
```

#### The line

```{r}
#| label: line
ggplot(stores, aes(Price, Sales)) +
  geom_point(alpha = 0.5, size = 1.8, color = "#182C4B") +
  geom_smooth(method = "lm", se = FALSE, color = "#C41230", linewidth = 1.1) +
  labs(x = "Price (dollars)", y = "Sales (thousands of units)") +
  theme_mlba()
```

:::

Negative slope: higher list price, lower sales, on average. That is the sign you expect from a demand curve — **as a description**. The cloud is wide. A store at \$120 can sell 4 thousand units or 12. Price is not the whole story. Correlation of our price with the competitor’s price is `r sprintf("%.2f", cor(stores$Price, stores$CompPrice))`. If you read **our** price off this cloud alone, you are mixing our list price with the market the store sits in.

::: {.callout-tip collapse="true"}
### Check — which job?

“In these 400 stores, higher own price goes with lower sales.” Descriptive, explanatory, or predictive? Causal?

<details>
<summary>Answer</summary>

Descriptive, and a crude associational claim if you stop at one $X$. Not predictive: you have not produced $\hat Y$ for a new store. Not causal: prices were not randomized.
</details>
:::

## Least squares

We need a rule to pick the line. Least squares picks $\hat\beta_0$ and $\hat\beta_1$ to make the **vertical** misses small:

$$
\mathrm{RSS} = \sum_{i=1}^{n} \bigl( y_i - \hat\beta_0 - \hat\beta_1 x_i \bigr)^2 = \sum_{i=1}^{n} e_i^2.
$$

Why vertical? Because $Y$ is the outcome we care about — sales — and $e_i$ is a miss in units of sales. Why square? Large misses count more than small ones, and the calculus is clean. The minimizers have a closed form. We compute them rather than derive them here.

If the job is descriptive, the line is the linear summary of the cloud that makes the squared vertical gaps as small as they can be. If the job is associational, $\hat\beta_1$ is the slope of that summary. It is still the slope that best fits *this* cloud, not the slope from an experiment. If the job is prediction, RSS — and RSE, below — is the in-sample size of the forecast errors. Someone asking how many seats we are off by wants that leftover, not a story about $\beta_1$ alone.

```{r}
#| label: fit-m1
m1 <- lm(Sales ~ Price, data = stores)
coef(m1)
sum(resid(m1)^2)
```

`Sales ~ Price` means sales as a function of our price, plus an intercept. RSS = `r format(round(sum(resid(m1)^2), 1), big.mark = ",")`. Residual degrees of freedom: $n-2 = 398$ (two numbers were estimated: intercept and slope).

## Reading the slope as a pricing number

```{r}
#| label: summary-m1
#| echo: false
knitr::kable(coef_tab(m1), row.names = FALSE, align = "lrrrr")
```

```{r}
#| label: summary-print
summary(m1)$coefficients
```

**Estimate, intercept.** Fitted sales if price were \$0. There is no such store.

**Estimate, Price.** A one-dollar higher list price is associated with `r sprintf("%.3f", coef(m1)["Price"])` thousand fewer units — about `r abs(round(coef(m1)["Price"] * 1000))` seats — **when we use only price**. Say “associated with,” not “a one-dollar cut increases sales.” We did not cut anything.

For the \$10 conversation (associational, still not causal):

$$
10 \times (`r sprintf("%.3f", coef(m1)["Price"])`) = `r sprintf("%.2f", 10 * coef(m1)["Price"])`
$$

thousand units, about `r abs(round(10 * coef(m1)["Price"] * 1000))` seats. That is a crude own-price association. It is not an elasticity until we divide by levels, and it is not a demand curve until price is as-if random. It is also a **quantity** response. Revenue is price times quantity, and both would move.

**Std. Error.** How wiggly $\hat\beta_1$ is across samples of 400 stores like these, in the same units as the slope.

**$t$ and $p$.** $t = \hat\beta / \mathrm{SE}(\hat\beta)$. $H_0{:}\ \beta_1 = 0$ says there is no linear association. We reject it here ($t \approx `r sprintf("%.1f", coef(summary(m1))["Price","t value"])`$). That answers a descriptive question: the cloud is not flat. It does not answer “should we cut.”

::: {.callout-warning}
**Intercept:** not a store.  
**Small $p$:** not a price change.  
**\$10 $\times$ slope:** association in these rows, not the effect of an experimental cut.
:::

## How sure is the slope?

If the errors are uncorrelated with constant variance $\sigma^2$,

$$
\mathrm{SE}(\hat\beta_1)^2 = \frac{\sigma^2}{\sum_{i=1}^{n}(x_i - \bar x)^2}.
$$

More spread in price (the denominator) gives a more precise slope. We learn more when stores actually posted different prices. $\sigma$ is unknown; we estimate it from the residuals. For Price, $\mathrm{SE}(\hat\beta_1) = `r sprintf("%.4f", coef(summary(m1))["Price","Std. Error"])`$.

An approximate 95% interval is $\hat\beta_1 \pm 2\,\mathrm{SE}(\hat\beta_1)$:

$$
[`r sprintf("%.3f", coef(m1)["Price"] - 2 * coef(summary(m1))["Price","Std. Error"])`,\
 `r sprintf("%.3f", coef(m1)["Price"] + 2 * coef(summary(m1))["Price","Std. Error"])`].
$$

It does not contain zero. Same information as the $t$.

If the job is descriptive or associational, the interval is a range for the slope in this specification. It is not a range for the effect of a cut. If the job is prediction, uncertainty about $\beta_1$ is only part of forecast uncertainty. Even if we knew $\beta_1$, $\varepsilon$ would remain. That is the next section.

## Is ten dollars large?

A star tells you the cloud is not flat. The decision in front of the firm is a different question: is a \$10 list-price gap large enough to matter **in this file**? McCloskey and Ziliak (1996) is the economics version of that distinction. Statistical significance is not the same as a coefficient that is large in the units of the decision. We will not debate their survey of journals. We will do the arithmetic on these 400 rows.

```{r}
#| label: ten-dollars
b10 <- 10 * coef(m1)["Price"]
b10
mean(stores$Sales)
sigma(m1)
b10 / mean(stores$Sales)
b10 / sigma(m1)
```

If the question is seats, a \$10 gap is associated with about `r sprintf("%.2f", 10 * coef(m1)["Price"])` thousand units — on the order of `r abs(round(10 * coef(m1)["Price"] * 1000))` seats — when we use only price. Mean sales are `r sprintf("%.2f", mean(stores$Sales))` thousand units, so that is about `r sprintf("%.0f", 100 * abs(10 * coef(m1)["Price"]) / mean(stores$Sales))`% of typical sales in the file.

If the question is leftover, RSE is `r sprintf("%.2f", sigma(m1))` thousand units. The \$10 association is smaller than a typical miss. A store can easily be off by more seats than this pricing gap “explains.”

If the question is profit, this file does not answer it. There are no costs or margins. Prices were not randomized, so this is not the effect of *cutting* \$10. And `10 * coef(m1)["Price"]` is not a revenue number: revenue is price $\times$ quantity, and both would move.

A slope can be precisely estimated and still be modest next to leftover (McCloskey and Ziliak 1996). We will run the same arithmetic once the competitor, advertising, and the shelf are in the model. The \$10 association will change, because the object we are calling “the price coefficient” will have changed.

## How far off is the guess?

Question 4 was prediction: a sales number for next quarter, for a store like these.

Even with $\beta_0$ and $\beta_1$ known, $\varepsilon$ is still there. The **residual standard error** estimates that leftover **in thousands of units**:

$$
\mathrm{RSE} = \sqrt{\frac{\mathrm{RSS}}{n-2}}.
$$

```{r}
#| label: rse
sigma(m1)
```

RSE $= `r sprintf("%.2f", sigma(m1))`$ thousand units. Mean sales are `r sprintf("%.2f", mean(stores$Sales))`. A typical miss is large next to the \$10 quantity association above. If this line is used as a revenue input, that leftover has to be on the slide. RMSE divides by $n$ instead of $n-2$; we report **RSE**.

$R^2$ is the descriptive companion: what share of the variation in sales does the price line catch, relative to guessing the mean every time?

$$
R^2 = 1 - \frac{\mathrm{RSS}}{\mathrm{TSS}}, \qquad
\mathrm{TSS} = \sum_{i=1}^{n}(y_i - \bar y)^2.
$$

```{r}
#| label: r2
summary(m1)$r.squared
```

$R^2 = `r sprintf("%.3f", summary(m1)$r.squared)`$. About one fifth. In data of this kind, that is ordinary. It is not a failed exam. It is also not a reason to ignore ads and the shelf, and it is not the number to quote if someone asks how many seats we are off by.

If the question is “how far off is the forecast, in units?”, use RSE. If the question is “how much of the scatter did price catch?”, use $R^2$. They are different questions.

::: {.callout-tip collapse="true"}
### Check

(a) “How far off is the forecast, in units?” (b) “How much of the scatter did price catch?” RSE or $R^2$ for each?

<details>
<summary>Answer</summary>

(a) RSE — units of sales. (b) $R^2$ — a share, no units.
</details>
:::

## A short ads line

Price was a pricing problem. Local advertising is a promotion problem. Before we put several $X$s in one model, look at the raw association of sales with ads. Same jobs, shorter than the price section.

Advertising in the file is already in **thousands of dollars**. One unit is \$1,000 of local spend. Mean spend is about \$`r sprintf("%.0f", 1000 * mean(stores$Advertising))`. `r sprintf("%.0f", 100 * mean(stores$Advertising == 0))`% of stores spend nothing.

```{r}
#| label: fit-ads
m_ads <- lm(Sales ~ Advertising, data = stores)
summary(m_ads)$coefficients
sigma(m_ads)
summary(m_ads)$r.squared
coef(m_ads)["Advertising"] / mean(stores$Sales)
coef(m_ads)["Advertising"] / sigma(m_ads)
```

An extra \$1,000 of local ads is associated with `r sprintf("%.3f", coef(m_ads)["Advertising"])` thousand extra units — on the order of `r round(coef(m_ads)["Advertising"] * 1000)` seats — **when we use only ads**.

If the question is seats, that is about `r sprintf("%.0f", 100 * coef(m_ads)["Advertising"] / mean(stores$Sales))`% of typical sales. If the question is leftover, RSE is `r sprintf("%.2f", sigma(m_ads))`; the ads association is small next to a typical miss. $R^2$ is `r sprintf("%.3f", summary(m_ads)$r.squared)` — ads catch less of the scatter than price did.

If the question is a return on spend, this file does not answer it. The column *is* the spend. What is missing is margin, and whether the extra seats would still be there if we *raised* ads. Stores that advertise more may already be the stores that sell more for other reasons.

Ads and sales move together. The association is not huge next to leftover. We have not held price or the shelf still. We will hold them still below. Correlation of ads with our price is `r sprintf("%.2f", cor(stores$Advertising, stores$Price))` — unlike Price and CompPrice, these two are almost uncorrelated in the file. That is useful. It is not a reason to stop at one $X$.

## Why one $X$ is the wrong specification

The pricing question we actually care about is **our** list price, holding the market still. Price and CompPrice move together (`r sprintf("%.2f", cor(stores$Price, stores$CompPrice))`). A store that charges more often sits next to a competitor that charges more. Leave CompPrice out, and part of $\hat\beta_{\mathrm{Price}}$ is the local market.

That is an omitted-variable problem you have seen in econometrics: the coefficient on the included variable absorbs the excluded one when the two move together. The book’s version uses three advertising media: newspaper looks associated with sales on its own, then dies once radio is in the model, because newspaper was standing in for radio (James et al. 2021, pp. 73–75). Here the same logic is Price and CompPrice. The ice-cream / shark-attacks story is the same point with the temperature left out. Here the “temperature” is the local price level.

A second problem with three separate simple regressions — one for price, one for ads, one for the competitor — is that you then have three equations and no single sales guess for a store that has all three numbers. You want one specification, with each coefficient read **holding the others in the equation fixed**.

Advertising and the shelf are still sitting in $\varepsilon$ of `m1`. Question 3 was: do ads move sales *once price and the shelf are in the picture*? That sentence is multiple regression.

## One model, several $X$s

Give each predictor its own slope in a single equation:

$$
Y = \beta_0 + \beta_1 X_1 + \cdots + \beta_p X_p + \varepsilon,
$$

or

$$
\mathrm{E}(Y \mid X_1,\ldots,X_p) = \beta_0 + \beta_1 X_1 + \cdots + \beta_p X_p.
$$

$\beta_j$ is the average change in $Y$ associated with a one-unit change in $X_j$, **holding the other included predictors fixed**. Least squares still minimizes RSS. The geometry is a plane (or a hyperplane) instead of a line. The algebra is the same idea; software computes $\hat\beta$. We do not need the matrix formula this week.

```{r}
#| label: fit-m2
m2 <- lm(Sales ~ Price + CompPrice + Income, data = stores)
knitr::kable(coef_tab(m2), align = "lrrrr", row.names = FALSE)
```

$R^2 = `r sprintf("%.3f", summary(m2)$r.squared)`$. RSE $= `r sprintf("%.2f", sigma(m2))`$.

The coefficient we care about for pricing is Price, now holding CompPrice and Income fixed. A one-dollar higher own price is associated with `r sprintf("%.3f", coef(m2)["Price"])` thousand fewer units.

CompPrice, holding our price and Income fixed: a one-dollar higher competitor price is associated with `r sprintf("%.3f", coef(m2)["CompPrice"])` thousand *extra* units. That is the substitute sign you want as a description: when the rival is more expensive, our sales are higher, own price held still.

Income is a neighborhood shifter. Not something we set this quarter.

## The Price coefficient is not the same object

```{r}
#| label: nested-price
#| echo: false
nested <- data.frame(
  Model = c("Sales ~ Price", "+ CompPrice + Income"),
  Price = sprintf("%.3f", c(coef(m1)["Price"], coef(m2)["Price"])),
  R2 = sprintf("%.3f", c(summary(m1)$r.squared, summary(m2)$r.squared)),
  RSE = sprintf("%.2f", c(sigma(m1), sigma(m2)))
)
knitr::kable(nested, align = "lrrr")
```

After CompPrice and Income are included, $|\hat\beta_{\mathrm{Price}}|$ is **larger** (`r sprintf("%.3f", coef(m1)["Price"])` versus `r sprintf("%.3f", coef(m2)["Price"])`). A one-dollar gap in our list price is associated with a bigger sales gap once we stop mixing our price with the local market.

```{r}
#| label: ten-dollars-controlled
b10_c <- 10 * coef(m2)["Price"]
b10_c
b10_c / mean(stores$Sales)
b10_c / sigma(m2)
```

A \$10 own-price gap is now about `r sprintf("%.2f", 10 * coef(m2)["Price"])` thousand units — on the order of `r abs(round(10 * coef(m2)["Price"] * 1000))` seats — holding competitor price and income fixed. Versus mean sales that is about `r sprintf("%.0f", 100 * abs(10 * coef(m2)["Price"]) / mean(stores$Sales))`%. Versus this model’s leftover (`r sprintf("%.2f", sigma(m2))`) it is still smaller than a typical miss, but it is a larger pricing number than the simple regression gave you.

Same limits as before: no experiment, no margins. What changed is the **object**. “The price coefficient” is a function of what else is in the equation.

## Advertising, holding price still

Now the promotion variable, holding price and the market still.

```{r}
#| label: fit-m3
m3 <- lm(Sales ~ Price + CompPrice + Income + Advertising, data = stores)
knitr::kable(coef_tab(m3), align = "lrrrr", row.names = FALSE)
```

The coefficient we care about for advertising is `r sprintf("%.3f", coef(m3)["Advertising"])`. Holding price, competitor price, and income fixed, an additional \$1,000 of local advertising is associated with `r sprintf("%.3f", coef(m3)["Advertising"])` thousand extra units — about `r round(coef(m3)["Advertising"] * 1000)` seats.

```{r}
#| label: ads-controlled
b_ads <- coef(m3)["Advertising"]
b_ads
b_ads / mean(stores$Sales)
b_ads / sigma(m3)
10 * coef(m3)["Price"]
```

If the question is seats, \$1,000 of local ads is about `r sprintf("%.2f", coef(m3)["Advertising"])` thousand extra units, holding the other three still — on the order of `r sprintf("%.0f", 100 * coef(m3)["Advertising"] / mean(stores$Sales))`% of typical sales. If the question is leftover, RSE is now `r sprintf("%.2f", sigma(m3))`; the ads move is still small next to a typical miss.

If the question is how ads compare with price **in the same specification**, a \$10 own-price gap is about `r sprintf("%.2f", 10 * coef(m3)["Price"])` thousand units. In this file, a \$10 list-price difference is a larger sales association than a \$1,000 ads difference. That is not “ignore advertising.” It is a ranking of associations in these rows, with no costs attached.

$R^2$ moves from `r sprintf("%.3f", summary(m2)$r.squared)` to `r sprintf("%.3f", summary(m3)$r.squared)`. The Price coefficient stays about `r sprintf("%.3f", coef(m3)["Price"])`. Ads were not hiding inside the price slope the way CompPrice was — we already saw that ads and price are almost uncorrelated.

If the question is a return on spend, there is no margin in the file. The coefficient is extra seats associated with stores that spent another \$1,000, holding the included columns fixed. It is not the effect of *raising* the ad budget. The shelf is still in $\varepsilon$.

## Is anyone in this model doing work?

Once several predictors are in the model, you can ask a different question. Is anyone in this specification associated with sales, or are we looking at noise?

The overall $F$ test asks that. For a model with $p$ slopes, the null is that every slope is zero:

$$
H_0{:}\ \beta_1 = \beta_2 = \cdots = \beta_p = 0
$$

against the alternative that at least one $\beta_j$ is not zero. The statistic is

$$
F = \frac{(\mathrm{TSS}-\mathrm{RSS})/p}{\mathrm{RSS}/(n-p-1)}.
$$

If $H_0$ is true and the usual conditions hold, $F$ is near 1. Software reports a $p$-value from an $F$ distribution.

That is not the same question as “is the $t$ on Price large?” If you look at many $t$ statistics and keep the stars, some of those stars will be accidents. The book has a simple version of this: one hundred noise predictors, and you should expect about five $p$-values below 0.05 by chance (James et al. 2021, pp. 77–78). The $F$ is one test for the whole specification. It pays a price for the number of predictors.

```{r}
#| label: overall-f
summary(m3)$fstatistic
```

In this file $n = 400$ and $p$ is small, so $F$ will be large. Rejecting the null is not a pricing recommendation. It is not an advertising recommendation. It tells you the surface is not flat. Please make sure you can say the null in words, not only in Greek.

Three tests, three questions:

- one $t$ for Price — does *this* coefficient show up, others held fixed;
- the overall $F$ — does *anyone* in the specification show up;
- a nested $F$, next — do *these extra columns* show up, given the ones already in.

## Does this extra block belong?

A middle question is whether a *block* of extra columns belongs, given the ones already in. Not “all slopes zero,” and not “this one $t$.”

Fit a smaller model (RSS$_0$) and a larger one that contains it (RSS). If the extra block has $q$ coefficients,

$$
F = \frac{(\mathrm{RSS}_0 - \mathrm{RSS})/q}{\mathrm{RSS}/(n-p-1)}.
$$

A single-variable $t$ is this test with $q = 1$. For the shelf, $q = 2$ (two dummies), so you cannot read “does shelf belong?” off one dummy’s $t$.

In R the comparison is `anova` of the small model, then the large one.

```{r}
#| label: nested-f
anova(m1, m2)
```

CompPrice and Income, added to Price, drop RSS by enough that the nested $F$ is large. If you want a coefficient on our price that holds the local market still, the competitor has to be in the equation.

The book then names a further task: choose predictors among $2^p$ subsets, with $C_p$, AIC, BIC, adjusted $R^2$, and forward, backward, or mixed selection (James et al. 2021, pp. 78–79). That is Chapter 6, and a later week here. We will not run stepwise on these 400 rows. We add variables that match the questions above: price, then the competitor, then advertising, then the shelf.

What we do take from that discussion this week is adjusted $R^2$. Raw $R^2$ cannot fall when you add a column. That is mechanical. It is not evidence that the new column belongs. Adjusted $R^2$ penalizes the extra parameters.

## $R^2$ always goes up

In multiple regression, $R^2 = \mathrm{Cor}(Y,\hat Y)^2$. Add a predictor and RSS on these 400 rows falls, or stays. So $R^2$ rises, or stays.

RSE is

$$
\mathrm{RSE} = \sqrt{\frac{\mathrm{RSS}}{n-p-1}}.
$$

The denominator shrinks when $p$ rises. If RSS barely falls, leftover can get *worse*. The book shows that with an extra media variable that barely moves the fit (pp. 80–81). Watch for the same pattern in *our* nested table.

```{r}
#| label: nested-fit
#| echo: false
fit_row <- function(m, label) {
  data.frame(
    Model = label,
    Price = sprintf("%.3f", unname(coef(m)["Price"])),
    R2 = sprintf("%.3f", summary(m)$r.squared),
    `Adj. R2` = sprintf("%.3f", summary(m)$adj.r.squared),
    RSE = sprintf("%.2f", sigma(m)),
    check.names = FALSE
  )
}
knitr::kable(
  rbind(
    fit_row(m1, "Price"),
    fit_row(m2, "+ CompPrice + Income"),
    fit_row(m3, "+ Advertising")
  ),
  align = "lrrrr"
)
```

Each added block in this sequence raises $R^2$ and, in this file, also raises adjusted $R^2$ and cuts RSE. That is *this* sequence, not a law. Two questions when a column is added: did the coefficient you care about **change**, and did leftover fall enough to matter for the guess? If the coefficient does not move and leftover barely moves, a tiny $R^2$ bump is not a reason to lengthen the specification.

## A guess for next quarter, and two kinds of interval

Once the surface is fitted, $\hat y = \hat\beta_0 + \hat\beta_1 x_1 + \cdots + \hat\beta_p x_p$ is the guess for a store with those $x$ values. Two intervals sit around that number. They are centered at the same guess. They answer different questions (James et al. 2021, pp. 82–83).

If the question is typical sales for stores with this mix of price, competitor, income, and ads, that is a **confidence interval**. It is about the mean $f(x)$. Uncertainty here is that $\hat\beta$ is not $\beta$.

If the question is sales for *this* store next quarter, that is a **prediction interval**. Even if we knew $\beta$, $\varepsilon$ is still there. The prediction interval is always wider, because it has to live with that leftover.

There is a third issue: the linear form is itself an approximation. We name that and we do not chase it this week.

Mixing the two intervals is how a chain average and a store-level forecast talk past each other. Someone asking how far off we might be for store 1 wants the wide one.

```{r}
#| label: two-intervals
nd <- stores[1, ]
predict(m3, newdata = nd, interval = "confidence")
predict(m3, newdata = nd, interval = "prediction")
stores[1, c("Sales", "Price", "CompPrice", "Advertising", "ShelveLoc")]
```

Same center; the prediction interval is much wider. Store 1’s actual sales can sit outside the tight interval. That interval was never a promise about one store.

We will refit these two intervals after the shelf is in.

::: {.callout-tip collapse="true"}
### Check — which interval?

(a) “Typical sales for stores with this price, competitor, income, and ads.” (b) “Sales for store 1 next quarter.” Confidence or prediction?

<details>
<summary>Answer</summary>

(a) Confidence interval — mean at this mix. (b) Prediction interval — one store, $\varepsilon$ included.
</details>
:::

The book’s Figure 3.5 shows residuals from a two-predictor plane that fan out in a way that looks like **synergy** (two advertising media together versus either alone). That pattern is an **interaction**. It is the next section of the book (p. 87). We name it and stop. This week the leftover is one number and one histogram, not a hunt for products of $X$s.

## When the predictor is not a number

`Urban` is Yes or No. It is not a dollar amount. For two groups, the linear model uses **one dummy**. One group is the baseline (here, No).

$$
y_i =
\begin{cases}
\beta_0 + \beta_1 + \varepsilon_i & \text{if Urban is Yes}\\
\beta_0 + \varepsilon_i & \text{if Urban is No.}
\end{cases}
$$

$\beta_0$ is mean sales among non-urban stores. $\beta_1$ is the difference, urban versus not. Which group is coded 0 is a choice. The fitted values do not care. The sentence you read off the coefficient does.

```{r}
#| label: urban-dummy
m_u <- lm(Sales ~ Urban, data = stores)
knitr::kable(coef_tab(m_u), align = "lrrrr", row.names = FALSE)

m_u_flip <- lm(Sales ~ relevel(Urban, ref = "Yes"), data = stores)
knitr::kable(coef_tab(m_u_flip), align = "lrrrr", row.names = FALSE)

max(abs(fitted(m_u) - fitted(m_u_flip)))
```

With No as baseline, the intercept is mean sales in non-urban stores. The Yes coefficient is a small negative number, and the $t$ is near zero. In this file, urban versus not is not a sales story. I am showing it so you can see the coding on a small object, not because anyone should act on `Urban`.

Flip the baseline. The intercept becomes the urban mean, the coefficient flips sign, and every fitted value is the same. Please make sure you can say that in words.

Machine learning often calls this dummy coding **one-hot encoding**. Same object. For two levels you still leave one out; otherwise the dummies and the intercept are redundant.

The variable that actually moves sales in this file is the shelf.

## Three shelf locations

`ShelveLoc` takes three values: Bad, Medium, Good. A qualitative predictor with $k$ levels uses **$k-1$ dummy variables**. The omitted level is the baseline. We code Bad as the baseline.

$$
\begin{aligned}
\mathrm{E}(\mathrm{Sales}\mid \ldots)
&= \beta_0 + \beta_1 \mathrm{Price} + \cdots \\
&\quad + \beta_{\mathrm{Med}}\,\mathbf{1}\{\mathrm{Medium}\}
+ \beta_{\mathrm{Good}}\,\mathbf{1}\{\mathrm{Good}\}.
\end{aligned}
$$

- $\beta_{\mathrm{Good}}$: difference in mean sales, Good versus Bad, holding the other included predictors fixed
- $\beta_{\mathrm{Med}}$: the same comparison for Medium versus Bad

If you put all three dummies *and* an intercept in the model, the columns are linearly dependent. Software will drop one or refuse. Use $k-1$.

```{r}
#| label: fit-m4
m4 <- lm(Sales ~ Price + CompPrice + Income + Advertising + ShelveLoc,
         data = stores)
knitr::kable(coef_tab(m4), align = "lrrrr", row.names = FALSE)
```

$R^2 = `r sprintf("%.3f", summary(m4)$r.squared)`$. RSE $= `r sprintf("%.2f", sigma(m4))`$. Overall $F$ is still huge. That is “someone in this equation is doing work,” not a ranking of price versus ads versus shelf.

```{r}
#| label: shelf-block
anova(m3, m4)
```

The shelf is **two** coefficients. The nested $F$ asks whether both are zero, given price, competitor, income, and advertising. It is large. The shelf belongs in the specification as a block. You would miss that if you only stared at one dummy’s $t$.

Change the baseline. Fitted values should not move.

```{r}
#| label: relevel-shelf
m4b <- lm(
  Sales ~ Price + CompPrice + Income + Advertising +
    relevel(ShelveLoc, ref = "Good"),
  data = stores
)
knitr::kable(coef_tab(m4b), align = "lrrrr", row.names = FALSE)
max(abs(fitted(m4) - fitted(m4b)))
```

Good is now the omitted group. The intercept is a different number. Bad versus Good is about `r sprintf("%.2f", coef(m4b)["relevel(ShelveLoc, ref = \"Good\")Bad"])` thousand units — the same gap as Good versus Bad in `m4`, with the opposite sign. $\hat y$ matches.

Individual dummy $t$s **do** depend on which level you omitted. The $F$ for the factor does not. If you need “does shelf matter at all?”, use the nested $F$, not a shopping trip among dummy $p$-values.

## Price, advertising, and the shelf, in the same units

Gelman’s useful move here is not “standardize every column.” It is: put a continuous move and a dummy in the **same units of $Y$**, then read the comparison (Gelman 2008). We already have those units: thousands of seats.

```{r}
#| label: three-levers
b10_m4 <- 10 * coef(m4)["Price"]
b_ads_m4 <- coef(m4)["Advertising"]
b_good <- coef(m4)["ShelveLocGood"]
c(
  price_10 = unname(b10_m4),
  ads_1000 = unname(b_ads_m4),
  shelf_good_vs_bad = unname(b_good)
)
b10_m4 / sigma(m4)
b_ads_m4 / sigma(m4)
b_good / sigma(m4)
b10_m4 / b_good
b_ads_m4 / b_good
```

The coefficient we care about for pricing is still Price, now holding the other included columns fixed. A \$10 own-price gap is about `r sprintf("%.2f", 10 * coef(m4)["Price"])` thousand units. A \$1,000 advertising gap is about `r sprintf("%.2f", coef(m4)["Advertising"])` thousand units. Good versus Bad is about `r sprintf("%.2f", coef(m4)["ShelveLocGood"])` thousand units. Leftover is now `r sprintf("%.2f", sigma(m4))`.

If the question is which association is large in *seats* in this file, the shelf is the large one, and a \$10 price gap is larger than \$1,000 of ads. If the question is what we should *do*, we still do not have costs or margins, and we still do not have an experiment. Those are different questions. Cost of a nicer shelf is not in the file. Cost of ads is in the file as spend, but margin is not, so you still do not have a return.

Two stores that differ **only** in shelf (Good vs Bad), with the other included $X$ the same, differ in fitted sales by $\hat\beta_{\mathrm{Good}}$. That is what “holding the other predictors fixed” means.

```{r}
#| label: nested-final
#| echo: false
knitr::kable(
  rbind(
    fit_row(m1, "Price"),
    fit_row(m2, "+ CompPrice + Income"),
    fit_row(m3, "+ Advertising"),
    fit_row(m4, "+ ShelveLoc")
  ),
  align = "lrrrr"
)
```

When a predictor is added, two questions:

1. Does the coefficient you care about **change**?
2. Does leftover fall enough to matter for the guess? (RSE, not only $R^2$.)

$R^2$ always rises. Adjusted $R^2$ and the nested $F$ are the tools that do not automatically reward a longer equation.

Store 1 again, now with the shelf in the model. Same two intervals.

```{r}
#| label: two-intervals-m4
nd <- stores[1, ]
predict(m4, newdata = nd, interval = "confidence")
predict(m4, newdata = nd, interval = "prediction")
```

The guess moved because a Bad shelf is now in the equation explicitly. The prediction interval is still the one you want for *this* store.

## Residuals; what the estimates are not

$e_i = y_i - \hat y_i$. RSE $= `r sprintf("%.2f", sigma(m4))`$ thousand units (simple price line: `r sprintf("%.2f", sigma(m1))`).

```{r}
#| label: resid
#| fig-height: 3.15
#| fig-alt: Histogram of residuals from the multiple regression with shelf location.
ggplot(data.frame(e = resid(m4)), aes(e)) +
  geom_histogram(bins = 28, fill = "#182C4B", color = "#F4F6F8") +
  geom_vline(xintercept = 0, color = "#C41230", linewidth = 1) +
  labs(
    x = "Residual (thousands of units)",
    y = "Number of stores"
  ) +
  theme_mlba()
```

**Not an experiment.** Prices, ads, and shelves were not randomly assigned. $\hat\beta_{\mathrm{Price}}$ is not the effect of cutting the price. $\hat\beta_{\mathrm{Advertising}}$ is not the effect of raising the budget.

**Not a prediction for new stores.** Every $R^2$ and RSE here uses the same 400 rows used to fit. In-sample fit overstates accuracy out of sample. Week 3 is where we get honest about that.

**Not immune to collinearity.** Price and CompPrice are correlated. If they were nearly the same series, the individual coefficients would be unstable even if fitted values remained usable. We are not running a VIF table this week. The English is enough: the competitor is in the model because leaving it out mixed the local market into our price slope.

**Not an interaction model.** If the association of price with sales *depends on* ads, that is page 87. We stop.

Gelman and Stern (2006) put it this way: “The difference between ‘significant’ and ‘not significant’ is not itself statistically significant.” Do not treat “ads matter in `m3` and the shelf is the real story in `m4`” as a finding off a change in stars. Read the coefficients in units, and the nested $F$ if you need a test of a block.

## Readings

James, G., Witten, D., Hastie, T., and Tibshirani, R. (2021). *An Introduction to Statistical Learning*, 2nd edition. Free at [statlearning.com](https://www.statlearning.com/). **Assigned.**

- Chapter 1, pages 1–6 — supervised vs unsupervised; regression vs classification
- Chapter 2, pages 15–22 — $Y = f(X)+\varepsilon$; prediction vs inference
- Chapter 3, pages 59–70 — simple regression
- Chapter 3, pages 71–82 — multiple regression (the four questions)
- Chapter 3, pages 83–86 — qualitative predictors (stop before page 87)

If you want another pass on standard errors, confidence intervals, and $t$-statistics, pages 63–68 are the same material we used for the simple slope.

**Extras, not on the syllabus.** Cited above where we used them. You are not required to read them for Quiz 1.

- Shmueli, G. (2010). “To Explain or to Predict?” *Statistical Science* 25(3): 289–310. [Open PDF](https://arxiv.org/pdf/1101.0891).
- Holland, P. W. (1986). “Statistics and Causal Inference.” *JASA* 81(396): 945–960.
- Angrist, J. D., and Pischke, J.-S. (2009). *Mostly Harmless Econometrics*, chapter 2.
- McCloskey, D. N., and Ziliak, S. T. (1996). “The Standard Error of Regressions.” *Journal of Economic Literature* 34(1): 97–114.
- Gelman, A., and Stern, H. (2006). “The Difference Between ‘Significant’ and ‘Not Significant’ is not Itself Statistically Significant.” *The American Statistician* 60(4): 328–331. [PDF](https://sites.stat.columbia.edu/gelman/research/published/signif4.pdf).
- Gelman, A. (2008). “Scaling Regression Inputs by Dividing by Two Standard Deviations.” *Statistics in Medicine* 27: 2865–2873. We used the comparison of a continuous move to a dummy, not a requirement that you rescale the file. [PDF](https://sites.stat.columbia.edu/gelman/research/published/standardizing7.pdf).

## Monday and Wednesday

Monday: the same 400 stores — competitor, advertising, shelf, overall $F$, nested $F$, two kinds of interval — and Quiz 1 at the **end** of class. The quiz covers Wednesday and Monday.

Wednesday: first lab, a **different** file. After Monday I will share a short problem on that dataset and starter code. Please make sure you review those before you sit down, and please finish [Setup](../setup/index.qmd) so we do not spend the lab on installation.
