---
title: "Lab 3: Selecting a model for Balance on a QNB credit card"
subtitle: "70-467 · Monday 21 September · bring a laptop"
author: "Andrés Castaño Zuluaga"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: right
    body-classes: "mlba-page"
    code-fold: false
    code-tools: true
execute:
  echo: true
  warning: false
  message: false
---

<!--
Student starter for the hour. Code is filled on purpose. Run each chunk,
read the comments, and talk through the Interpret / Question / Discuss
prompts in the room. The 120 held-out accounts stay closed.
-->

# Introduction to the Business Problem

We have been using the Credit data in class. For this lab, treat the 400
rows as credit-card accounts on QNB (Qatar National Bank) cards. The
outcome is `Balance`: the average amount on the card, in dollars. Mean
Balance on this file is about $520. Ninety accounts sit at zero.

QNB has already issued the cards. For each account we observe income,
credit limit, credit rating, number of cards, age, years of education,
whether the person owns a home, is married, is a student, and region
(East, South, or West). That is many predictors, so there are many
possible least-squares models. If we put every predictor into one fit,
or we keep adding predictors because the training fit looks better, we
can overfit these 400 accounts. A closer fit here does not mean the
model will predict well for a new customer.

The question for the hour is the one from class: how can we select a
good predictive model for Balance among many, and how can we avoid
overfitting?

We will search with three methods from class: best subset, forward
stepwise, and backward stepwise. That gives the best model of each size.
Then we will pick a size in the two ways we discussed.

- Cp, AIC, BIC, and adjusted $R^2$ adjust the training error. We will
  compute those on all 400 accounts and label that fact.
- Cross-validation scores accounts we did not use to pick the
  predictors. We will do that only on the 280 training accounts.

Do not pick predictors from the $p$-values of the model that uses every
predictor. We will see why that fails when two predictors move together.

Open the case PDF before you run code and keep the codebook in view.
Income is in thousands of dollars. Balance, Limit are in dollars. Rating
is in points. `Student`, `Own`, and `Married` are Yes/No. `Region` is
East, South, or West.

**Think 1.** Why can a model that fits these 400 accounts very well still
be a poor model for the next QNB customer?

**Think 2.** Ninety accounts have Balance exactly zero, and the mean is
about $520. What should you keep in mind when you read a least-squares
slope on this file?

# Loading Packages and Data

```{r}
#| label: setup
# leaps: regsubsets() for best subset, forward, and backward search.
# ggplot2 / patchwork / scales: figures that we can read in the room.
# Why? Base par(mfrow) stacks plots. patchwork keeps the comparison
# aligned and labeled, which is the point of the criteria panel later.
if (!requireNamespace("leaps", quietly = TRUE)) {
  install.packages("leaps")
}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
if (!requireNamespace("patchwork", quietly = TRUE)) {
  install.packages("patchwork")
}
if (!requireNamespace("scales", quietly = TRUE)) {
  install.packages("scales")
}

library(leaps)
library(ggplot2)
library(patchwork)
library(scales)

options(scipen = 999)

# Dollar-scale RMSE. Why? Balance is in dollars. A unitless MSE is
# harder to hold next to the $520 mean.
rmse <- function(y, yhat) {
  sqrt(mean((y - yhat)^2))
}

# Names of the predictors in a regsubsets() model of a given size.
# Why? We will compare teams often. One helper keeps the output short.
selected_at <- function(object, size) {
  setdiff(names(coef(object, id = size)), "(Intercept)")
}

# Turn a regsubsets team into an lm() formula.
# Why? leaps reports dummy names (StudentYes, RegionWest). lm() wants
# the original factor (Student, Region). After class, if you change the
# seed, use this helper so the fit matches the new team.
team_formula <- function(object, size) {
  vars <- selected_at(object, size)
  vars <- sub("Yes$", "", vars)
  vars <- sub("^Region.*", "Region", vars)
  vars <- unique(vars)
  as.formula(paste("Balance ~", paste(vars, collapse = " + ")))
}
```

```{r}
#| label: load-data
# Read the shipped file. Do not load a different Credit table from a
# package: this file uses Region (East / South / West) and has no extra
# demographic predictor beyond the codebook.
credit <- read.csv("data/Credit.csv")

# Why? read.csv() leaves Yes/No and Region as character. lm() and
# regsubsets() need factors so the dummy coding is explicit.
credit$Own <- factor(credit$Own)
credit$Student <- factor(credit$Student)
credit$Married <- factor(credit$Married)
credit$Region <- factor(credit$Region)

str(credit)
summary(credit)

n <- nrow(credit)
n
mean(credit$Balance)
sum(credit$Balance == 0)
sum(is.na(credit))

# Mean Balance by the Yes/No and Region predictors.
# Why? The Discuss below is about who actually sits away from $520.
tapply(credit$Balance, credit$Student, mean)
tapply(credit$Balance, credit$Own, mean)
tapply(credit$Balance, credit$Married, mean)
tapply(credit$Balance, credit$Region, mean)

# Question: How many accounts? How many predictors sit beside Balance?
# Confirm there are no missing values before you search.
#
# Interpret: Mean Balance should be about $520. The 90 zeros are real
# accounts, not missing values. Least squares will spend some fit on
# that pile at zero.
#
# Discuss: Own, Married, and Region means sit close to the file mean.
# Student does not. Why might a student indicator move Balance when
# home ownership barely does?
```

```{r}
#| label: hist-balance
#| fig-width: 8
#| fig-height: 4.2
ggplot(credit, aes(x = Balance)) +
  geom_histogram(bins = 36, fill = "#156082", color = "white", linewidth = 0.2) +
  geom_vline(
    xintercept = mean(credit$Balance),
    color = "#e97132",
    linewidth = 1.1
  ) +
  scale_x_continuous(labels = label_dollar(accuracy = 1)) +
  labs(
    x = "Balance (dollars)",
    y = "Number of accounts",
    title = "Balance on 400 QNB-card accounts",
    subtitle = "Orange line: mean, about $520. The spike at zero is 90 accounts."
  ) +
  theme_minimal()

# Interpret: The mass at zero and the long right tail both matter. A
# single linear model is a compromise across those accounts.
# Question: If QNB cares more about high-balance accounts than about
# the zeros, is training RMSE on all 400 the loss they would write
# down in a product meeting?
```

```{r}
#| label: split
# Hold out 120 accounts. Search and cross-validation use the other 280.
# Seed 46704 so the room shares one split.
# Why? If we pick predictors on all 400 and then congratulate ourselves
# on a holdout that already influenced the pick, the holdout is no
# longer a test. That is leakage.
set.seed(46704)
test_id <- sample(nrow(credit), 120)
train <- credit[-test_id, ]
test <- credit[test_id, ]

nrow(train)
nrow(test)
mean(train$Balance)
mean(test$Balance)

# Question: The two means need not match. What would it mean for
# selection if we now used the 120 to choose a size?
#
# Discuss: Suppose someone runs best subset on all 400, then reports
# error on these 120. Which step used information from the 120? Why
# is that a problem for a claim about new customers?
#
# The 120 stay closed for the rest of this notebook. We print their
# size and mean only to record the split. Do not score models on test.
```

# Best Subset Selection

Best subset, for each size $k$, finds the $k$ predictors that minimize
training RSS. It is a search over teams, not a search over $p$-values.

As we add predictors the fit can follow the 280 more closely (lower
bias) and also follow noise (higher variance). The outmat is the team
sheet. An asterisk means that predictor is in the best model of that
size.

```{r}
#| label: best-subset
# nvmax = 11 because Region has three levels (two dummies) and the
# three Yes/No variables each contribute one dummy. That is 11
# predictors after dummy coding, not 10 raw names.
# Why? If nvmax is too small, we never see the kitchen-sink team and
# the criteria plots get cut off.
best_train <- regsubsets(Balance ~ ., data = train, nvmax = 11)
best_train_sum <- summary(best_train)
best_train_sum$outmat

cat("Size 1:", paste(selected_at(best_train, 1), collapse = ", "), "\n")
cat("Size 2:", paste(selected_at(best_train, 2), collapse = ", "), "\n")
cat("Size 3:", paste(selected_at(best_train, 3), collapse = ", "), "\n")
cat("Size 4:", paste(selected_at(best_train, 4), collapse = ", "), "\n")

# Interpret: Who enters at size 1? At size 2? At size 3? Student
# typically arrives by size 3. Limit and Rating do not both stay in
# the size-4 team — we will come back to that.
#
# Question: Size 1 is a single numeric score. Why is that predictor
# a plausible first pick for Balance on a credit card?
#
# Discuss: Best subset re-optimizes at every size. The size-4 team
# is not required to contain the size-3 team. Find the first size
# where the nested pattern breaks. What swapped?
#
# Business implication: A search that is allowed to drop a predictor
# later is more honest than a story that says "we already included
# Rating, so it has to stay."
```

```{r}
#| label: best-heatmap
#| fig-width: 8.5
#| fig-height: 4.8
# Optional picture of the same outmat. Dark tiles are selected.
# Why? A table is precise. A tile plot makes the Limit / Rating swap
# and the late arrival of Education visible from across the room.
om <- best_train_sum$outmat
best_tiles <- do.call(
  rbind,
  lapply(seq_len(nrow(om)), function(i) {
    data.frame(
      size = i,
      predictor = colnames(om),
      selected = om[i, ] == "*",
      row.names = NULL
    )
  })
)
best_tiles$predictor <- factor(best_tiles$predictor, levels = colnames(om))

ggplot(best_tiles, aes(x = size, y = predictor, fill = selected)) +
  geom_tile(color = "white", linewidth = 0.4) +
  scale_fill_manual(values = c("TRUE" = "#156082", "FALSE" = "#e8eef2"),
                    labels = c("TRUE" = "In the team", "FALSE" = "Out")) +
  scale_x_continuous(breaks = 1:11) +
  labs(
    x = "Number of predictors",
    y = NULL,
    fill = NULL,
    title = "Best subset on the 280 training accounts",
    subtitle = "Each row is a predictor (after dummy coding). Dark = selected at that size."
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Interpret: Education stays out for a long time. Region dummies
# arrive late. The interesting movement is in Limit, Rating, Cards,
# and Student.
```

# Forward Stepwise Selection

Forward stepwise starts at the intercept and adds one predictor at a
time. Once a predictor is in, it stays. The search is nested by
construction.

```{r}
#| label: forward
fwd_train <- regsubsets(Balance ~ ., data = train, nvmax = 11, method = "forward")
fwd_train_sum <- summary(fwd_train)
fwd_train_sum$outmat

# Student Exercise: Walk sizes 1 through 5 on both outmats. Find the
# first size where the forward team is not the same as the best-subset
# team. Write the two lists of predictors before you look at the
# size-4 section below.
for (i in 1:5) {
  cat(
    "Size", i,
    "| best:", paste(selected_at(best_train, i), collapse = ", "),
    "| forward:", paste(selected_at(fwd_train, i), collapse = ", "),
    "\n"
  )
}
```

```{r}
#| label: forward-adds
#| fig-width: 8
#| fig-height: 5.2
# Which predictor is new at each step?
fwd_path <- data.frame(size = 1:11, added = NA_character_)
already <- character(0)
for (i in 1:11) {
  now <- selected_at(fwd_train, i)
  fwd_path$added[i] <- setdiff(now, already)
  already <- now
}
fwd_path
fwd_path$added <- factor(fwd_path$added, levels = rev(fwd_path$added))

# Why a lollipop, not one row of labels? Eleven names on one line
# overwrite each other. The path should be readable from the back row.
ggplot(fwd_path, aes(x = size, y = added)) +
  geom_segment(aes(x = 0.8, xend = size, yend = added), color = "#8eb4cc") +
  geom_point(size = 3.2, color = "#156082") +
  scale_x_continuous(breaks = 1:11, limits = c(0.8, 11.4)) +
  labs(
    x = "Forward step (model size after the add)",
    y = NULL,
    title = "Predictor added by forward stepwise",
    subtitle = "Once a name appears, forward will not drop it later."
  ) +
  theme_minimal()

# Interpret: Forward commits early. If Rating enters at size 1, Rating
# is still there at size 4. Best subset is allowed to swap Rating for
# Cards when that swap wins on training RSS.
#
# Question: Is a method that cannot change its mind safer, or just
# cheaper to compute?
#
# Discuss: For 11 predictors, best subset is feasible. When would you
# still prefer forward in a business setting?
```

# Backward Stepwise Selection

Backward stepwise starts from the model that uses every predictor and
drops the one whose removal hurts training RSS the least. The path is
nested in the other direction.

```{r}
#| label: backward
bwd_train <- regsubsets(Balance ~ ., data = train, nvmax = 11, method = "backward")
bwd_train_sum <- summary(bwd_train)
bwd_train_sum$outmat

cat("Backward size 1:", paste(selected_at(bwd_train, 1), collapse = ", "), "\n")
cat("Best subset size 1:", paste(selected_at(best_train, 1), collapse = ", "), "\n")
cat("Backward size 4:", paste(selected_at(bwd_train, 4), collapse = ", "), "\n")

# Question: The size-1 backward model need not match the size-1 best
# subset model. Why can those two searches disagree at the small end?
#
# Discuss: Look across the three outmats at sizes 1, 4, and 6. Where
# do the methods agree, and where do they fork? If three searches
# disagree, what should you not do with the first asterisk you see?
#
# Interpret: Backward has already lived with Limit and Rating in the
# same model. When it is forced to shrink, it has to drop one of them
# while the rest of the kitchen sink is still there. That is a
# different question than "which single predictor is best alone."
```

# Training Error by Model Size

Training RMSE is the in-sample fit of the best model of each size on
the 280. It should fall as we add predictors. That is why it cannot
choose a size.

```{r}
#| label: train-rmse
#| fig-width: 7.5
#| fig-height: 4.2
mm_train <- model.matrix(Balance ~ ., data = train)
train_rmse <- sapply(1:11, function(i) {
  cf <- coef(best_train, id = i)
  pred <- as.numeric(mm_train[, names(cf), drop = FALSE] %*% cf)
  rmse(train$Balance, pred)
})

train_err <- data.frame(size = 1:11, rmse = train_rmse)
train_err

ggplot(train_err, aes(x = size, y = rmse)) +
  geom_line(color = "#156082", linewidth = 1) +
  geom_point(color = "#156082", size = 2.4) +
  scale_x_continuous(breaks = 1:11) +
  scale_y_continuous(labels = label_dollar(accuracy = 1)) +
  labs(
    x = "Number of predictors",
    y = "Training RMSE",
    title = "Training RMSE falls as we add predictors",
    subtitle = "280 training accounts. Best subset at each size."
  ) +
  theme_minimal()

# Interpret: The big drop is in the first few sizes. After that the
# curve is almost flat. Education arrives late on this training path
# (size 9 in the heatmap). It will still nudge training RMSE down.
# That nudge is not evidence that Education belongs in a model we
# will take off these 280 accounts.
#
# Question: If you handed this plot to a colleague and said nothing
# else, which size would they pick? Why is that the wrong prompt?
#
# Discuss: Why do we need Cp / AIC / BIC / adjusted R^2, or
# cross-validation, if training RMSE already tells a clear story?
#
# Business implication: A dashboard that only reports training fit
# will keep asking for more predictors. That is a process failure,
# not a data failure.
```

# Comparing Best Subset and Forward Stepwise at Size 4

Size 4 is the first place the two searches usually disagree on this
file. The disagreement is about Limit, Rating, and Cards.

```{r}
#| label: size4-teams
cat("Best subset, size 4:\n")
print(selected_at(best_train, 4))
cat("\nForward, size 4:\n")
print(selected_at(fwd_train, 4))
cat("\nBackward, size 4:\n")
print(selected_at(bwd_train, 4))

# Same size, different teams. Fit each one on the 280 so the RSS
# comparison is in dollars.
size4 <- list(
  "Income + Limit + Cards + Student"  = Balance ~ Income + Limit + Cards + Student,
  "Income + Limit + Rating + Student" = Balance ~ Income + Limit + Rating + Student,
  "Income + Rating + Cards + Student" = Balance ~ Income + Rating + Cards + Student,
  "Income + Limit + Rating + Cards"   = Balance ~ Income + Limit + Rating + Cards
)

size4_err <- data.frame(
  team = names(size4),
  rmse = sapply(size4, function(f) rmse(train$Balance, fitted(lm(f, data = train)))),
  row.names = NULL
)
size4_err
```

```{r}
#| label: size4-plot
#| fig-width: 8
#| fig-height: 4.4
size4_err$team <- factor(size4_err$team, levels = rev(size4_err$team))

ggplot(size4_err, aes(x = rmse, y = team)) +
  geom_col(fill = "#156082", width = 0.65) +
  geom_text(
    aes(label = dollar(rmse, accuracy = 0.1)),
    hjust = -0.08,
    size = 3.2
  ) +
  scale_x_continuous(
    labels = label_dollar(accuracy = 1),
    expand = expansion(mult = c(0.02, 0.18))
  ) +
  labs(
    x = "Training RMSE on the 280",
    y = NULL,
    title = "Four size-4 teams, same training accounts",
    subtitle = "The last bar drops Student. That omission is expensive."
  ) +
  theme_minimal()

# Interpret: Best subset's size-4 team keeps Cards and drops Rating.
# Forward keeps Rating and drops Cards. The training RMSEs are close.
# The team that drops Student is not close.
#
# Question: Limit and Rating move together. If you could keep only
# one of them at size 4, what else do you want on the team?
#
# Discuss: A colleague says "both Limit and Rating are important, so
# both should be in the model." What does this plot say back?
#
# Business implication: Importance in a sentence is not the same as
# a seat on a four-predictor team. Student is easy to miss in a
# credit-limit conversation and expensive to omit on this file.
```

# Model Comparison (Cp, AIC, BIC, Adjusted R²)

These are the criteria from class that adjust training error. We compute
them on **all 400 accounts** and we label that. They are not an estimate
of error on new customers. They are a ranking of sizes on the file we
already have.

leaps reports Cp, BIC, and adjusted $R^2$. AIC we get by refitting the
best model of each size with `lm()`.

```{r}
#| label: criteria-all400
# Why? The outline for the hour is: criteria on all 400, labeled;
# cross-validation on the 280. Mixing those two without a label is
# how people later claim an "honest" number that used every account.
best_all <- regsubsets(Balance ~ ., data = credit, nvmax = 11)
best_all_sum <- summary(best_all)

aic_vals <- numeric(11)
mm_all <- model.matrix(Balance ~ ., data = credit)
for (i in 1:11) {
  cf <- coef(best_all, id = i)
  vars <- setdiff(names(cf), "(Intercept)")
  dat_i <- data.frame(
    Balance = credit$Balance,
    mm_all[, vars, drop = FALSE],
    check.names = FALSE
  )
  aic_vals[i] <- AIC(lm(Balance ~ ., data = dat_i))
}

crit <- data.frame(
  size = 1:11,
  Cp = as.numeric(best_all_sum$cp),
  AIC = aic_vals,
  BIC = as.numeric(best_all_sum$bic),
  adjR2 = as.numeric(best_all_sum$adjr2)
)
crit

cat("Cp minimum at size", which.min(crit$Cp), "\n")
cat("AIC minimum at size", which.min(crit$AIC), "\n")
cat("BIC minimum at size", which.min(crit$BIC), "\n")
cat("Adjusted R^2 maximum at size", which.max(crit$adjR2), "\n")

cat("\nBest subset on all 400, size 4:", paste(selected_at(best_all, 4), collapse = ", "), "\n")
cat("Best subset on all 400, size 6:", paste(selected_at(best_all, 6), collapse = ", "), "\n")
cat("Best subset on all 400, size 7:", paste(selected_at(best_all, 7), collapse = ", "), "\n")
```

```{r}
#| label: criteria-panel
#| fig-width: 9
#| fig-height: 6.4
# Four panels, one comparison. Orange marks the preferred size.
# Why? A single number from one criterion is easy to over-trust.
# Seeing the four together is the exercise.
# Orange marker for the preferred size. Built as a one-row data frame
# so we do not fight ggplot's inherited y mapping.
mark <- function(df, y, which_row) {
  hit <- df[which_row, , drop = FALSE]
  geom_point(
    data = data.frame(size = hit$size, y = hit[[y]]),
    aes(x = size, y = y),
    color = "#e97132",
    size = 3.2,
    inherit.aes = FALSE
  )
}

vline_at <- function(size) {
  geom_vline(xintercept = size, color = "#e97132", linetype = "dashed",
             linewidth = 0.5, alpha = 0.7)
}

p_cp <- ggplot(crit, aes(x = size, y = Cp)) +
  vline_at(which.min(crit$Cp)) +
  geom_line(color = "#156082", linewidth = 0.9) +
  geom_point(color = "#156082", size = 2) +
  mark(crit, "Cp", which.min(crit$Cp)) +
  scale_x_continuous(breaks = 1:11) +
  labs(x = "Number of predictors", y = expression(C[p]), title = "Mallows Cp") +
  theme_minimal()

p_aic <- ggplot(crit, aes(x = size, y = AIC)) +
  vline_at(which.min(crit$AIC)) +
  geom_line(color = "#156082", linewidth = 0.9) +
  geom_point(color = "#156082", size = 2) +
  mark(crit, "AIC", which.min(crit$AIC)) +
  scale_x_continuous(breaks = 1:11) +
  labs(x = "Number of predictors", y = "AIC", title = "AIC from lm() at each size") +
  theme_minimal()

p_bic <- ggplot(crit, aes(x = size, y = BIC)) +
  vline_at(which.min(crit$BIC)) +
  geom_line(color = "#156082", linewidth = 0.9) +
  geom_point(color = "#156082", size = 2) +
  mark(crit, "BIC", which.min(crit$BIC)) +
  scale_x_continuous(breaks = 1:11) +
  labs(x = "Number of predictors", y = "BIC", title = "BIC") +
  theme_minimal()

p_adj <- ggplot(crit, aes(x = size, y = adjR2)) +
  vline_at(which.max(crit$adjR2)) +
  geom_line(color = "#156082", linewidth = 0.9) +
  geom_point(color = "#156082", size = 2) +
  mark(crit, "adjR2", which.max(crit$adjR2)) +
  scale_x_continuous(breaks = 1:11) +
  labs(x = "Number of predictors", y = expression("Adjusted " ~ R^2),
       title = "Adjusted R-squared") +
  theme_minimal()

(p_cp | p_aic) / (p_bic | p_adj) +
  plot_annotation(
    title = "Strategy 1 criteria on all 400 accounts",
    subtitle = "Orange point: preferred size for that criterion. These numbers use every account."
  )

# Interpret: Cp and AIC agree on one size. BIC prefers a smaller
# model. Adjusted R^2 prefers a larger one. That disagreement is
# the lesson, not a bug in the code.
#
# Question: BIC penalizes extra predictors more heavily than AIC.
# On a file whose mean Balance is about $520, which way of scoring
# extra predictors is closer to how you would defend a model?
#
# Discuss: These four numbers all used the same 400 accounts we
# might later want to claim as "new." What can you still use them
# for, and what must you not use them for?
#
# Student Exercise: Defend BIC's size against a colleague who wants
# the adjusted-R^2 size. Use the training-RMSE curve, the $520 mean,
# and the list of predictors that enter after size 4. Write four
# sentences, not a dump of coefficients.
```

# Limit, Rating, and p-values

Limit and Rating almost duplicate each other on this file. That is why
selection and $p$-values tell different stories.

```{r}
#| label: limit-rating
cor(credit$Limit, credit$Rating)
cor(credit$Limit, credit$Balance)
cor(credit$Rating, credit$Balance)

fit_limit  <- lm(Balance ~ Limit, data = credit)
fit_rating <- lm(Balance ~ Rating, data = credit)
fit_both   <- lm(Balance ~ Limit + Rating, data = credit)

summary(fit_limit)
summary(fit_rating)
summary(fit_both)

# Kitchen sink on all 400. Why? The Question below is about stars in
# the full model, not only Limit + Rating. On this file both Limit
# and Rating stay starred once everything else is in the equation.
fit_sink <- lm(Balance ~ ., data = credit)
summary(fit_sink)

# Slopes, standard errors, and p-values side by side.
# Why format.pval()? scipen = 999 makes raw p-values unreadable.
brief <- function(fit, label) {
  s <- summary(fit)$coefficients
  data.frame(
    model = label,
    predictor = rownames(s),
    slope = round(s[, "Estimate"], 4),
    se = round(s[, "Std. Error"], 4),
    p = format.pval(s[, "Pr(>|t|)"], digits = 3, eps = 0.001),
    r2 = round(summary(fit)$r.squared, 4),
    rse = round(summary(fit)$sigma, 2),
    row.names = NULL
  )
}
rbind(
  brief(fit_limit, "Limit only"),
  brief(fit_rating, "Rating only"),
  brief(fit_both, "Limit + Rating")
)

# Why not select by p-values?
# 1. A p-value in a joint model asks a conditional question: does this
#    predictor help given the others already in the equation? That is
#    not a search over models.
# 2. When two predictors move together, the same slope can look
#    decisive alone and look like noise once its twin is in the room.
# 3. A small p-value can sit next to almost no gain in fit. Look at
#    R^2 and RSE for Rating alone versus Limit + Rating.
#
# Interpret: Alone, Limit and Rating each look like the model. Together,
# one slope keeps a small p-value and the other does not. The RSE of
# the two-predictor fit is not an improvement on Rating alone.
#
# Question: Look at the kitchen-sink stars. Do Limit and Rating both
# stay starred there? If you had only those stars, would you have
# kept both? Why is that a bad selection rule on this file?
#
# Discuss: Someone says "Rating is significant, Limit is not, so drop
# Limit." What did they condition on? What would they have said if
# they had run the two simple models first?
#
# Business implication: Do not let a software star decide a credit
# model when the two strongest scores are nearly the same number
# twice. Selection is a comparison of models, not a row of p-values.
```

# Choosing Among Models with Cross-Validation

The criteria above used all 400 accounts. Cross-validation on the 280
is the second way from class: score accounts that were not used to pick
the team of a given size.

`regsubsets()` has no `predict()` method, so we add one. The helper
rebuilds the model matrix, lines up the coefficients, and multiplies.
We then run best subset **inside each training fold**. That is the
step people skip, and it is the step that keeps selection from leaking
across the fold.

```{r}
#| label: predict-helper
# Why? Without this helper we would have to rebuild the matrix math
# in every fold. The function is the same idea as predict(lm()):
# new accounts in, fitted Balance out, for a chosen size.
predict.regsubsets <- function(object, newdata, id, ...) {
  form <- as.formula(object$call[[2]])
  mat <- model.matrix(form, newdata)
  coefi <- coef(object, id = id)
  xvars <- names(coefi)
  mat[, xvars] %*% coefi
}
```

```{r}
#| label: cv-size
#| fig-width: 7.8
#| fig-height: 7.2
# 10-fold CV over size, on the 280 only.
# Why a new set.seed(46704)? The fold assignment is random. Fixing it
# means the room is looking at the same picture. It is not a claim
# that this fold cut is magic.
k <- 10
set.seed(46704)
folds <- sample(rep(1:k, length = nrow(train)))
cv_mse <- matrix(NA, k, 11, dimnames = list(NULL, paste(1:11)))

for (j in 1:k) {
  # Selection happens on 9 folds only.
  fold_fit <- regsubsets(
    Balance ~ .,
    data = train[folds != j, ],
    nvmax = 11
  )
  for (i in 1:11) {
    pred <- predict(fold_fit, train[folds == j, ], id = i)
    y_fold <- train$Balance[folds == j]
    cv_mse[j, i] <- mean((y_fold - pred)^2)
  }
}

# Average MSE across folds, then take the square root so the axis is
# in dollars. Why? That matches how we have been reading training RMSE.
cv_rmse <- sqrt(colMeans(cv_mse))
cv_rmse
as.integer(which.min(cv_rmse))

err_both <- rbind(
  data.frame(size = 1:11, rmse = train_rmse, source = "Training (280)"),
  data.frame(size = 1:11, rmse = as.numeric(cv_rmse), source = "10-fold CV (280)")
)

# Two panels: the full drop, then a zoom so the flat band is visible.
# Why? Size 1–2 stretch the y-axis. The decision among sizes 4–7 lives
# in a few dollars, and that is the part a single plot hides.
p_cv_full <- ggplot(err_both, aes(x = size, y = rmse, color = source, group = source)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.4) +
  scale_color_manual(values = c("Training (280)" = "#156082",
                                "10-fold CV (280)" = "#e97132")) +
  scale_x_continuous(breaks = 1:11) +
  scale_y_continuous(labels = label_dollar(accuracy = 1)) +
  labs(
    x = "Number of predictors",
    y = "RMSE (dollars)",
    color = NULL,
    title = "Training RMSE versus 10-fold CV RMSE",
    subtitle = "Selection is re-run inside each fold. The 120 held-out accounts are not used."
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

p_cv_zoom <- ggplot(
  subset(err_both, size >= 3),
  aes(x = size, y = rmse, color = source, group = source)
) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.4) +
  scale_color_manual(values = c("Training (280)" = "#156082",
                                "10-fold CV (280)" = "#e97132")) +
  scale_x_continuous(breaks = 3:11) +
  scale_y_continuous(labels = label_dollar(accuracy = 1)) +
  labs(
    x = "Number of predictors",
    y = "RMSE (dollars)",
    color = NULL,
    title = "Same comparison, sizes 3 to 11",
    subtitle = "CV sits in a band. The exact minimum is less important than the band."
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

p_cv_full / p_cv_zoom

# Interpret: Training RMSE keeps falling. CV RMSE drops sharply at
# first and then sits in a band. The exact minimum of a wiggly CV
# curve is less important than that band.
#
# Question: Which sizes are in the flat region? Does BIC's size from
# the 400-account panel sit in that region?
#
# Discuss: We re-ran best subset inside each fold. What would go
# wrong if we picked the size-4 team once on all 280 and only
# cross-validated that one team?
#
# Business implication: A one-dollar CV gap on a $520 outcome is not
# a reason to add three more predictors. Flat is a finding.
```

# Choosing a Model

We need one model we are willing to take off the 280 accounts. BIC on
the 400 preferred a small model. CV on the 280 is already flat in that
neighborhood. We take the size-4 best-subset team and fit it with
`lm()` on the training accounts.

The 120 held-out accounts stay closed. We are not reporting a test
RMSE in this notebook.

```{r}
#| label: chosen-model
# Size 4 on the 280: on this seed, Income, Limit, Cards, Student.
# Why this size? It is BIC's pick on the 400, and it sits in the
# flat CV region on the 280. Size 6 can win a CV minimum by a
# fraction of a dollar. That is not a strong argument for the
# extra predictors on a $520 outcome.
team_formula(best_train, 4)
chosen <- lm(team_formula(best_train, 4), data = train)
summary(chosen)
coef(chosen)

# Same team from regsubsets(), for the coefficient table without
# the extra lm() printout.
coef(best_train, 4)

# Interpret the slopes on the training fit:
# - Income is in thousands of dollars. Holding Limit, Cards, and
#   Student fixed, a $10,000 higher income is 10 times the Income
#   coefficient. The sign is negative on this file.
# - Limit is in dollars. A $1,000 higher limit is 1000 times the
#   Limit coefficient.
# - Cards is a count. One extra card on the account, holding the
#   rest fixed.
# - StudentYes is the difference in fitted Balance for a student
#   versus the reference (not a student), holding the rest fixed.
#
# Question: Why can Income have a negative slope after we control
# for Limit? What would go wrong if someone read that slope as
# "raise income so that balances fall"?
#
# Business implication: This is a predictive model for Balance on
# accounts QNB has already issued. It is not a lever list. Do not
# talk as if the bank can flip Student, or as if raising Limit will
# cause Balance to rise by the slope. Those are associations on
# this file, after the cards exist.
#
# Discuss: If you had to defend this four-predictor model to a
# colleague who wants the kitchen sink because "more information
# is better," what two pictures from this notebook do you open?
#
# Student Exercise: Fit the size-6 best-subset team on the same 280
# accounts. Compare the Limit slope, the Student gap, and the
# training RMSE with size 4. What moved enough to matter at a $520
# mean, and what did not?
```

```{r}
#| label: chosen-size6-exercise
# Filled so you can run the comparison in the room. Write the
# interpretation yourself.
size6_team <- selected_at(best_train, 6)
size6_team
team_formula(best_train, 6)
fit6 <- lm(team_formula(best_train, 6), data = train)
coef(chosen)
coef(fit6)
rmse(train$Balance, fitted(chosen))
rmse(train$Balance, fitted(fit6))
```

# After class

Work these on your own after the hour. Do not open the 120 accounts
unless you are later told to.

1. **Second seed.** Change the seed that draws the 120 held-out
   accounts (or the seed that draws the 10 folds). Rerun the search
   and the CV plot. Does the flat region still cover size 4? What
   moved: the team, the size, or only the dollars?

2. **Validation versus 10-fold.** On the 280, hold out one validation
   slice (for example 70 accounts). Pick a size by validation RMSE.
   Compare that size with the 10-fold curve. Why can a single split
   be noisier than 10 folds as a way to choose among sizes?

3. **Three reasons against $p$-value selection.** Write three short
   reasons, in your own words, that you would not pick predictors
   from the kitchen-sink $p$-values on this file. Use Limit and
   Rating in at least one reason.

4. **Optional: Education.** Add Education to the size-4 training
   model. What happens to RMSE and to the other slopes? When would
   you still keep Education because a person asked for it, even if
   the fit does not want it?
