---
title: "Lab 3: Model selection. Selecting a model for Balance on a QNB credit card"
subtitle: "70-467 · Monday 21 September"
author: "Andrés Castaño Zuluaga"
date: today
format:
  html:
    toc: true
    toc-depth: 2
    toc-location: right
    body-classes: "mlba-page"
    code-fold: false
    code-tools: true
execute:
  echo: true
  warning: false
  message: false
---

<!--
Student starter for the hour. Code is filled on purpose. Run each chunk
and read the comments. We will talk through only some of the Question
and Discuss prompts in the room. The 120 held-out accounts stay unused
until we have chosen a model. Then we score that model once.
-->

# Introduction to the Business Problem

We have been using the Credit data in class. For this lab, treat the 400
observations as credit-card accounts on QNB (Qatar National Bank) cards. The
outcome is `Balance`: the average amount on the card, in dollars. Mean
Balance in this data is about $520. Ninety accounts sit at zero.

QNB has already issued the cards. For each account we observe income,
credit limit, credit rating, number of cards, age, years of education,
whether the person owns a home, is married, is a student, and region
(East, South, or West). That is many predictors, so there are many
possible least-squares models. If we put every predictor into one fit,
or we keep adding predictors because the training fit looks better, we
can overfit these 400 accounts. A closer fit here does not mean the
model will predict well for a new customer.

The question for the hour is the one from class: how can we select a
good predictive model for Balance among many, and how can we avoid
overfitting?

We will search with three methods from class: best subset, forward
stepwise, and backward stepwise. That gives the best model of each size.
Then we will pick a size in the two ways we discussed in class.

We hold out 120 accounts. Search and the two size rules use the other
280. The 120 stay unused until we have a model. Then we score that
model once.

- **Indirectly.** Cp, AIC, BIC, and adjusted $R^2$ adjust the training
  error. We compute those on the 280 training accounts.
- **Directly.** Cross-validation estimates the error on a new customer
  without using the 120. In each fold we score training accounts that
  were not used to pick the predictors of that size. A single
  validation sample is in the after-class questions.

We will also see why you should not pick predictors from the
$p$-values of the model that uses every predictor. If Limit and Rating
both have small $p$-values in one summary, does that mean both belong
in a model for Balance?

Keep the case PDF and the codebook open while you run code. Income is
in thousands of dollars. Balance and Limit are in dollars. Rating is
in points. `Student`, `Own`, and `Married` are Yes/No. `Region` is
East, South, or West. A slope is only readable if you know the unit.

**Before you run anything.** A model can fit these 400 accounts very
well and still be a poor model for the next QNB customer. Why? What
goes wrong for the bank if you use that model? Talk this through with
the person next to you. We will come back to it when training RMSE
keeps falling.

Ninety accounts have Balance exactly zero. Are those missing values,
or real accounts that carry no balance? We treat them as real, and we
keep them. A model for who carries a balance, and another for how
large it is, is not this lab.

# Roadmap

1. **Load the data and hold out 120 accounts.** See Balance, then
   set 120 accounts aside so they cannot influence the search or the
   size.
2. **Find the best model of each size.** Best subset, then forward,
   then backward.
3. **See why training RMSE cannot choose a size.** The training fit
   keeps falling as we add predictors.
4. **See when two searches disagree.** Best subset and forward at
   size 4. Skip the ranking plot if the hour is short.
5. **Pick a size by adjusting training error.** Cp, AIC, BIC, and
   adjusted $R^2$ on the 280.
6. **See why $p$-values are not a search.** Limit and Rating move
   together.
7. **Pick a size by estimating test error.** Cross-validation on the
   280, with the search inside each fold.
8. **Commit to one model and score it once.** Refit on the 280, then
   open the 120.
9. **After class.** A validation sample, a second seed, and the
   remaining questions.

# Load the data and hold out 120 accounts

```{r}
#| label: setup
# leaps: regsubsets() runs best subset, forward, and backward selection.
# ggplot2, patchwork, and scales: plots for the lab.
# Why patchwork? It places several plots in one figure with shared
# labels, which we will need for the four criteria later.
if (!requireNamespace("leaps", quietly = TRUE)) {
  install.packages("leaps")
}
if (!requireNamespace("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}
if (!requireNamespace("patchwork", quietly = TRUE)) {
  install.packages("patchwork")
}
if (!requireNamespace("scales", quietly = TRUE)) {
  install.packages("scales")
}

library(leaps)
library(ggplot2)
library(patchwork)
library(scales)

options(scipen = 999)

# RMSE in dollars. Balance is in dollars, so this is easier to
# compare with the mean of about $520 than a unitless MSE.
rmse <- function(y, yhat) {
  sqrt(mean((y - yhat)^2))
}

# Names of the predictors in a regsubsets() model of a given size.
# We will compare models often. This helper keeps the printed lists short.
selected_at <- function(object, size) {
  setdiff(names(coef(object, id = size)), "(Intercept)")
}

# Build an lm() formula from a regsubsets() model of a given size.
# leaps reports dummy names (StudentYes, RegionWest). lm() needs the
# original factor names (Student, Region). If only one Region dummy
# was selected, this formula puts the whole Region factor back. On
# this seed the size-4 model does not include Region. For the test
# score we use predict() on the regsubsets fit, not this formula.
team_formula <- function(object, size) {
  vars <- selected_at(object, size)
  vars <- sub("Yes$", "", vars)
  vars <- sub("^Region.*", "Region", vars)
  vars <- unique(vars)
  as.formula(paste("Balance ~", paste(vars, collapse = " + ")))
}
```

```{r}
#| label: load-data
# Read the csv in the data folder. Do not load a different Credit
# table from a package: this csv uses Region (East / South / West)
# and matches the codebook.
credit <- read.csv("data/Credit.csv")

# read.csv() leaves Yes/No and Region as character. lm() and
# regsubsets() need factors so the dummy coding is explicit.
credit$Own <- factor(credit$Own)
credit$Student <- factor(credit$Student)
credit$Married <- factor(credit$Married)
credit$Region <- factor(credit$Region)

str(credit)
summary(credit)

n <- nrow(credit)
n
mean(credit$Balance)
sum(credit$Balance == 0)
sum(is.na(credit))

# Mean Balance for each group. Compare these with the overall mean
# of about $520.
tapply(credit$Balance, credit$Student, mean)
tapply(credit$Balance, credit$Own, mean)
tapply(credit$Balance, credit$Married, mean)
tapply(credit$Balance, credit$Region, mean)

# Question: How many accounts? How many predictors are there besides Balance?
# Confirm there are no missing values before you search.
#
# Interpret: Mean Balance should be about $520. The 90 zeros are real
# accounts, not missing values. Least squares will try to fit those
# zeros as well as the rest of the accounts.
#
# Discuss: The means for Own, Married, and Region are close to the
# overall mean. The mean for Student is not. Why might a student
# indicator move Balance when home ownership barely does?
```

```{r}
#| label: hist-balance
#| fig-width: 8
#| fig-height: 4.2
ggplot(credit, aes(x = Balance)) +
  geom_histogram(bins = 36, fill = "#156082", color = "white", linewidth = 0.2) +
  geom_vline(
    xintercept = mean(credit$Balance),
    color = "#e97132",
    linewidth = 1.1
  ) +
  scale_x_continuous(labels = label_dollar(accuracy = 1)) +
  labs(
    x = "Balance (dollars)",
    y = "Number of accounts",
    title = "Balance on 400 QNB-card accounts",
    subtitle = "Orange line: mean, about $520. The spike at zero is 90 accounts."
  ) +
  theme_minimal()

# Interpret: The spike at zero and the long right tail both matter.
# One linear model is a compromise across those accounts. We still
# keep the zeros. They are accounts, not missing values.
```

```{r}
#| label: split
# Hold out 120 accounts. Search and cross-validation use the other 280.
# Seed 46704 so everyone in class uses the same split.
# Why? If we choose predictors using all 400 accounts, and then
# evaluate on a holdout that already influenced that choice, the
# holdout is no longer a test. That is leakage.
set.seed(46704)
test_id <- sample(nrow(credit), 120)
train <- credit[-test_id, ]
test <- credit[test_id, ]

nrow(train)
nrow(test)
mean(train$Balance)
mean(test$Balance)

# Question: The two means can differ. What goes wrong if we use the
# 120 to choose a size?
#
# Discuss: Suppose someone runs best subset on all 400 accounts, then
# reports error on these 120. Which step already used the 120? Why
# is that error not the error on a new customer?
#
# We print the size and mean of the 120 only to record the split.
# Do not use these accounts to pick a size. We will score one model
# on them at the end.
```

# Best subset: the best model of each size

Best subset, for each size $k$, finds the $k$ predictors that minimize
training RSS. It is a search over models, not a search over $p$-values.

As we add predictors the fit can follow the 280 more closely (lower
bias) and also follow noise (higher variance). The table that
`regsubsets` prints shows, for each size, which predictors are in
the best model. An asterisk means that predictor is included.

```{r}
#| label: best-subset
# nvmax = 11 because Region has three levels (two dummies) and the
# three Yes/No variables each contribute one dummy. That is 11
# predictors after dummy coding, not 10 raw names.
best_train <- regsubsets(Balance ~ ., data = train, nvmax = 11)
best_train_sum <- summary(best_train)
# $outmat is the table of selected predictors by model size.
best_train_sum$outmat

cat("Size 1:", paste(selected_at(best_train, 1), collapse = ", "), "\n")
cat("Size 2:", paste(selected_at(best_train, 2), collapse = ", "), "\n")
cat("Size 3:", paste(selected_at(best_train, 3), collapse = ", "), "\n")
cat("Size 4:", paste(selected_at(best_train, 4), collapse = ", "), "\n")

# Interpret: Which predictor enters at size 1? At size 2? At size 3?
# On this seed Student is in by size 3. Limit and Rating do not both
# stay in the size-4 model. We will come back to that.
#
# Question: Size 1 is a single numeric score. Why is that predictor
# a plausible first pick for Balance on a credit card?
#
# Discuss: Best subset chooses the best model of each size from
# scratch. The size-4 model does not have to include every predictor
# from the size-3 model. Find the first size where a predictor that
# was included is later left out. What changed?
#
# Business implication: A method that can drop a predictor later is
# more flexible than a rule that says "we already included Rating,
# so it has to stay."
```

```{r}
#| label: best-heatmap
#| fig-width: 8.5
#| fig-height: 4.8
# Same information as the table above, as a picture. Dark tiles are
# selected.
# Why? The table is precise. The tile plot makes it easier to see
# when Limit and Rating replace each other, and when Education
# enters the model.
om <- best_train_sum$outmat
best_tiles <- do.call(
  rbind,
  lapply(seq_len(nrow(om)), function(i) {
    data.frame(
      size = i,
      predictor = colnames(om),
      selected = om[i, ] == "*",
      row.names = NULL
    )
  })
)
best_tiles$predictor <- factor(best_tiles$predictor, levels = colnames(om))

ggplot(best_tiles, aes(x = size, y = predictor, fill = selected)) +
  geom_tile(color = "white", linewidth = 0.4) +
  scale_fill_manual(values = c("TRUE" = "#156082", "FALSE" = "#e8eef2"),
                    labels = c("TRUE" = "In the model", "FALSE" = "Out")) +
  scale_x_continuous(breaks = 1:11) +
  labs(
    x = "Number of predictors",
    y = NULL,
    fill = NULL,
    title = "Best subset on the 280 training accounts",
    subtitle = "Each row is a predictor (after dummy coding). Dark = selected at that size."
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Interpret: Education stays out for a long time. The Region
# indicators enter late. The movement that matters is in Limit,
# Rating, Cards, and Student.
```

# Forward stepwise: add one predictor at a time

Forward stepwise starts at the intercept and adds one predictor at a
time. Once a predictor is in, it stays. The search is nested by
construction.

```{r}
#| label: forward
fwd_train <- regsubsets(Balance ~ ., data = train, nvmax = 11, method = "forward")
fwd_train_sum <- summary(fwd_train)
fwd_train_sum$outmat

# Student exercise: Walk sizes 1 through 5 in both tables. Find the
# first size where the forward model is not the same as the best-subset
# model. Write the two lists of predictors before you look at the
# size-4 section below.
for (i in 1:5) {
  cat(
    "Size", i,
    "| best:", paste(selected_at(best_train, i), collapse = ", "),
    "| forward:", paste(selected_at(fwd_train, i), collapse = ", "),
    "\n"
  )
}
```

```{r}
#| label: forward-adds
#| fig-width: 8
#| fig-height: 5.2
# Which predictor is new at each step?
fwd_path <- data.frame(size = 1:11, added = NA_character_)
already <- character(0)
for (i in 1:11) {
  now <- selected_at(fwd_train, i)
  fwd_path$added[i] <- setdiff(now, already)
  already <- now
}
fwd_path
fwd_path$added <- factor(fwd_path$added, levels = rev(fwd_path$added))

# Why this plot? Eleven names on one line would overlap. This plot
# keeps the path readable.
ggplot(fwd_path, aes(x = size, y = added)) +
  geom_segment(aes(x = 0.8, xend = size, yend = added), color = "#8eb4cc") +
  geom_point(size = 3.2, color = "#156082") +
  scale_x_continuous(breaks = 1:11, limits = c(0.8, 11.4)) +
  labs(
    x = "Forward step (model size after the add)",
    y = NULL,
    title = "Predictor added by forward stepwise",
    subtitle = "Once a name appears, forward will not drop it later."
  ) +
  theme_minimal()

# Interpret: Once forward adds a predictor, that predictor stays in
# every larger model. If Rating is added first, it is still in the
# size-4 model. Best subset can replace Rating with Cards if that
# combination has a smaller training RSS.
#
# Question: Forward is faster, and it never drops a predictor it has
# already chosen. Is that an advantage for the bank, or only a
# computational shortcut? What could go wrong if an early choice is
# a poor one?
#
# Discuss: With 11 predictors we can run best subset. When would you
# still use forward in a business setting?
```

# Backward stepwise: drop one predictor at a time

Backward stepwise starts from the model that uses every predictor and
drops the one whose removal hurts training RSS the least. The path is
nested in the other direction. If the hour is short, read the two
printed lines and the Interpret comment, then move on.

```{r}
#| label: backward
bwd_train <- regsubsets(Balance ~ ., data = train, nvmax = 11, method = "backward")
bwd_train_sum <- summary(bwd_train)
bwd_train_sum$outmat

cat("Backward size 1:", paste(selected_at(bwd_train, 1), collapse = ", "), "\n")
cat("Best subset size 1:", paste(selected_at(best_train, 1), collapse = ", "), "\n")
cat("Backward size 4:", paste(selected_at(bwd_train, 4), collapse = ", "), "\n")

# Question: The size-1 backward model need not match the size-1 best
# subset model. Why can those two searches disagree when the model is
# still very small?
#
# Discuss: Look at the three selection tables at sizes 1, 4, and 6.
# Where do the methods agree, and where do they differ? If the three
# searches disagree, what should you not do with the first predictor
# that appears selected?
#
# Interpret: Backward starts from the model that uses every predictor,
# so Limit and Rating are already in together. When it has to drop
# one of them, the other predictors are still in the model. That is
# a different question than "which single predictor is best alone."
```

# Why training RMSE cannot choose a size

Training RMSE is the in-sample fit of the best model of each size on
the 280. It should fall as we add predictors. That is why it cannot
choose a size.

```{r}
#| label: train-rmse
#| fig-width: 7.5
#| fig-height: 4.2
mm_train <- model.matrix(Balance ~ ., data = train)
train_rmse <- sapply(1:11, function(i) {
  cf <- coef(best_train, id = i)
  pred <- as.numeric(mm_train[, names(cf), drop = FALSE] %*% cf)
  rmse(train$Balance, pred)
})

train_err <- data.frame(size = 1:11, rmse = train_rmse)
train_err

ggplot(train_err, aes(x = size, y = rmse)) +
  geom_line(color = "#156082", linewidth = 1) +
  geom_point(color = "#156082", size = 2.4) +
  scale_x_continuous(breaks = 1:11) +
  scale_y_continuous(labels = label_dollar(accuracy = 1)) +
  labs(
    x = "Number of predictors",
    y = "Training RMSE",
    title = "Training RMSE falls as we add predictors",
    subtitle = "280 training accounts. Best subset at each size."
  ) +
  theme_minimal()

# Interpret: The large drop is in the first few sizes. After that
# the curve is almost flat. Education enters late on this training
# path (size 9 in the heatmap). It will still reduce training RMSE
# a little. That small drop is not evidence that Education belongs
# in a model we would use for a new customer.
#
# Question: If you handed this plot to a colleague and said nothing
# else, which size would they pick? Why is that the wrong question?
#
# Discuss: Why do we need Cp, AIC, BIC, adjusted R^2, or
# cross-validation, if training RMSE already tells a clear story?
#
# Business implication: If you only report training fit, you will
# keep adding predictors. That is a problem with the process, not
# with the data.
```

# When two searches disagree

Size 4 is the first place the two searches disagree in this data. The
difference is Limit, Rating, and Cards. Best subset saw every
four-predictor model. Forward saw one path.


```{r}
#| label: size4-teams
# Print the two models before we rank them. The difference is Rating
# versus Cards. Income, Limit, and Student appear in both.
cat("Best subset, size 4:\n")
print(selected_at(best_train, 4))
cat("\nForward, size 4:\n")
print(selected_at(fwd_train, 4))
cat("\nBackward, size 4:\n")
print(selected_at(bwd_train, 4))

coef(best_train, 4)
coef(fwd_train, 4)
```

```{r}
#| label: size4-all330
#| fig-width: 8.4
#| fig-height: 4.6
# There are choose(11, 4) = 330 models with four predictors. Best
# subset ranked all 330 by training RSS. Forward followed one path
# and kept one of those models. We fit every four-predictor model
# on the 280 accounts and plot the ranking so we can see where
# forward lands, and where training RMSE jumps.
X_train <- model.matrix(Balance ~ ., data = train)[, -1]
y_train <- train$Balance
teams4  <- combn(colnames(X_train), 4, simplify = FALSE)
length(teams4)   # should be 330

rmse4 <- vapply(teams4, function(team) {
  fit <- lm.fit(cbind(1, X_train[, team]), y_train)
  sqrt(mean(fit$residuals^2))
}, numeric(1))

team_key <- function(v) paste(sort(v), collapse = " + ")
size4_rank <- data.frame(
  team = vapply(teams4, team_key, character(1)),
  rmse = rmse4,
  stringsAsFactors = FALSE
)
size4_rank <- size4_rank[order(size4_rank$rmse), ]
size4_rank$rank <- seq_len(nrow(size4_rank))

key_best4 <- team_key(names(coef(best_train, 4))[-1])
key_fwd4  <- team_key(names(coef(fwd_train, 4))[-1])
size4_rank$pick <- ifelse(
  size4_rank$team == key_best4, "Best subset",
  ifelse(size4_rank$team == key_fwd4, "Forward", "Other size-4 model")
)

head(size4_rank, 8)
size4_rank[size4_rank$pick != "Other size-4 model", ]
size4_rank[14:17, ]   # RMSE jumps when Student is left out. Read this with the plot.

top40 <- head(size4_rank, 40)
ggplot(top40, aes(rank, rmse, color = pick)) +
  geom_point(size = 2.6) +
  scale_color_manual(
    values = c(
      "Best subset" = "#C41230",
      "Forward" = "#156082",
      "Other size-4 model" = "grey70"
    )
  ) +
  scale_x_continuous(breaks = c(1, 5, 10, 20, 30, 40)) +
  scale_y_continuous(labels = label_dollar(accuracy = 1)) +
  labs(
    x = "Rank among all 330 four-predictor models (training accounts)",
    y = "Training RMSE (USD)",
    color = NULL,
    title = "All 330 four-predictor models, ranked on the 280",
    subtitle = "Red = best subset. Blue = forward. RMSE rises after about rank 15 when Student is left out."
  ) +
  theme_minimal()

# Interpret: Rank 1 is the best-subset model, by construction.
# Forward keeps Rating (added at step 1) and therefore cannot add Cards.
# Where does forward land, and how many dollars separate it from
# rank 1? The top of the ranking is almost flat: Limit and Rating
# are nearly the same information, so swapping them barely changes
# training RMSE.
#
# Question: Training RMSE rises by about $50 after the first fifteen
# or so models. Look at size4_rank[14:17, ]. What do the first
# fifteen models all include that the rest do not? That one
# predictor matters more than the choice between Cards and Rating.
#
# Question: The top three models are a dollar or two apart. Is the
# gap between best subset and forward at size 4 something a bank
# would notice on a new customer? Training RMSE cannot answer that.
#
# Discuss: A colleague says "both Limit and Rating are important, so
# both should be in the model." What does a flat top of this ranking
# say? Being important in a sentence is not the same as belonging
# in a four-predictor model.
#
# Business implication: Student is easy to leave out of a
# conversation about limits, and expensive to omit in this data.
# The jump in RMSE is the dollar cost of that omission on the
# training accounts.
#
# Student exercise: In two sentences, tell a manager why best subset
# and forward disagree at size 4, using the ranks and the dollar gap
# from the table above. Then one sentence on why leaving Student out
# matters more than that disagreement.
```

# Pick a size by adjusting training error

These are the criteria from class that adjust training error. We compute
them on the **same 280 training accounts** we used for the search. They
are a ranking of sizes on those accounts. They are not a score on new
customers.

leaps reports Cp, BIC, and adjusted $R^2$. AIC we get by refitting the
best model of each size with `lm()`. For least squares, Cp and AIC
are the same idea, so they often pick the same size. BIC charges more
per extra predictor, so it often prefers a smaller model.

```{r}
#| label: criteria-train280
# Same 280 accounts as best_train. Do not recompute on all 400:
# that would use the 120 we are holding out.
aic_vals <- numeric(11)
for (i in 1:11) {
  cf <- coef(best_train, id = i)
  vars <- setdiff(names(cf), "(Intercept)")
  dat_i <- data.frame(
    Balance = train$Balance,
    mm_train[, vars, drop = FALSE],
    check.names = FALSE
  )
  aic_vals[i] <- AIC(lm(Balance ~ ., data = dat_i))
}

crit <- data.frame(
  size = 1:11,
  Cp = as.numeric(best_train_sum$cp),
  AIC = aic_vals,
  BIC = as.numeric(best_train_sum$bic),
  adjR2 = as.numeric(best_train_sum$adjr2)
)
crit

cat("Cp minimum at size", which.min(crit$Cp), "\n")
cat("AIC minimum at size", which.min(crit$AIC), "\n")
cat("BIC minimum at size", which.min(crit$BIC), "\n")
cat("Adjusted R^2 maximum at size", which.max(crit$adjR2), "\n")

cat("\nBest subset on the 280, size 4:", paste(selected_at(best_train, 4), collapse = ", "), "\n")
cat("Best subset on the 280, size 6:", paste(selected_at(best_train, 6), collapse = ", "), "\n")
cat("Best subset on the 280, size 7:", paste(selected_at(best_train, 7), collapse = ", "), "\n")
```

```{r}
#| label: criteria-panel
#| fig-width: 9
#| fig-height: 6.4
# Four panels, one comparison. Orange marks the preferred size.
# Why? A single number from one criterion is easy to over-trust.
# Seeing the four together is the point of this figure.
# Orange marker for the preferred size. Built as a one-row data frame
# so it does not conflict with ggplot's y mapping.
mark <- function(df, y, which_row) {
  hit <- df[which_row, , drop = FALSE]
  geom_point(
    data = data.frame(size = hit$size, y = hit[[y]]),
    aes(x = size, y = y),
    color = "#e97132",
    size = 3.2,
    inherit.aes = FALSE
  )
}

vline_at <- function(size) {
  geom_vline(xintercept = size, color = "#e97132", linetype = "dashed",
             linewidth = 0.5, alpha = 0.7)
}

p_cp <- ggplot(crit, aes(x = size, y = Cp)) +
  vline_at(which.min(crit$Cp)) +
  geom_line(color = "#156082", linewidth = 0.9) +
  geom_point(color = "#156082", size = 2) +
  mark(crit, "Cp", which.min(crit$Cp)) +
  scale_x_continuous(breaks = 1:11) +
  labs(x = "Number of predictors", y = expression(C[p]), title = "Mallows Cp") +
  theme_minimal()

p_aic <- ggplot(crit, aes(x = size, y = AIC)) +
  vline_at(which.min(crit$AIC)) +
  geom_line(color = "#156082", linewidth = 0.9) +
  geom_point(color = "#156082", size = 2) +
  mark(crit, "AIC", which.min(crit$AIC)) +
  scale_x_continuous(breaks = 1:11) +
  labs(x = "Number of predictors", y = "AIC", title = "AIC from lm() at each size") +
  theme_minimal()

p_bic <- ggplot(crit, aes(x = size, y = BIC)) +
  vline_at(which.min(crit$BIC)) +
  geom_line(color = "#156082", linewidth = 0.9) +
  geom_point(color = "#156082", size = 2) +
  mark(crit, "BIC", which.min(crit$BIC)) +
  scale_x_continuous(breaks = 1:11) +
  labs(x = "Number of predictors", y = "BIC", title = "BIC") +
  theme_minimal()

p_adj <- ggplot(crit, aes(x = size, y = adjR2)) +
  vline_at(which.max(crit$adjR2)) +
  geom_line(color = "#156082", linewidth = 0.9) +
  geom_point(color = "#156082", size = 2) +
  mark(crit, "adjR2", which.max(crit$adjR2)) +
  scale_x_continuous(breaks = 1:11) +
  labs(x = "Number of predictors", y = expression("Adjusted " ~ R^2),
       title = "Adjusted R-squared") +
  theme_minimal()

(p_cp | p_aic) / (p_bic | p_adj) +
  plot_annotation(
    title = "Cp, AIC, BIC, and adjusted R-squared on the 280 training accounts",
    subtitle = "Orange point: preferred size for that criterion. These numbers do not use the 120."
  )

# Interpret: Cp and AIC often agree. BIC prefers a smaller model.
# Adjusted R^2 prefers a larger one. That disagreement is the
# lesson, not a bug in the code.
#
# Question: BIC penalizes extra predictors more heavily than AIC.
# Mean Balance here is about $520. Which way of scoring extra
# predictors is closer to how you would defend a model?
#
# Discuss: These four numbers used the 280 training accounts. What
# can you still use them for, and what must you not use them for?
#
# Student exercise: Defend BIC's size against a colleague who wants
# the adjusted-R^2 size. Use the training-RMSE curve, the $520 mean,
# and the list of predictors that enter after size 4. Write four
# sentences, not a list of coefficients.
```

# Why p-values are not a search

Limit and Rating almost duplicate each other in this data. That is why
selection and $p$-values tell different stories. If the hour is short,
run the correlations and the three fits, then read the three reasons.

```{r}
#| label: limit-rating
cor(credit$Limit, credit$Rating)
cor(credit$Limit, credit$Balance)
cor(credit$Rating, credit$Balance)

fit_limit  <- lm(Balance ~ Limit, data = credit)
fit_rating <- lm(Balance ~ Rating, data = credit)
fit_both   <- lm(Balance ~ Limit + Rating, data = credit)

summary(fit_limit)
summary(fit_rating)
summary(fit_both)

# Model with every predictor, on all 400 accounts. The question
# below is about the p-values in that summary, not only Limit and
# Rating. In this data both Limit and Rating still have small
# p-values once everything else is in the model.
fit_sink <- lm(Balance ~ ., data = credit)
summary(fit_sink)

# Slopes, standard errors, and p-values side by side.
# format.pval() keeps the p-values readable. scipen = 999 would
# print them as long decimals.
brief <- function(fit, label) {
  s <- summary(fit)$coefficients
  data.frame(
    model = label,
    predictor = rownames(s),
    slope = round(s[, "Estimate"], 4),
    se = round(s[, "Std. Error"], 4),
    p = format.pval(s[, "Pr(>|t|)"], digits = 3, eps = 0.001),
    r2 = round(summary(fit)$r.squared, 4),
    rse = round(summary(fit)$sigma, 2),
    row.names = NULL
  )
}
rbind(
  brief(fit_limit, "Limit only"),
  brief(fit_rating, "Rating only"),
  brief(fit_both, "Limit + Rating")
)

# Why not select by p-values? Three reasons, in order:
# 1. A p-value in a joint model asks a conditional question: does this
#    predictor help given the others already in the equation? Change
#    the other predictors and the p-value changes. That is not a search
#    over models, and it is not a property of the predictor alone.
# 2. When two predictors move together, the same slope can look
#    decisive alone and look like noise once the other is in the model.
#    With near-duplicates, which one has a small p-value is unstable
#    across samples.
# 3. A small p-value can sit next to almost no gain in fit. Look at
#    R^2 and residual standard error for Rating alone versus Limit +
#    Rating. Eleven tests on the same accounts, with 0.05 as the cutoff,
#    also say nothing about the dollar error on a new customer. No
#    held-out account was used.
#
# Interpret: Three different questions, not one contradiction.
# Alone, Limit and Rating each look like a strong model. Together,
# one slope keeps a small p-value and the other does not, and the
# residual standard error is not an improvement on Rating alone.
# In the model that uses every predictor, both can have small
# p-values. The p-value depends on who else is already in the model.
#
# Question: If you had only the p-values from the model that uses
# every predictor, would you have kept both Limit and Rating? Why
# is that a poor selection rule in this data?
#
# Discuss: Someone says "Rating is significant, Limit is not, so drop
# Limit." What did they condition on? What would they have said if
# they had run the two simple models first?
#
# Business implication: Do not let a software p-value decide a credit
# model when the two strongest scores are nearly the same number
# twice. Selection is a comparison of models, not a row of p-values.
```

# Pick a size by estimating test error

The criteria above used the 280 training accounts. They adjust
training error. They are not a score on new customers.

Cross-validation on the same 280 is the second way from class:
estimate the test error directly, using accounts that were not used to
pick the predictors of a given size. That estimate is what we use to
choose a size. We still do not use the 120 held-out accounts to
choose. Those 120 are for one score after we already have a model.

`regsubsets()` has no `predict()` method, so we add one. The helper
rebuilds the model matrix, lines up the coefficients, and multiplies.
We then run best subset **inside each training fold**. That is the
step that keeps selection from leaking across the fold.

```{r}
#| label: predict-helper
# Without this helper we would have to rebuild the matrix math in
# every fold. The function is the same idea as predict(lm()): new
# accounts in, fitted Balance out, for a chosen size.
predict.regsubsets <- function(object, newdata, id, ...) {
  form <- as.formula(object$call[[2]])
  mat <- model.matrix(form, newdata)
  coefi <- coef(object, id = id)
  xvars <- names(coefi)
  mat[, xvars] %*% coefi
}
```

```{r}
#| label: cv-size
#| fig-width: 7.8
#| fig-height: 7.2
# 10-fold cross-validation over size, on the 280 accounts only.
# Why a new set.seed(46704)? The fold assignment is random. Fixing
# it means everyone is looking at the same picture. It is not a
# claim that this particular fold split is special.
k <- 10
set.seed(46704)
folds <- sample(rep(1:k, length = nrow(train)))
cv_mse <- matrix(NA, k, 11, dimnames = list(NULL, paste(1:11)))

for (j in 1:k) {
  # Selection uses nine folds only.
  fold_fit <- regsubsets(
    Balance ~ .,
    data = train[folds != j, ],
    nvmax = 11
  )
  for (i in 1:11) {
    pred <- predict(fold_fit, train[folds == j, ], id = i)
    y_fold <- train$Balance[folds == j]
    cv_mse[j, i] <- mean((y_fold - pred)^2)
  }
}

# Average MSE across folds, then take the square root so the axis
# is in dollars. That matches how we have been reading training RMSE.
cv_mse_mean <- colMeans(cv_mse)
cv_mse_se <- apply(cv_mse, 2, sd) / sqrt(k)
cv_rmse <- sqrt(cv_mse_mean)
cv_rmse

# Smallest CV error, and the one-standard-error rule from class.
# The rule is applied to the fold MSEs, as in the slides. We plot
# RMSE so the axis stays in dollars.
cv_size <- as.integer(which.min(cv_mse_mean))
cv_size_1se <- min(which(cv_mse_mean <= cv_mse_mean[cv_size] + cv_mse_se[cv_size]))
one_se_line <- sqrt(cv_mse_mean[cv_size] + cv_mse_se[cv_size])
cv_band <- data.frame(
  size = 1:11,
  lo = sqrt(pmax(cv_mse_mean - cv_mse_se, 0)),
  hi = sqrt(cv_mse_mean + cv_mse_se)
)
cat("Smallest CV RMSE at size", cv_size, "\n")
cat("One-standard-error rule: size", cv_size_1se, "\n")
cat("BIC on the 280: size", which.min(crit$BIC), "\n")

err_both <- rbind(
  data.frame(size = 1:11, rmse = train_rmse, source = "Training (280)"),
  data.frame(size = 1:11, rmse = as.numeric(cv_rmse), source = "10-fold CV (280)")
)

# Two panels: the full drop, then a zoom so the flat region is
# visible. Sizes 1 and 2 stretch the y-axis. The choice among sizes
# 4 to 7 is a few dollars, and that is the part a single plot hides.
p_cv_full <- ggplot(err_both, aes(x = size, y = rmse, color = source, group = source)) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.4) +
  scale_color_manual(values = c("Training (280)" = "#156082",
                                "10-fold CV (280)" = "#e97132")) +
  scale_x_continuous(breaks = 1:11) +
  scale_y_continuous(labels = label_dollar(accuracy = 1)) +
  labs(
    x = "Number of predictors",
    y = "RMSE (dollars)",
    color = NULL,
    title = "Training RMSE versus 10-fold CV RMSE",
    subtitle = "Selection is re-run inside each fold. The 120 held-out accounts are not used."
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

p_cv_zoom <- ggplot(
  subset(err_both, size >= 3),
  aes(x = size, y = rmse, color = source, group = source)
) +
  geom_ribbon(
    data = subset(cv_band, size >= 3),
    aes(x = size, ymin = lo, ymax = hi),
    fill = "#e97132",
    alpha = 0.15,
    inherit.aes = FALSE
  ) +
  geom_hline(
    yintercept = one_se_line,
    color = "#C41230",
    linetype = "dotted",
    linewidth = 0.6
  ) +
  geom_vline(xintercept = cv_size, color = "#C41230", linewidth = 0.6) +
  geom_vline(
    xintercept = cv_size_1se,
    color = "#C41230",
    linetype = "dashed",
    linewidth = 0.6
  ) +
  geom_line(linewidth = 1) +
  geom_point(size = 2.4) +
  scale_color_manual(values = c("Training (280)" = "#156082",
                                "10-fold CV (280)" = "#e97132")) +
  scale_x_continuous(breaks = 3:11) +
  scale_y_continuous(labels = label_dollar(accuracy = 1)) +
  labs(
    x = "Number of predictors",
    y = "RMSE (dollars)",
    color = NULL,
    title = "Same comparison, sizes 3 to 11",
    subtitle = "Band: ±1 SE of CV RMSE. Dotted line: minimum + 1 SE. Solid red: CV min. Dashed red: 1-SE size."
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

p_cv_full / p_cv_zoom

# Interpret: Training RMSE keeps falling. Cross-validation RMSE
# drops sharply at first and then stays in a narrow range. The
# solid red line is the size with the smallest CV RMSE: size 6.
#
# The one-standard-error rule: the ten folds are one draw, so the
# CV minimum is estimated with error. The dotted line is that
# minimum plus one standard error. Take the smallest size whose
# CV error is still at or below that line. Those sizes are close
# enough to the minimum that these folds cannot tell them apart.
# On this graph that is the dashed red line: size 3. It is a
# shorter model in the same flat region.
#
# Question: Which sizes are in the flat region? Does BIC's size from
# the 280-account panel sit in that region?
#
# Question: Left of the CV minimum is more bias. Right of it is more
# variance. In a meeting, would you take the CV minimum or the
# one-standard-error size? Why?
#
# Discuss: We re-ran best subset inside each fold. What would go
# wrong if we picked one model on all 280 accounts and only
# cross-validated that one model?
#
# Business implication: A one-dollar CV gap on a $520 outcome is not
# a reason to add three more predictors. A flat curve is itself a
# result. We still have not used the 120 to choose.
```

# Commit to one model and score it once

We now choose one size from the 280 training accounts, before we look
at the 120.

The CV minimum is an estimate. The ten folds are one draw, so the
curve would move a little if we redrew the folds. A size that looks a
dollar or two worse than the minimum may not be worse.

The one-standard-error rule takes the smallest model whose CV error
is still within one SE of that minimum. Those models are not
distinguishable from the minimum given the folds we have. BIC prefers
a smaller model by a different route. On this seed the three answers
are the 1-SE size, BIC's size, and the CV minimum. They need not
agree. That is what the case said.

We take BIC's size: **4**. It sits in the flat CV region, between the
1-SE size and the CV minimum. It does not put Limit and Rating in
together. If every dollar of estimated RMSE were the only concern, we
would take the CV minimum instead. We commit to size 4 before any
test line runs.

Then we refit that model on all 280 training accounts and open the 120
**once**. Test RMSE scores the model we already chose. It is not a
second search. If you use the 120 to pick a size, you no longer have a
test.

```{r}
#| label: chosen-model
# BIC on the 280 is size 4: Income, Limit, Cards, Student.
# Commit before any test line.
chosen_size <- 4
chosen_size
selected_at(best_train, chosen_size)
team_formula(best_train, chosen_size)
chosen <- lm(team_formula(best_train, chosen_size), data = train)
summary(chosen)
coef(best_train, chosen_size)

# Interpret the slopes on the training fit. Read the printed names.
# - Income is in thousands of dollars. Holding Limit, Cards, and
#   Student fixed, a $10,000 higher income is 10 times the Income
#   coefficient. The sign is negative in this data.
# - Limit is in dollars. A $1,000 higher limit is 1000 times the
#   Limit coefficient.
# - Cards is a count. One extra card on the account, holding the
#   rest fixed.
# - StudentYes is the difference in fitted Balance for a student
#   versus the reference (not a student), holding the rest fixed.
#
# Question: Why can Income have a negative slope after we control
# for Limit? What would go wrong if someone read that slope as
# "raise income so that balances fall"?
#
# Question: The CV minimum on this seed is larger and usually keeps
# both Limit and Rating. We did not take that model. Why?
#
# Business implication: This is a predictive model for Balance on
# accounts QNB has already issued. It does not tell you what to
# change. Do not talk as if the bank can change Student, or as if
# raising Limit will cause Balance to rise by the slope. Those are
# associations in this data, after the cards exist.
```

```{r}
#| label: test-once
# Open the 120 once. Score the model we already chose, using the
# regsubsets coefficients (not a rebuilt lm() formula).
# The baseline is the training mean: the miss if we ignored every
# predictor.
pred_test <- predict(best_train, newdata = test, id = chosen_size)
rmse(test$Balance, pred_test)
rmse(test$Balance, rep(mean(train$Balance), nrow(test)))

# Interpret: Test RMSE is the typical miss, in dollars, on accounts
# we did not use to pick the size or the predictors. Put it next to
# the $520 mean and next to the training-mean baseline.
#
# Question: Suppose a colleague now wants to try size 5 and size 7
# on these same 120 and keep the winner. What happens to the claim
# that this number is a test RMSE?
#
# Business implication: One test number is what you can take to a
# new customer. A search on the 120 is not.
```

# After class: extra questions

Work these on your own after the hour. You have already scored the
120 once. Do not go back and pick a different size from that one
number.

1. **Second seed.** Change the seed that draws the 120 held-out
   accounts (or the seed that draws the 10 folds). Rerun the search
   and the CV plot. Does the CV minimum stay at the same size? What
   moved: the predictors, the size, or only the dollars?

2. **Validation versus 10-fold.** On the 280, hold out one validation
   slice (for example 70 accounts). Pick a size by validation RMSE.
   Compare that size with the 10-fold curve. Why can a single split
   be noisier than 10 folds as a way to choose among sizes?

3. **Three reasons against $p$-value selection.** Write three short
   reasons, in your own words, that you would not pick predictors
   from the $p$-values of the model that uses every predictor. Use
   Limit and Rating in at least one reason.

4. **Optional: Education.** Add Education to the model you chose.
   What happens to training RMSE, test RMSE, and the other slopes?
   When would you still keep Education because a person asked for
   it, even if the fit does not want it?

5. **The CV-minimum model, after you already chose.** Size 6 is the
   CV minimum on this seed. Fit it on the 280 and score it on the
   120. Compare training RMSE and test RMSE with size 4. This is no
   longer a test. Would you change your pick after seeing the 120?
   Why not?

```{r}
#| label: after-class-size6
# After class only. Size 6 is the CV minimum. The 120 have already
# been opened. These two test numbers are not a new search.
selected_at(best_train, cv_size)
fit_cvmin <- lm(team_formula(best_train, cv_size), data = train)
coef(best_train, chosen_size)
coef(best_train, cv_size)
rmse(train$Balance, fitted(chosen))
rmse(train$Balance, fitted(fit_cvmin))
rmse(test$Balance, predict(best_train, newdata = test, id = chosen_size))
rmse(test$Balance, predict(best_train, newdata = test, id = cv_size))
```
