---
title: "Lab 2: A price you can take to a house that has not sold yet"
subtitle: "70-467 · Wednesday 9 September"
author: "Andrés Castaño Zuluaga"
date: last-modified
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    code-fold: false
    theme: cosmo
execute:
  echo: true
  warning: false
  message: false
---

Read the case first, including the codebook. You will produce a value for a house that is not in the file, using only facts the office would have before the sale closes, and report test MSE next to the $180,000 mean. Please say whether neighborhood dummies belong, and whether Partial and New belong.

In the room we get through In class. Most chunks already run. After class is the rest of the notebook. The code is already typed.

```{r}
#| label: setup
options(scipen = 999)
library(readr)
library(dplyr)
library(ggplot2)

theme_lab <- function() {
  theme_minimal(base_size = 13) +
    theme(
      panel.grid.minor = element_blank(),
      plot.title = element_text(face = "bold", size = 14),
      legend.position = "top"
    )
}
col_ink <- "#182C4B"
col_red <- "#C41230"
col_grey <- "#6D6E71"
usd <- scales::label_dollar(accuracy = 1)
```

# In class (~75 minutes)

```{r}
#| label: intro-problem
######################################
# Introduction to the business problem
######################################
# Business problem:
# The Ames City Assessor's Office used 2,925 residential sales from 2006 to 2010
# to keep its valuation model current. Mean sale is about $180,000. The office
# still has to put a number on a house that is not in this file, using only facts
# it would have before the sale closes.
#
# A common rule is living area. In this file overall quality moves tens of
# thousands of dollars; leaving it out will change the posted value. Partial and
# New sales sit around $275,000; the rest around $172,000. Those labels exist
# only after the deal is done. Twenty-eight neighborhoods can look good on one
# training sample and fail on the next.
#
# How the Monday tools help:
# Last week residual standard error was on the houses already in the fit. Monday:
# that training error is the wrong number to report for a house the office has
# not seen. Test error is the miss on rows the rule did not use to fit. We will
# estimate test error three ways you saw in class (ISLR Ch. 5, and Monday's
# validation hour): a validation set, leave-one-out CV, and k-fold CV. The goal
# is to lock a rule we can take to the next house, not to get a pretty fit on
# the 2,925 we already have.
#
# Description of variables (from the codebook in the case):
# - SalePrice: sale price in dollars (the outcome)
# - Gr.Liv.Area: above-grade living area in square feet
# - Overall.Qual: overall material and finish, 1 to 10
# - Full.Bath, Garage.Cars, Year.Built: listing facts
# - Neighborhood: 28 names in Ames
# - Sale.Condition, Sale.Type: recorded at closing (Partial, New)
#
# Source: De Cock (2011), Ames housing; OpenIntro extract. Five sales with
# living area above 4,000 sq ft are already out.
#
# Think: 1) What would you explore before you post a number?
#        2) What is the outcome, and which columns would you have before close?
```

```{r}
#| label: load
#############################
# Loading packages and data
#############################
ames <- read_csv("data/ames_housing.csv", show_col_types = FALSE)
ames$Garage.Cars[is.na(ames$Garage.Cars)] <- 0
ames$Kitchen.Qual[ames$Kitchen.Qual == "Po"] <- "Fa"

nrow(ames)                 # Why? Confirm n matches the case (2,925).
mean(ames$SalePrice)       # Why? The case asks you to report test MSE next to this mean.
sd(ames$SalePrice)
table(ames$Sale.Condition) # Partial is how the sale closed, not a listing fact.
```

```{r}
#| label: plot-price-area
#| fig-height: 5.0
#| fig-width: 7.8
# Before running regressions, look at the comparables.
# Why? Visuals help check whether a living-area line is even plausible, and they
# communicate findings. Color is overall quality, the large listing lever in this file.
ggplot(ames, aes(Gr.Liv.Area, SalePrice, color = Overall.Qual)) +
  geom_point(alpha = 0.35, size = 1.2) +
  geom_smooth(method = "lm", se = FALSE, color = col_red, linewidth = 0.8) +
  scale_color_viridis_c(option = "cividis", breaks = c(2, 4, 6, 8, 10)) +
  scale_y_continuous(labels = usd) +
  labs(
    x = "Above-grade living area (sq ft)",
    y = "Sale price (USD)",
    color = "Overall quality"
  ) +
  theme_lab()
# Question: Does the red line fit well? Hold living area at about 1,500 sq ft.
# A quality-5 sale and a quality-8 sale are not in the same band. Any pattern
# the line is missing?
# Discuss: The case said leaving quality out will change the posted value.
# Roughly how many dollars sit between those two bands, next to the $180,000 mean?
# The house you will post is quality 6.
```

```{r}
#| label: plot-price-hist
#| fig-height: 4.0
#| fig-width: 7.4
ggplot(ames, aes(SalePrice)) +
  geom_histogram(bins = 35, fill = col_ink, color = "white") +
  geom_vline(xintercept = mean(ames$SalePrice), color = col_red, linewidth = 0.9) +
  scale_x_continuous(labels = usd) +
  labs(x = "Sale price (USD)", y = "Number of sales") +
  theme_lab()
# Question: The red line is the mean, about $180,000. If a rule is typically
# $40,000 off on a house it did not use to fit, is that large or small here?
```

```{r}
#| label: simple-lm
#############################
# Simple linear regression
#############################
# Simple linear regression estimates the relationship between one predictor and
# the outcome. Here, how sale price changes with living area.
# Why important? This is the common rule in the case. If the slope is small next
# to $180,000, living area barely moves the posted value; if it is large, the
# office cannot ignore square footage.
m_area_all <- lm(SalePrice ~ Gr.Liv.Area, data = ames)  # SalePrice = b0 + b1 * living area + e
summary(m_area_all)
# Interpret: What does the coefficient for Gr.Liv.Area tell us in dollars per
# square foot? Is it significant? R-squared shows explained variance on these
# 2,925 sales, not on a house that is not in the file.
# Business implication: Is a living-area-only asking price good enough, or is
# the residual still huge next to the $180,000 mean?
```

```{r}
#| label: listing-lm
#############################
# Multiple linear regression
#############################
# Multiple regression includes several predictors to control for confounding.
# Why? The scatter already showed quality fanning at a given square footage.
# Isolating living area after quality is in the line is closer to how the office
# actually looks at a listing.
m_list_all <- lm(
  SalePrice ~ Gr.Liv.Area + Overall.Qual + Full.Bath + Garage.Cars + Year.Built,
  data = ames
)
summary(m_list_all)
# Interpret new coefficients: How does Overall.Qual affect sale price, in dollars
# per step, holding living area, baths, garage, and year? Did R-squared improve?
# Business implication: The case said leaving quality out will change the posted
# value. Does this coefficient agree, next to $180,000?
#
# Student exercise: Look at Full.Bath. Did the sign come out the way you would
# explain to an assessor ("add a bath, add dollars")? If not, what else in the
# line might baths be standing in for?
```

```{r}
#| label: cv-header
####################################
# Estimating the miss on a new house
####################################
# These methods assess performance on unseen sales and estimate variability.
# Reference: ISLR 2nd edition, Chapter 5 (the pages assigned for Monday) and
# Monday's hour on training MSE, test MSE, and resampling.
#
# Cross-validation evaluates how well a rule generalizes by simulating prediction
# on held-out subsets.
# Why important in business analytics? Prevents treating a good fit on the 2,925
# sales we already have as a good value for the next house. The case is exactly
# that next house.
#
# On Monday the flexibility knob was extra columns (and, in the book, polynomial
# degree). Here the same idea, in this file:
#   1. Living area only          (too simple: underfit on the scatter)
#   2. Listing facts             (living area + quality + baths + garage + year)
#   3. Listing + 28 neighborhoods (more flexible; rare streets may not travel)
# Goal: Choose the rule that minimizes estimated test error, then post a number.
# Traditional in-sample comparison (R-squared, residual standard error) does not
# assess generalization. That was last week.
```

```{r}
#| label: validation-set
# The Validation Set Approach (ISLR Ch. 5.1.1; Monday: shuffle, split, fit, score).
# Definition: Randomly split the sales into a training sample and a validation
# (test) sample. Fit on train. Evaluate MSE on validation.
# Advantages: Simple to implement.
# Disadvantages: High variability (depends on the split); potential bias if the
# training sample is small.
# Business note: Useful for a quick check, but repeat the split before you lock
# a valuation rule. The case already said one training sample is not enough.
set.seed(467)
n <- nrow(ames)
train_id <- sample.int(n, floor(0.5 * n))  # about half the sales to train
train <- ames[train_id, ]
test  <- ames[-train_id, ]
nrow(train)
nrow(test)

# Train a living-area rule. Score the sales it did not use to fit.
m_area <- lm(SalePrice ~ Gr.Liv.Area, data = train)
mean((test$SalePrice - predict(m_area, test))^2)
# The MSE formula: MSE = (1 / n_test) * sum (y_i - yhat_i)^2
# where y_i is the actual sale price, yhat_i is the predicted price, and n_test
# is the number of held-out sales.
# Why? Estimates generalization error. Lower is better out of sample.
# In this office: use it to compare rules you would take to a house not in the file.

# RMSE in dollars is easier to put next to the $180,000 mean.
sqrt(mean((test$SalePrice - predict(m_area, test))^2))

# Repeat for listing facts, then listing + neighborhood, to compare flexibility.
m_list <- lm(
  SalePrice ~ Gr.Liv.Area + Overall.Qual + Full.Bath + Garage.Cars + Year.Built,
  data = train
)
mean((test$SalePrice - predict(m_list, test))^2)
sqrt(mean((test$SalePrice - predict(m_list, test))^2))

m_nb <- lm(
  SalePrice ~ Gr.Liv.Area + Overall.Qual + Full.Bath + Garage.Cars + Year.Built + Neighborhood,
  data = train
)
nb_hat <- try(predict(m_nb, test), silent = TRUE)
if (inherits(nb_hat, "try-error")) {
  NA_real_  # a neighborhood in test was not in train, so this split cannot score 28 names
} else {
  ok <- !is.na(nb_hat)
  c(
    test_MSE = mean((test$SalePrice[ok] - nb_hat[ok])^2),
    scored = sum(ok),
    n_test = nrow(test)
  )
}
# Interpret: Did listing facts beat living area on test MSE? Did 28 names beat
# listing facts, or did predict() fail? Training residual will look better as
# you add columns. That is not the question. The question is test.
```

```{r}
#| label: plot-pred-actual
#| fig-height: 4.8
#| fig-width: 8.2
# Why? The red line is y = x, a perfect prediction on the held-out sales.
# Underfit is a cloud that fans off that line (Monday: living-area underfit).
eval_df <- rbind(
  data.frame(rule = "Living area", predicted = predict(m_area, test), actual = test$SalePrice),
  data.frame(rule = "Listing facts", predicted = predict(m_list, test), actual = test$SalePrice)
)
eval_df$rule <- factor(eval_df$rule, levels = c("Living area", "Listing facts"))
ggplot(eval_df, aes(predicted, actual)) +
  geom_abline(slope = 1, intercept = 0, color = col_red, linewidth = 0.85) +
  geom_point(alpha = 0.22, size = 1.15, color = col_ink) +
  facet_wrap(~rule) +
  coord_cartesian(xlim = c(0, 625000), ylim = c(0, 625000)) +
  scale_x_continuous(labels = usd) +
  scale_y_continuous(labels = usd) +
  labs(
    x = "Predicted sale price on the validation sample (USD)",
    y = "Actual sale price (USD)"
  ) +
  theme_lab()
# Question: Does the line fit well on the left panel? Any pattern in the
# residuals for expensive houses? Which panel would you take to the next house?
```

```{r}
#| label: plot-nbhd
#| fig-height: 5.0
#| fig-width: 7.4
# 28 neighborhoods are extra flexibility. predict() needs every test street to
# have appeared in train. Rare names sit on an axis.
nb_lev <- sort(unique(ames$Neighborhood))
nb_df <- data.frame(
  Neighborhood = nb_lev,
  n_train = as.integer(table(factor(train$Neighborhood, levels = nb_lev))),
  n_test  = as.integer(table(factor(test$Neighborhood,  levels = nb_lev)))
)
ggplot(nb_df, aes(n_train, n_test)) +
  geom_abline(slope = 1, intercept = 0, color = col_grey, linetype = "dashed", linewidth = 0.6) +
  geom_point(size = 2.3, color = col_ink) +
  geom_point(data = subset(nb_df, n_train == 0 | n_test == 0), color = col_red, size = 3.2) +
  annotate("text", x = 24, y = -2, label = "Landmrk", size = 3.3, color = col_red, hjust = 0) +
  annotate("text", x = 24, y = 36, label = "GrnHill", size = 3.3, color = col_ink, hjust = 0) +
  coord_cartesian(xlim = c(-4, max(nb_df$n_train) * 1.06), ylim = c(-10, max(nb_df$n_test) * 1.08)) +
  labs(x = "Sales in the training sample", y = "Sales in the test sample") +
  theme_lab()
# Question: Which street is on the axis with seed 467? GrnHill has two sales in
# the whole file; Landmrk has one. What happens to a 28-dummy rule on a street
# the training sample never saw?
```

```{r}
#| label: second-split
# Different split to show variability (ISLR Fig. 5.2; Monday: a second 50/50).
# Why? Demonstrates how results change with a different random split, which is
# the main drawback of the validation set.
set.seed(468)
train_id2 <- sample.int(n, floor(0.5 * n))
train2 <- ames[train_id2, ]
test2  <- ames[-train_id2, ]

mse_val <- function(form, tr, te) {
  m <- lm(form, data = tr)
  yh <- try(predict(m, te), silent = TRUE)
  if (inherits(yh, "try-error")) return(NA_real_)
  ok <- !is.na(yh)
  mean((te$SalePrice[ok] - yh[ok])^2)
}
f_area <- SalePrice ~ Gr.Liv.Area
f_list <- SalePrice ~ Gr.Liv.Area + Overall.Qual + Full.Bath + Garage.Cars + Year.Built
f_nb   <- SalePrice ~ Gr.Liv.Area + Overall.Qual + Full.Bath + Garage.Cars + Year.Built + Neighborhood

val467 <- c(
  mse_val(f_area, train, test),
  mse_val(f_list, train, test),
  mse_val(f_nb,   train, test)
)
val468 <- c(
  mse_val(f_area, train2, test2),
  mse_val(f_list, train2, test2),
  mse_val(f_nb,   train2, test2)
)
data.frame(
  rule = c("Living area", "Listing facts", "Listing + neighborhood"),
  MSE_seed467 = val467,
  MSE_seed468 = val468,
  RMSE_seed467 = sqrt(val467),
  RMSE_seed468 = sqrt(val468)
)
# Compare to previous: Different MSE? Did neighborhood even score on both splits?
# Student exercise: Change seed to 469 and run the split again.
# Discuss: How does variability affect whether you lock 28 neighborhood names?
# In this office, average over several splits before you take a rule to the next house.
#
# Problems with the Validation Set approach (ISLR; Monday):
# 1) MSE can be highly variable depending on which sales landed in training.
# 2) Performance is worse when models are trained on fewer observations, so
#    validation MSE can overestimate the test error you would see with more data.
```

```{r}
#| label: fig52
#| fig-height: 4.8
#| fig-width: 7.8
# Analog of ISLR Figure 5.2, in this file. Flexibility on the x-axis (three
# rules, simple to more flexible). Two random splits. If the two lines do not
# pick the same winner, you cannot lock from one validation set.
fig52 <- data.frame(
  flexibility = factor(
    rep(c("1 Living area", "2 Listing facts", "3 + neighborhood"), 2),
    levels = c("1 Living area", "2 Listing facts", "3 + neighborhood")
  ),
  split = rep(c("Seed 467", "Seed 468"), each = 3),
  rmse = sqrt(c(val467, val468))
)
ggplot(fig52, aes(flexibility, rmse, color = split, group = split)) +
  geom_line(linewidth = 0.9, na.rm = TRUE) +
  geom_point(size = 3, na.rm = TRUE) +
  scale_color_manual(values = c("Seed 467" = col_ink, "Seed 468" = col_red)) +
  scale_y_continuous(labels = usd) +
  labs(
    x = "Flexibility of the rule (this file)",
    y = "Validation RMSE (USD)",
    color = NULL
  ) +
  theme_lab()
# Question: What do you see? Is there consensus between the two splits about
# which rule has the lowest validation error? Monday's Figure-5.2 point: the
# quadratic often won, but not every split agreed. Same idea, Ames instead of mpg.
```

```{r}
#| label: kfold
# k-Fold Cross-Validation (ISLR Ch. 5.1.3; Monday: k = 5 or 10).
# Definition: Split the sales into k blocks (folds). Train on k-1, test on the
# held-out fold, repeat k times, average the k MSEs.
# Advantages:
# - Lower variance than LOOCV (the k models are less correlated)
# - Computationally faster (k fits vs n)
# Disadvantages:
# - Slightly higher bias than LOOCV (each training sample is smaller than n-1)
# Common k = 5 or 10; balances the bias-variance trade-off.
# In business analytics: preferred on a file this size (n = 2,925). The case
# asked you to average test MSE across several training samples; this is that
# average. The number you report is the mean of the k MSEs, not the smallest,
# not the last (Monday).
#
# How the picture works: each block is the validation sample once. Cross-validation
# error = average(Error 1, Error 2, ..., Error k).
```

```{r}
#| label: plot-kfold-schematic
#| fig-height: 3.6
#| fig-width: 5.4
kf <- expand.grid(pass = 1:5, block = 1:5)
kf$held <- kf$pass == kf$block
ggplot(kf, aes(block, 6 - pass, fill = held)) +
  geom_tile(color = "white", width = 0.92, height = 0.92) +
  scale_fill_manual(values = c("FALSE" = col_ink, "TRUE" = col_red), guide = "none") +
  coord_equal() +
  labs(x = "Block of sales", y = "Pass") +
  theme_lab() +
  theme(axis.text.y = element_blank(), panel.grid = element_blank())
# Question: On pass 3, which block is the test sample? How many times does each
# sale get to be in test across the five passes? Red is held out; ink is train.
```

```{r}
#| label: cv5-loop
set.seed(467)
fold5 <- sample(rep(1:5, length.out = n))

cv_k <- function(form, fold) {
  k <- max(fold)
  vapply(1:k, function(j) {
    tr <- ames[fold != j, ]
    te <- ames[fold == j, ]
    mse_val(form, tr, te)
  }, numeric(1))
}

cv5_area <- cv_k(f_area, fold5)
cv5_list <- cv_k(f_list, fold5)
cv5_nb   <- cv_k(f_nb,   fold5)
data.frame(
  rule = c("Living area", "Listing facts", "Listing + neighborhood"),
  mean_test_MSE = c(mean(cv5_area, na.rm = TRUE), mean(cv5_list, na.rm = TRUE), mean(cv5_nb, na.rm = TRUE)),
  mean_test_RMSE = sqrt(c(mean(cv5_area, na.rm = TRUE), mean(cv5_list, na.rm = TRUE), mean(cv5_nb, na.rm = TRUE))),
  folds_scored = c(sum(!is.na(cv5_area)), sum(!is.na(cv5_list)), sum(!is.na(cv5_nb)))
)
# Interpret: Lowest mean test error for which rule? Did every fold score for
# neighborhood? Business: choose the parsimonious rule you can actually take
# to a street that might be missing from a training sample.
```

```{r}
#| label: plot-cv5
#| fig-height: 4.6
#| fig-width: 7.6
cv5_plot <- data.frame(
  fold = rep(1:5, 3),
  rule = factor(
    rep(c("Living area", "Listing facts", "Listing + neighborhood"), each = 5),
    levels = c("Living area", "Listing facts", "Listing + neighborhood")
  ),
  rmse = sqrt(c(cv5_area, cv5_list, cv5_nb))
)
ggplot(cv5_plot, aes(fold, rmse, color = rule)) +
  geom_line(linewidth = 0.8, na.rm = TRUE) +
  geom_point(size = 2.6, na.rm = TRUE) +
  scale_color_manual(
    values = c(
      "Living area" = col_grey,
      "Listing facts" = col_ink,
      "Listing + neighborhood" = col_red
    )
  ) +
  scale_x_continuous(breaks = 1:5) +
  scale_y_continuous(labels = usd) +
  labs(x = "Held-out fold (Error 1 ... Error 5)", y = "Test RMSE (USD)", color = NULL) +
  theme_lab()
# Question: What do you see? Neighborhood has no point on a fold that could not
# score a street. The CV number is the average of the points that exist, not
# the lowest point. Would you lock 28 names from one fold that looked good?
```

```{r}
#| label: close
#############################
# Information you would not have
#############################
# Monday: leakage is using a column you would not have at the moment of the
# decision. Partial and New are recorded at closing. They came in high in this
# file (~$275,000 vs ~$172,000). That does not mean they belong in a value for
# a house that has not closed.
train$Partial <- as.integer(train$Sale.Condition == "Partial")
train$New     <- as.integer(train$Sale.Type == "New")
test$Partial  <- as.integer(test$Sale.Condition == "Partial")
test$New      <- as.integer(test$Sale.Type == "New")
ames$Partial  <- as.integer(ames$Sale.Condition == "Partial")
ames$New      <- as.integer(ames$Sale.Type == "New")

m_close <- lm(
  SalePrice ~ Gr.Liv.Area + Overall.Qual + Full.Bath + Garage.Cars + Year.Built + Partial + New,
  data = train
)
c(
  listing_test_MSE = mse_val(f_list, train, test),
  close_test_MSE = mean((test$SalePrice - predict(m_close, test))^2)
)
coef(m_close)[c("Partial", "New")]
# Interpret: Did test MSE fall when Partial and New went in? Holding listing
# facts, how many dollars is Partial? Business implication: would you have that
# label on a house the office is valuing before it sells?
```

```{r}
#| label: plot-partial
#| fig-height: 4.8
#| fig-width: 7.8
ames$partial_lab <- ifelse(ames$Sale.Condition == "Partial", "Partial (after close)", "Other")
ggplot(ames, aes(Gr.Liv.Area, SalePrice, color = partial_lab)) +
  geom_point(data = ames[ames$Sale.Condition != "Partial", ], alpha = 0.28, size = 1.15) +
  geom_point(data = ames[ames$Sale.Condition == "Partial", ], alpha = 0.8, size = 1.45) +
  scale_color_manual(values = c("Other" = col_ink, "Partial (after close)" = col_red)) +
  scale_y_continuous(labels = usd) +
  labs(x = "Above-grade living area (sq ft)", y = "Sale price (USD)", color = NULL) +
  theme_lab()
# Question: At the same living area, do the red points still sit above the ink
# cloud, or is Partial only "newer, larger houses"? Would you have the red
# label before the sale closes?
```

```{r}
#| label: post
#############################
# Post the number
#############################
# Last in-class block is the decision in the case. Lock a rule you can defend
# with estimated test error, using only facts before close. Then predict.
new_house <- data.frame(
  Gr.Liv.Area = 1440,
  Overall.Qual = 6,
  Full.Bath = 2,
  Garage.Cars = 2,
  Year.Built = 1973
)
m_lock <- lm(f_list, data = train)  # listing facts; 28 names and Partial/New stay out
posted <- as.numeric(predict(m_lock, new_house))
posted
sqrt(mse_val(f_list, train, test))  # test RMSE next to the $180,000 mean
mean(ames$SalePrice)
# Interpret: Why is the post not the mean of the file? Next to $180,000, is the
# typical miss on held-out sales small enough to post without saying how wide
# the miss is? Did neighborhood dummies go into this lock? Did Partial?
```

```{r}
#| label: plot-posted
#| fig-height: 4.2
#| fig-width: 7.4
ggplot(ames, aes(SalePrice)) +
  geom_histogram(bins = 35, fill = col_ink, color = "white") +
  geom_vline(xintercept = mean(ames$SalePrice), color = col_red, linewidth = 0.9) +
  geom_vline(xintercept = posted, color = col_grey, linewidth = 0.95, linetype = "dashed") +
  annotate("text", x = posted, y = 505, label = "Posted", hjust = 1.12, size = 3.4, color = col_grey) +
  annotate("text", x = mean(ames$SalePrice), y = 505, label = "Mean", hjust = -0.12, size = 3.4, color = col_red) +
  coord_cartesian(ylim = c(0, 530)) +
  scale_x_continuous(labels = usd) +
  labs(x = "Sale price (USD)", y = "Number of sales") +
  theme_lab()
```

# After class

The code is already typed.

```{r}
#| label: loocv
# LOOCV (Leave-One-Out Cross-Validation) - ISLR Ch. 5.1.2; Monday: n fits, or
# the OLS shortcut.
# Definition: For n sales, train on n-1, test on the held-out one, repeat n
# times, average the n MSEs.
# Advantages:
# - Low bias (trains on almost all the data)
# - No randomness in the splits
# Disadvantages:
# - High variance (the n models are highly correlated)
# - Computationally expensive if you really refit n times (here n = 2,925)
# Why use? Nearly unbiased estimate of test error. For lm, ISLR p. 193: you do
# not run 2,925 regressions. One fit and the leverages (hatvalues) give the
# same number. Neighborhood with a one-sale street is a bad place for this
# shortcut; we do listing facts and living area only.
m_area_all <- lm(f_area, data = ames)
h_area <- hatvalues(m_area_all)
loocv_area <- mean(((ames$SalePrice - fitted(m_area_all)) / (1 - h_area))^2)

m_list_all2 <- lm(f_list, data = ames)
h_list <- hatvalues(m_list_all2)
loocv_list <- mean(((ames$SalePrice - fitted(m_list_all2)) / (1 - h_list))^2)

c(living_area_LOOCV = loocv_area, listing_LOOCV = loocv_list)
sqrt(c(living_area_LOOCV = loocv_area, listing_LOOCV = loocv_list))
# Interpret: Lowest for listing facts? Compare to the one validation split.
# Business: a nearly unbiased number, no shuffle. Still not a reason to put
# Partial in the rule.
```

```{r}
#| label: tenfold
# 10-fold CV, same loop as 5-fold. Student exercise in L4: try K = 5. We already
# ran K = 5 in class; this is K = 10 on the same three rules.
set.seed(467)
fold10 <- sample(rep(1:10, length.out = n))
cv10_area <- cv_k(f_area, fold10)
cv10_list <- cv_k(f_list, fold10)
cv10_nb   <- cv_k(f_nb,   fold10)
data.frame(
  rule = c("Living area", "Listing facts", "Listing + neighborhood"),
  mean_test_MSE = c(mean(cv10_area, na.rm = TRUE), mean(cv10_list, na.rm = TRUE), mean(cv10_nb, na.rm = TRUE)),
  folds_scored = c(sum(!is.na(cv10_area)), sum(!is.na(cv10_list)), sum(!is.na(cv10_nb)))
)
# Interpret: Minimum still listing facts, or do 28 names win once you average
# ten blocks? Do all ten folds score? Compare to LOOCV: similar ranking, less
# drama than the two 50/50 splits.
# Student exercise: Does the rule you would lock change between K = 5 and K = 10?
#
# Bias-variance trade-off between LOOCV and k-fold (ISLR Ch. 5.1.3; Monday):
# LOOCV has low bias (trains on n-1) but high variance (n highly correlated models).
# k-fold (k = 10) has slightly higher bias (trains on 9/10 of the sales) but lower
# variance (10 less correlated models).
# Why prefer k-fold on this file? More stable test-error estimates, less risk of
# locking 28 names because one split looked good.
```

```{r}
#| label: fig54
#| fig-height: 5.0
#| fig-width: 8.0
# Analog of ISLR Figure 5.4: estimated test error against flexibility, for
# more than one resampling method. The book uses polynomial degree; this file
# uses the three rules in the case.
fig54 <- rbind(
  data.frame(
    flexibility = c("1 Living area", "2 Listing facts", "3 + neighborhood"),
    method = "Validation 467",
    rmse = sqrt(val467)
  ),
  data.frame(
    flexibility = c("1 Living area", "2 Listing facts", "3 + neighborhood"),
    method = "Validation 468",
    rmse = sqrt(val468)
  ),
  data.frame(
    flexibility = c("1 Living area", "2 Listing facts", "3 + neighborhood"),
    method = "5-fold",
    rmse = sqrt(c(mean(cv5_area, na.rm = TRUE), mean(cv5_list, na.rm = TRUE), mean(cv5_nb, na.rm = TRUE)))
  ),
  data.frame(
    flexibility = c("1 Living area", "2 Listing facts", "3 + neighborhood"),
    method = "10-fold",
    rmse = sqrt(c(mean(cv10_area, na.rm = TRUE), mean(cv10_list, na.rm = TRUE), mean(cv10_nb, na.rm = TRUE)))
  )
)
fig54$flexibility <- factor(
  fig54$flexibility,
  levels = c("1 Living area", "2 Listing facts", "3 + neighborhood")
)
fig54$method <- factor(
  fig54$method,
  levels = c("Validation 467", "Validation 468", "5-fold", "10-fold")
)
ggplot(fig54, aes(flexibility, rmse, color = method, group = method)) +
  geom_line(linewidth = 0.9, na.rm = TRUE) +
  geom_point(size = 2.8, na.rm = TRUE) +
  scale_color_manual(
    values = c(
      "Validation 467" = col_grey,
      "Validation 468" = "#8AA8C4",
      "5-fold" = col_ink,
      "10-fold" = col_red
    )
  ) +
  scale_y_continuous(labels = usd) +
  labs(
    x = "Flexibility of the rule (this file)",
    y = "Estimated test RMSE (USD)",
    color = NULL
  ) +
  theme_lab()
# Question: What do you see in this graph? The two validation-set lines jump.
# The 5-fold and 10-fold lines sit closer. Living area is high on every method.
# Neighborhood jumps or goes missing. Which rule would you lock for the house
# that is not in the file?
# Discuss: What does all of this have to do with overfitting? Extra columns can
# look good on train and on a lucky split. Averaging several held-out blocks is
# how Monday told you to pick complexity. The case asked for that average
# before you post.
```

```{r}
#| label: baths
# Full.Bath on the listing rule can change sign. That is not "add a bath, add
# dollars." It is baths once living area and quality are already in the line.
summary(m_list)$coefficients["Full.Bath", ]
aggregate(Overall.Qual ~ Full.Bath, data = ames, FUN = mean)
```

```{r}
#| label: plot-baths
#| fig-height: 4.6
#| fig-width: 7.6
bath_df <- subset(ames, Full.Bath %in% c(1, 2, 3))
bath_df$baths <- factor(paste0(bath_df$Full.Bath, "-bath"), levels = c("1-bath", "2-bath", "3-bath"))
ggplot(bath_df, aes(factor(Overall.Qual), SalePrice, fill = baths)) +
  geom_boxplot(
    position = position_dodge(width = 0.75),
    outlier.alpha = 0.2, width = 0.65, color = col_ink
  ) +
  scale_fill_manual(values = c(col_ink, col_red, col_grey)) +
  scale_y_continuous(labels = usd) +
  labs(x = "Overall quality", y = "Sale price (USD)", fill = NULL) +
  theme_lab()
# Interpret: Inside a quality score, do two-bath boxes sit above one-bath boxes,
# or do they overlap? Business implication: if they overlap, a negative Full.Bath
# coefficient once quality is in the line is not a puzzle.
```
